### 2017-02-15 pedestrian_forecasting_iros_2017.tex
Here's the paper repo

\documentclass[letterpaper,10pt,conference]{ieeeconf}
%\usepackage{times}

\IEEEoverridecommandlockouts 
\overrideIEEEmargins   % Needed to meet printer requirements.
\usepackage[T1]{fontenc}

%  For math environment 
\usepackage{amsmath,amssymb}
%\usepackage[linesnumbered]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{caption}
%  For todo notes
\usepackage{todonotes}
%  For drawing Bayesian networks
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage{tikz-cd}
\usepackage[justification=justified,font=footnotesize]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{array}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\usepackage{comment}
\usepackage{color}  % For Highlighting

%  For theorems, etc
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
%\usepackage{multicol}
\usepackage[bookmarks=true]{hyperref}
\newcommand{\Ram}[1]{\textcolor{red}{#1}}
\newcommand{\Hoj}[1]{\textcolor{blue}{#1}}

% Custom commands
\newcommand{\pder}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

\begin{document}

% paper title
\title{Real-Time Certified Probabilistic Pedestrian Forecasting}

%%% first author
\author{Henry O. Jacobs, Owen Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
 \thanks{ H.O. Jacobs, O. Hughes, and R.~Vasudevan are with the Department of Mechanical Engineering, University of Michigan, Ann Arbor, MI 48109
{\scriptsize \{\texttt{hojacobs,owhughes,ramv}\}~\texttt{@umich.edu}}}
\thanks{M. Johnson-Roberson is with the Department of Naval and Marine Engineering, University of Michigan, Ann Arbor, MI 48109\newline
{\scriptsize \texttt{mattjr@umich.edu}}}
}

% You will get a Paper-ID when submitting a pdf file to the conference system
%\newcommand{\umich}{ \begin{affiliation} Dept. of Mechanical Eng. \\ University of Michigan \\ Ann Arbor, Michigan \end{affiliation}}
%\author{
%	Henry O. Jacobs \\
%	\umich \\
%	\email{hojacobs@umich.edu}
%	\and
% 	Owen Hughes \\
%	\umich \\
%	\email{owhughes@umich.edu}
% 	\and Ram Vasudevan \\
%	\umich \\
%	\email{ramv@umich.edu} \\
%}

\maketitle

\begin{abstract}
That autonomous vehicles will come to dominate our streets is imminent.
This motivates the need for a real-time probabilistic forecasting algorithm for pedestrians, cyclists, and other agents,
as it forms a necessary step in assessing the risk (and therefore the cost) we should expect to incur.
In this paper, we present a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations.
The resulting algorithm is embarrassingly parallel, and trained on historical trajectory information within a fixed scene.
When compared with MDP-based methods, our algorithm appears to be superior from the standpoint of precision and recall.
\end{abstract}

%\IEEEpeerreviewmaketitle

\section{Introduction}

Autonomous systems are increasingly being deployed in and around humans. 
The ability to accurately model and anticipate human behavior is critical to maximizing safety, confidence, and effectiveness of these systems in human-centric environments. 
The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of an action over a set of possible behaviors. 
Since the set of plausible human behaviors is vast, this work focuses on anticipating the possible future locations of people within a bounded area. 
This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).

With a particular focus on the AV application several additional design criteria become important. 
First, false negative rates for unoccupied regions must be minimized. 
The misclassification of space in this way has obvious safety issues. 
Second, speed is paramount. 
To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories. 
The margins between humans and vehicles are small and both frequently operate at the boundary. 
Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.
This paper presents an algorithm for real-time, long-term prediction of pedestrian behavior which can subsequently be used by autonomous agents.
As depicted in Figure \ref{fig:gates-1-2}, this method works quickly to generate predictions that are precise while reducing the likelihood of false negative detections.

\begin{figure}
	\centering
	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=55.png}
		\caption{$t=1.83s$}
	\end{subfigure}
		\begin{subfigure}[b]{.45\linewidth}
			\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=145.png}
			\caption{$t=4.83$}
		\end{subfigure}

	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=235.png}
		\caption{$t=7.83s$}
	\end{subfigure}
	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=345.png}
		\caption{$t=11.5s$}
	\end{subfigure}
	\caption{The performance of the presented algorithm captures the most probable routes that a pedestrian chooses. In this example which was captured at thirty frames per second, the presented algorithm took 0.0158s per frame in a Python implementation.  In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.}
	\label{fig:gates-1-2}
\end{figure}

Autonomous systems are increasingly being deployed in and around humans.
 The ability to accurately model and anticipate human behavior is of great value to maximize safety, confidence, and effectiveness in human-centric environments. 
 The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of possible action over a set a of possible behaviors. 
 Since the set of plausible behaviors in arbitrary situations may be vast, this work constrains the problem to anticipating the possible future locations of people within a bounded area. This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).

 With a particular focus on the AV application several additional design criteria become important. 
 First, false negative rates for unoccupied regions must be minimized. 
 The misclassification of space in this way has obvious safety issues. 
 Second, speed is paramount. 
 To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories. 
 The margins between humans and vehicles are small and both frequently operate at the boundary. 
 Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.

This paper presents an algorithm for real-time, long-term prediction of pedestrian behavior which can subsequently be used by autonomous agents.
As depicted in Figure \ref{fig:}, this approach relies upon ...

%First, since safety is paramount, generated predictions must avoid misclassifying areas as unoccupied even though humans are present. 
%Though this may unduly reduce the precision of the algorithm

\subsection{Background}
Most forecasting algorithm are well characterized by the underlying evolution model they adopt.
Such models come in a variety flavors, and are adapted to the task at hand (e.g. crowd modeling \cite{Helbing1992}).
This paper is focused on the construction of useful motion models for pedestrians that can aide the task of real-time forecasting for autonomous agents.
The simplest approach to forecasting with motion models forward integrates a Kalman filter \cite{kalman1960new} based upon the observed heading.
Over short time scales this method may perform well, but the resulting distribution devolves into an imprecise Gaussian mass over longer time scales.
This is a typical property of the extended Kalman filter and the Kalman filter with deterministic nonlinear drift \cite{Schneider2013}.
In particular, such models are less useful for forecasts beyond two seconds, especially when a pedestrian turns.
Nonetheless, these stochastic linear models serve as a reasonable default in the absence of any contextual knowledge.
For example, \cite{Kooji2014} uses a motion model of this variety, parametrized by latent variables, such as the pedestrian's awareness of nearby vehicles, to predict when and how a pedestrian may cross a street without leveraging any environmental context.

More sophisticated models that attempt to leverage environmental data include inverse optimal control (IOC) based models \cite{Ziebart2008,Ziebart2009,Kitani2012,Xie2013,Karasev2016}.
These IOC models have the desirable property of attributing intention and goal-seeking behavior to the agents.
For example, \cite{Kitani2012} extracts a Markov Decision Process (MDP) evolving on a finite 2D lattice by training on a small collection of features and trajectories.
The resulting MDP is light-weight since it is parametrized by only a small collection of coefficients equal in number to that of the feature maps.
Moreover, given an initial and final state, the probability distribution of the pedestrian at intermediate states is computable using matrix multiplication.
The speed of this computation, makes the algorithm of \cite{Kitani2012} a reasonable baseline for comparison for the algorithm that is presented in this paper.

The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
For example, time-dependent information, such as traffic signals, are incorporated in \cite{Karasev2016}, by relaxing the Markov property and considering a switched Markov process.
Other generalizations in \cite{Karasev2016} include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method: a Rao-Blackwellized particle filter~\cite{Doucet2000} is employed.
The resulting accuracy of such methods can only be known in a probabilistic sense in that the error bounds are themselves random variables.
Moreover, accuracy can come at a large computational expensive which is prohibitive during real-time control.

One limitation of IOC models occurs in cases where there are many locally optimal solutions between a start and end goal.
In these cases, IOC type methods can yield non-robust and imprecise behavior.
This can occur, for example, when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch based on its feature and the behavior of previous agents on similar patches.
The motion model is a Markov jump process and the relevant posteriors were approximated using sample based techniques similar to \cite{Karasev2016}.
The objective of \cite{Ballan2016} is not only prediction, but the development of a motion model learned on one scene that could then subsequently be transferred to other scenes. 
This requires representations of ``objects'' in the scene that do not depend rigidly on the finitely many labels an engineer managed to think of in a late-night brainstorming session.
For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.

Along similar lines to \cite{Ballan2016}, \cite{Walker2014} constructed an unsupervised approach towards forecasting.
As before, the motion model was a Markov jump process, and the training set was a collection unlabeled videos.
Unlike all the approaches mentioned thus far, the agents in \cite{Walker2014} were not manually specified.
They were learned by detecting which sort of patches of video were likely to move, and how.
The resulting predictions outperformed \cite{Kitani2012} when comparing the most likely path with the ground truth using the mean Hausdorff distance.
As in all methods mentioned thus far, computational speed and accuracy of any predicted posteriors were not a concern of \cite{Walker2014}, so no such results were reported.
However, since the motion model was a Markov jump process which required the application of a sample based technique, we should expect the same painful trade-off between error and speed to occur as in \cite{Karasev2016} and \cite{Ballan2016}.

\subsection{Contributions}

The primary contributions of this paper are:
\begin{itemize}
	\item An accurate motion model for pedestrian forecasting.
	\item An expedient method for computing approximations of relevant posteriors generated by our motion model.
	\item Hard error bounds on the proposed approximations.
\end{itemize}

For clarification, we should mention that there are a number of things that we do not do.
For example, we do not concern ourselves with detection and tracking. 
Nor do we concern ourselves with updating our prediction as new data comes along \cite{Kitani2012}.
We largely work in a 2D environment with a bird's eye view, operating under the assumption that the data has been appropriately transformed by a third party.
This is in contrast to \cite{Karasev2016}, which operates from first person video.
While it would be a straight forward extension to consider such things, it would detract from the presentation.
%Such questions are important, but beyond the scope of this paper, which concerns the underlying dynamic model.

The rest of the paper is organized as follows:
\begin{itemize}
	\item \S \ref{sec:model} describes our motion model as a Bayesian network.
	\item \S \ref{sec:efficient} describes how to compute probability densities for an agent's position efficiently.
	\item \S \ref{sec:implementation} demonstrates the model by training and testing it on the Stanford drone dataset \cite{Robicquet2016}.
\end{itemize}

\section{Model}\label{sec:model}
This paper's goal is to generate a time-dependent probability density over $\mathbb{R}^2$, which predicts the true location of an agent in the future.
The input to the algorithm at runtime is a noisy measurement of position and velocity, $\hat{x}_0, \hat{v}_0 \in \mathbb{R}^2$.
If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$.

To numerically compute $\rho_t$, we build a probabilistic graphical model.
Our model assumes we have noisy information about agents, and each agent moves with some intention through the world in a way that is roughly approximated by a model.
Our model can be divided into three parts:
\begin{enumerate}
	\item Reality: This is parametrized by the true position for all time, $x_t$, and the initial velocity of the agent $v_0$.
	\item Measurements:  This is represented by our sensor readings $\hat{x}_0$ and $\hat{v}_0$ and are independent of all other variables given the true initial position and velocity, $x_o, v_0$.
	\item Motion Model: This is represented by a trajectory $\check{x}_t$ and depends on a variety of other variables which are described below.
\end{enumerate}
We now elaborate on these three components, and relate them to one another.
Many of the choices we make are motivated by a balance between model quality and computational speed (see \S \ref{sec:efficient}).

\subsection{The Variables of the Model}
The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,T]$ for some $T>0$.
We denote the position and velocity at time $t=0$ by $x_0$ and $v_0$ respectively.
At $t=0$, we obtain a measurement of position and velocity, denotes by $\hat{x}_0$ and $\hat{v}_0$.
Lastly, we have a variety of motion models, parametrized by a set $\mathcal{M}$ (described in the sequel).
For each model $m \in \mathcal{M}$, a trajectory $\check{x}_t$ given the initial position and velocity $x_0$ and $v_0$.
All these variables are probabilistically related to one another in a (sparse) Bayesian network, which we will describe next.

\subsection{The Sensor Model}
At time $t=0$, we obtain a noisy reading of position, $\hat{x}_0 \in \mathbb{R}^2$.
We assume that given the true position, $x_0 \in \mathbb{R}^2$, that the measurement $\hat{x}_0$ is independent of all other variables and the posterior $\Pr( \hat{x}_0 \mid x_0)$ is known.
We assume a similar measurement model for the measured initial velocity $\hat{v}_0$.

\subsection{The Agent Model}
All agents are initialized within some rectangular region $D \subset \mathbb{R}^2$.
We denote the true position of an agent by $x_t$.
We should never expect to know $x_t$ and the nature of its evolution precisely, and any model should account for its own (inevitable) imprecision.
We do this by fitting a deterministic model to the data and then smoothing the results.
Specifically, our motion model consists of a modeled trajectory $\check{x}_t$, which is probabilistically related to the true position by $x_t$ via a known and easily computable posterior, $\Pr(x_t \mid \check{x}_t)$.

Once initialized, agents come in two flavors: linear and nonlinear.
The linear agent model evolves according to the equation $\check{x}_t = x_0 + t v_0$ and so we have the posterior:
\begin{align}
	\Pr( \check{x}_t  \in A \mid x_0, v_0, lin) = \int_A \delta( \check{x}_t - x_0 - t v_0 ) d\check{x}_t.
\end{align}
for all measurable sets $A \subset \mathbb{R}^2$, where $\delta( \cdot )$ denotes the multivariate Dirac-delta distribution.
For the sake of convenience, from here on we drop the set $A$ and the integral when defining such posteriors since this equation is true for all measurable sets $A$.
We also assume the posteriors, $\Pr(x_0 \mid lin)$ and $\Pr( v_0 \mid lin, x_0)$ are known.

If the agent is of nonlinear type, then we assume the dynamics take the form:
\begin{align}n
	\frac{d \check{x}_t}{dt} = s \cdot X_k(\check{x}_t) \label{eq:ode}
\end{align}
where $X_k$ is a vector-field\footnote{A vector-field, generally speaking, is an assignment of a velocity to each position in some space.  A vector-field on $\mathbb{R}^n$ is nothing but a map from $\mathbb{R}^n \to \mathbb{R}^n$.} drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
More specifically, we assume that each $X_k$ has the property that $\| X_k(x) \| = 1$ for all $x \in D$.
This property ensures that speed is constant in time
% and it has the further advantage of being the (local) generator of solutions to an optimal navigation problem (see Appendix \ref{app:ioc}).
As we describe in \S \ref{sec:implementation}, the stationary vector-fields $X_1,\dots,X_n$ are learned from the dataset.

It is assumed that $k$ and $s$ are both constant in time, so that $\check{x}_t$ is determined from the triple $(x_0,k,s)$ by integrating \eqref{eq:ode} with the initial condition $x_0$.
This insight allows us to use the motion model to generate the posterior for $\Pr( \check{x}_t \mid x_0, k, s)$.
For each initial condition, $x_0$, we can solve \eqref{eq:ode} as an initial value problem, to obtain a point $\check{x}_t$ with initial condition $\check{x}_0 = x_0$.
This process, which takes $\check{x}_0$ and outputs $\check{x}_t$, constitutes a map which is termed the flow-map \cite[Ch 4]{MTA}, and which we denote by $\Phi_{k,s}^t$.
Explicitly, we have the posterior:
\begin{align}
	\Pr( \check{x}_t \mid x_0 , k , s) = \delta( \check{x}_t - \Phi^{t}_{k,s}( x_0) )  d\check{x}_t \label{eq:x_check | ksx}
\end{align}
where $\Phi^{t}_{k,s}$ is the flow-map of the vector field $s \,X_k$ up to time $t$.
Note that this flow-map can be evaluated for an initial condition efficiently by just integrating the vector field from that initial condition.
Note the variables $k,s$ and $x_0$ determine $v_0$.
Thus we have the posterior:
\begin{align}
	\Pr( v_0 \mid k, s, x_0) = \delta( v_0 -s X_k( x_0) ) dv_0. \label{eq:v | ksx}
\end{align}
In summary, the agent models are parametrized by the set:
\begin{align}
	\mathcal{M} = \{ lin \} \cup \left( \mathbb{R} \times \{ 1 , \dots, n \} \right) \label{eq:models}
\end{align}
and each flavor determines the type of motion we should expect from the agent model.

\subsection{The Full Model}
Concatenating the measurement model with our motion models yields the Bayesian network, where $M \in \mathcal{M}$ denotes the model of the agent:
\begin{align}
\begin{tikzpicture}[thick, var/.style={circle,draw,thin,rounded corners,shade,top color=blue!50,minimum size = 4mm}]
	\node[var] (M) {$M$};
	\node[var] (x)[right=of M] {$x_0$};
	\node[var] (v)[below=of x] {$v_0$};
	\node[var] (x_hat) [right=of x] {$\hat{x}_0$};
	\node[var] (v_hat) [right=of v] {$\hat{v}_0$};
	\node[var] (x_check_t) [left=of v] {$\check{x}_t$};
	\node[var] (x_t) [left=of x_check_t] {$x_t$};
	\draw[->] (M) to (x);
	\draw[->] (M) to (v);
	\draw[->] (M) to (x_check_t);
	\draw[->] (x) to (x_hat);
	\draw[->] (x) to (x_check_t);
	\draw[->] (x) to (v);
	\draw[->] (v) to (x_check_t);
	\draw[->] (v) to (v_hat);
	\draw[->] (x_check_t) to (x_t); 
\end{tikzpicture}.\label{eq:pgm}
\end{align}
We may use this Bayesian network to compute $\rho_t$ efficiently.
In particular
\begin{align}
	\rho_t(x_t ) &:= \Pr( x_t \mid \hat{x}_0, \hat{v}_0 ) \\
	&= \left( \sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds \right) \\
	&\quad + \Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 ). \label{eq:decomposition}
\end{align}
The final term $\Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 )$ is expressible in closed form when the posteriors $\Pr( x_0 \mid lin)$ and $\Pr( v_0 \mid lin,x_0)$ have known expressions ( e.g. Gaussians or uniform distributions). 
In this instance, the numerical computation of $\Pr(x_t, lin \mid \hat{x}_0, \hat{v}_0)$ poses a negligible burden.
Instead the primary computational burden derives from computing $\sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds$.

\section{Efficient Probability Propagation} \label{sec:efficient}
As mentioned, many of the modeling choices are born out of a balance between accuracy and real-time computability.
One of the major modeling choices, that agents move approximately according to a small number of ODEs, is the most prominent such choice.
This section details how this modeling choice can be leveraged to compute $\rho_t(x_t)$ quickly and accurately.

To begin, rather than focusing on computing $\rho_t(x_t)$, we describe how to compute the joint probability $\Pr( x_t , \hat{x}_0, \hat{v}_0)$.
We can obtain $\rho_t(x_t)$ by normalizing $\Pr( x_t , \hat{x}_0, \hat{v}_0)$ with respect to integration over $x_t$.
We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on the support of $\Pr(s)$, 
we can conclude that the integral term in \eqref{eq:decomposition} can be approximated by
\begin{align}
	\begin{split}
	&\sum_{k} \int \Pr( x_t, k , s , \hat{x}_0, \hat{v}_0 ) ds = \\
	 &\underbrace{\Delta s \sum_{j} \sum_{k} \Pr( x_t, k , s_j , \hat{x}_0, \hat{v}_0)}_{\text{approximation}}
	  +
	  \underbrace{\varepsilon_s}_{\text{error}}
	 \end{split} \label{eq:approximation 0}
\end{align}
where the error is bounded by $\int| \varepsilon_s | ds \leq  TV(x_t, \hat{x}_0, \hat{v}_0) \Delta s$
where $TV(x_t, \hat{x}_0, \hat{v}_0)$ is the sum, with respect to $k$, of the total variation of $\Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 )$ with respect to $s$ for fixed $x_t, \hat{x}_0, \hat{v}_0$.
Since this error term can be controlled, the problem of solving $\rho_t(x_t)$ is reduced to that of efficiently computing $ \Pr( x_t, k , s_j  \mid \hat{x}_0, \hat{v}_0)$
for a fixed collection of $s_j$'s.

$\hat{x}_0$ and $\hat{v}_0$ are measured and are assumed fixed for the remainder of this section.
%The only continuous variable we must discretize is $x_t$.
To begin, from \eqref{eq:pgm}  notice that:
\begin{align}
	&\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0) =\\
	&= \int \Pr( x_t, \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \\
	&= \int \Pr( x_t \mid \check{x}_t ) \Pr(\check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \label{eq:convolve}
\end{align}
Observe that from the last line that $\Pr( x_t, k,s, \hat{x}_0, \hat{v}_0)$ is a convolution of the joint distribution $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$.
Assuming, for the moment, that such a convolution can be performed efficiently, we focus on computation of $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$.

Again, \eqref{eq:pgm} implies:
\begin{align}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \\
	&= \int \Pr( \check{x}_t , x_0, \hat{x}_0, v_0, \hat{v}_0, k,s) dx_0 \, dv_0 \\
	&= \int \Pr( \check{x}_t \mid  x_0, k,s, v_0) \Pr( \hat{v}_0 \mid v_0) \cdot \\
	& \hspace{10ex}  \cdot \Pr( v_0 \mid k,s,x_0) \Pr(\hat{x}_0, x_0, k, s) dx_0 \, dv_0 \nonumber \\
		&= \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right) \delta\left( v_0 - s X_k (x_0) \right) \cdot \\
		&\hspace{10ex} \cdot\Pr( \hat{v}_0 \mid v_0) \Pr(\hat{x}_0, x_0,k, s) dx_0\, dv_0, \nonumber
\end{align}
where the last equality follows from substituting \eqref{eq:x_check | ksx} and \eqref{eq:v | ksx}.
 Carrying out the integration over $v_0$ we observe:
\begin{align}
\begin{split}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right)  \cdot \\
	&\hspace{10ex} \cdot \Pr(\hat{x}_0, x_0, k, s) \Psi(\hat{v}_0 ;k,s,x_0) dx_0,
\end{split}
\label{eq:push forward}
\end{align}
 where $\Psi( \hat{v}_0 ;k,s,x_0) := \left. \Pr( \hat{v}_0 \mid v_0) \right|_{v_0 = s X_k(x_0)}$.
 Assuming the density $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0)$ is of bounded variation in the variable $x_0$ (with all other variables held fixed),
 we may approximate it as a sum of weighted Dirac-delta distributions supported on a regular grid.
 %This is merely the definition of \emph{Riemann integrable} \cite{Rudin1991}.

 To accomplish this, let $S_L(\hat{x}_0)$ denote the square of side length $L>0$ centered around $\hat{x}_0$.
 Choose  $L>0$ to be such that $\int_{S_L(\hat{x}_0)} \Pr( x_0 \mid \hat{x}_0) dx_0 = 1 - \varepsilon_{tol}$ for some error tolerance $\varepsilon_{tol}>0$.
 Then, for a given resolution $N_x \in \mathbb{N}$ define the regular grid on $S_L(\hat{x}_0)$ as $\Gamma_L( \hat{x}_0 ; N_x ) := \left\{ x_0^{i,j} \mid i,j = -N_x,\dots,N_x \right\}$, where $x_0^{i,j} = \hat{x}_0 + \frac{L}{2N_x}(i,j)$.
 The grid spacing is given by $\Delta x = ( \frac{L}{2N_x}, \frac{L}{2N_x} )$.
Approximate the smooth distribution $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ as a weighted sum of Dirac-deltas (in the variable $x_0$) supported on $\Gamma_L( \hat{x}_0;N)$:
\begin{align}
	\begin{split}
		&\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 ) = \\
		&\qquad \underbrace{\left( \sum_{i,j=-N}^{N} W(k,s,i,j,\hat{x}_0) \delta( x_0 - x_0^{i,j} ) \right)}_{\text{approximation}}
		+ \underbrace{ \varepsilon_0(x_0) }_{\text{error}}
	\end{split}
	\label{eq:approximation 1}
\end{align}
where $W(k,s,i,j,\hat{x}_0)$ is merely the evaluation of $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ at the grid point $x_0 = x_0^{i,j} \in \Gamma( \hat{x}_0; N)$.
%That is to say
%\begin{align}
%	\begin{split}
%	&W(k,s,i,j,\hat{x}_0) =\\
%	&\quad \left. \left[ \Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0; k,s,x_0 )  | \Delta x | \right] \right|_{x_0 = x_0^{i,j}} 
%	\end{split}\label{eq:constants} 
%\end{align}
For fixed $k$ and $s$, the expression $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ is a density in $x_0$ and the error term in \eqref{eq:approximation 1} has a magnitude of $\| \varepsilon_0 \|_{L^1} \sim \mathcal{O}( | \Delta x |  + \varepsilon_{tol} )$ with respect to the $L^1$-norm in $x_0$.
%In particular, a feasible constant of proportionality would be the BV-norm.
%The function $W(k,s,i,j, \hat{x}_0)$, can be computed efficiently as product of the posteriors represented in our Bayesian network \eqref{eq:pgm}.
Substitution of \eqref{eq:approximation 1} into the final line of \eqref{eq:push forward} yields:
\begin{align}
	\begin{split}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \\
	&\quad \sum_{i,j} W(k,s,i,j,\hat{x}_0) \delta \left( \check{x}_t - \Phi_{k,s}( x_0^{i,j}) \right) + \varepsilon_t( \check{x}_t)
	\end{split}
\end{align}
where $\varepsilon_t( \check{x}_t) = \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right)  \varepsilon_0(x_0) dx_0$.
The first term is computable by flowing the points of the grid, $\Gamma_L(\hat{x}_0; N_x)$, along the vector field $s X_k$.
The second term, $\varepsilon_t$, may be viewed as an error term.
In fact, this method of approximating $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$ as a sum of Dirac-delta distributions is adaptive, in that the error term does not grow in total mass, which is remarkable since many methods for linear evolution equations accumulate error exponentially in time \cite{Leveque1992,Gottlieb2001}:

\begin{thm} \label{thm:error}
	The error term, $\varepsilon_t \sim \mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ in the $L^1$-norm, for fixed $k,s,\hat{x}_0$, and $\hat{v}_0$.
	Moreover, $\| \varepsilon_t \|$ is constant in time.
\end{thm}
\begin{proof}
	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
	We observe
\begin{align*}
	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
\end{align*}
As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
\end{proof}

While this allows us to compute posteriors over the output of our models, $\check{x}_t$, we ultimately care about densities over the true position.
The following corollary of Theorem \ref{thm:error} addresses this:
\begin{cor} \label{cor:error}
	The density
	\begin{align}
		\sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^\alpha) } \label{eq:approximation 2}
	\end{align}
	is an approximation of $\Pr( x_t, k, s, \hat{x}_0, \hat{v}_0)$ with a constant in time error bound of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$.
\end{cor}
\begin{proof}
	By \eqref{eq:convolve}
	\begin{align*}
		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
	\end{align*}
	Substitution of \eqref{eq:approximation 1} yields
	\begin{align*}
		\begin{split}
		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
		\end{split}
	\end{align*}
	where the error term is
	\begin{align}
		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
	\end{align}
	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
	\begin{align}
		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
	\end{align}
	Implementing the integration over $x_t$ first yields:
	\begin{align}
		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
	\end{align}
	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
\end{proof}

Corollary \ref{cor:error} justifies using \eqref{eq:approximation 2} as an approximation of $\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0)$.
In particular, this reduces the problem of computing $\rho_t(x_t)$ to the problem of computing the weights $W(k,s,i,j,\hat{x}_0)$ and the points $\Phi_{k,s}^t(x_0^{i,j})$ for all $k,s$ and points $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
We can reduce this burden further by exploiting the following symmetry:
\begin{thm} \label{thm:symmetry}
	$\Phi_{k,s}^t = \Phi_{k,1}^{st}$.
\end{thm}
\begin{proof}
	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
\end{proof}
This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$.
This yields the following theorem
\begin{thm} \label{thm:main}
	Let $\{s_1,\dots,s_n\}$ be a regular partition on the support of $\Pr(s)$.
	Then the density
	\begin{align}
		\begin{split}
		&\Delta s \sum_{i,j,k,\ell} W(k,s_\ell,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{s_\ell t}( x_0^{i,j}) } \\
		&+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 )
		\end{split}
		\label{eq:approximation 3}
	\end{align}
	approximates $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ with an error of size $\mathcal{O}( \Delta s + \Delta x + \varepsilon_{\rm tol})$.
\end{thm}
\begin{proof}
	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
\end{proof}

This is a powerful result, since if $\Pr(s)$ has a support contained within the domain $[-\bar{s}, \bar{s}]$, then to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $0, \Delta t, \dots, N_t \Delta t$, we only need to compute $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ for $\ell = -N_t,\dots,N_t$ once.
This is because, once $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ is computed for $\ell = -N_t, \dots ,N_t$ we can use the partition of $[-\bar{s}, \bar{s}]$ given by $\{ \bar{s} \frac{m}{\ell} \}_{m=-\ell}^{\ell}$ to compute a $\Delta s = \bar{s} / \ell$ accurate approximation for each $\ell$, courtesy of Theorem \ref{thm:main}.
%Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
The procedure to compute   $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.

\begin{algorithm}[t]
		  \caption{Algorithm to Compute  $\rho_t$.}
		  \label{alg:1}
		  \begin{algorithmic}[1]
		    \Require $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.
		    \ForEach {$m \in \{ -\ell,\dots,\ell \}$}
		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
		    \EndFor  
		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} \label{alg1:step2}.
		    \State Normalize with respect to $\check{x}_t$ to obtain an approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
		    \State Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$
		  \end{algorithmic}
\end{algorithm}

%\begin{algorithm}[t]
%		    \KwData{ $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.}
%		    \For{$\ell \in \{ -N_t, \dots, N_t\}$}{
%		    	Compute $\Phi_{k,1}^{\bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0,N)$ \;
%			\For{$m \in \{ -\ell,\dots,\ell \}$}{
%				Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ \;
%			}
%			Use Theorem \ref{thm:main} to obtain a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$\;
%			Normalize with respect to $\check{x}_t$ to obtain an equally accurate approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$ \;
%			Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$.
%		    }
%		    \KwResult{An approximation of $\rho_t(x_t)$ at $t= 0,\Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
%		    \caption{Our algorithm for computing $\rho_t$}
%		    		    \label{alg:main}
%\end{algorithm}

%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at $t = -N_t\bar{s}, (1-N_t)\bar{s} ,\dots, N_t \bar{s}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
%We summarize this in the following theorem:
%\begin{thm}
%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
%\end{thm}

\section{Implementation and Experimental results} \label{sec:implementation}
  Given this model, we describe a specific implementation and the process of fitting the model to a dataset.
  In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
  For the purposes of demonstration, we use the Stanford Drone Dataset \cite{Robicquet2016}.
  More generally, we assume that for a fixed scene we have a database of previously observed trajectories $\{ \hat{x}^1, \dots, \hat{x}^N\}$.
  From this data we tune the parameters of the model ($\{X_1, \dots, X_n\}$, $\Pr( x_0 \mid M)$ and $\Pr(M)$)  appropriately.

  \subsection{Learning the Vector Fields}
  Before we begin to learn vector-fields, we must learn the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
  In particular, we use the start and end point for each trajectory to obtain a point in $\mathbb{R}^4$.
  We then cluster in $\mathbb{R}^4$ using Affinity propagation \cite{FreyDueck2007}.
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

  For each set $S_k$ we learn a vector-field that is approximately compatible with that set.
  %We'd like to only consider vector-fields who's vectors have unit magnitude because we observe many of the trajectories in the data appear to have a roughly constant speed.
  Since most trajectories appearing in the dataset have roughly constant speed, we chose a vector-field that has unit magnitude. 
  That is, we assume the vector-field takes the form $X_k(x) = \left( \cos( \Theta_k(x) ) , \sin(\Theta_k(x)) \right)$ for some scalar function $\Theta_k(x)$.
  Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
  We assume $\Theta_k$ takes the form
  \begin{align*}
  	\Theta_k(x) = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x),
  \end{align*}
  for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
  In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
 $\Theta_k$ can be learned by computing the velocities observed in the cluster, $S_k$.
  These velocities may be obtained by a low order finite difference formula.
  Upon normalizing the velocities, we obtain a unit-length velocity vectors, $v_{i,k}$, anchored at each point, $x_{i,k}$, of $S_k$.
  We learn $\Theta_k$ by defining the cost-function:
  \begin{align*}
  	C[ \Theta_k] = \sum_i \langle v_{i,k} , ( \cos(\Theta_k( x_{i,k}) , \sin( \Theta_k( x_{i,k} ) ) \rangle
  \end{align*}
  which penalizes $\Theta_k$ for producing a misalignment with the observed velocities at the observed points of $S_k$.
  When $\Theta_{k}$ includes high order polynomials (e.g. beyond 5th order), we also include a regularization term to bias the cost towards smoother outputs.
  Using the $H^1$-norm times a fixed scalar suffices as a regularization term.

  \subsection{Learning $\Pr( x_0 \mid M)$ and $\Pr(M)$}

  We begin by considering the nonlinear models first.
%  That is to say, we are concerned with $M = (k,s)$ for an integer $k$ and scalar $s$.
%  We'd like to compute $\Pr( x_0 \mid k,s)$ for each cluster $k=1,\dots,n$ and each speed $s \in \mathbb{R}$.
  We make the modeling assumption that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
  Additionally, we assume that $s$ and $k$ are independent.
  This means that we only need to learn $\Pr( x_0 \mid k)$, $\Pr(k)$, and $\Pr(s)$.

  We let $\Pr(k) = (n+1)^{-1}$ and $\Pr(s) \sim \mathcal{U}( [-s_{\max}, s_{\max} ] )$  where $s_{\max}>0$ is the largest observed speed in the dataset.  
  This implies that $\Pr(lin) = (n+1)^{-1}$ as well.
  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.

%  To learn $\Pr( x_0 \mid k)$, we assume that its logarithm may be approximated by a low dimensional subspace of functions on $\mathbb{R}^2$.
%  That is to say, 
  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is given by:
  \begin{align*}
  	V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)
  \end{align*}
  for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
  We chose our basis functions to be the collection of tensor products from the first six Legendre polynomials, normalized to the size of the domain.
  Then, one may fit the coefficients $c_{\alpha}$ to the data by using a log-likelihood criterion.
  The resulting (convex) optimization problem takes the form:
  \begin{align*}
  	\mathbf{c}^* = \inf_{ |\mathbf{c}| } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
  \end{align*}
  Where the norm on $\mathbf{c}$ is a sup-norm.
  We can also bias this optimization towards more regular functions by adding a penalty to the cost function.
    Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.

  \subsection{Learning the Measurement Model}
  We assume a Gaussian noise model.  
  That is
  \begin{align*}
  	\Pr( \hat{x}_0 \mid x_0 ) \sim \mathcal{N}( x_0 , \sigma_x) \quad, \quad \Pr( \hat{v}_0 \mid v_0 ) \sim \mathcal{N}( v_0, \sigma_v).
  \end{align*}
  Therefore, our model is parametrized by the standard deviations $\sigma_x$ and $\sigma_v$.
  We assume that the true trajectory of an agent is smooth compared to the noisy output of our measurement device.
  This justifies smoothing the trajectories, and using the difference between the smoothed signals and the raw data to learn the variance $\sigma_x$.
  To obtain the results in this paper we have used a moving average of four time steps (this is $0.13$ seconds in realtime).
  We set $\sigma_v = 2 \sigma_x / \Delta t$ where $\Delta t > 0$ is the time-step size.  
  This choice is justified from the our use of finite differences to estimate velocity.
  In particular, if velocity is approximated via finite differencing as $v(t) \approx (x(t+h) - x(t))\,\Delta t^{-1} + \mathcal{O}(h)$ and the measurements are corrupted by Gaussian noise, then the measurement $\hat{v}(t)$ is related to $v(t)$ by Gaussian noise with roughly the same standard deviation as $(x(t+h) - x(t))\,\Delta t^{-1}$.

  \subsection{Learning the Noise Model}
  Finally, we assume that the true position is related to the model by Gaussian noise with a growing variance.
  In particular, we assume $\Pr( x_t \mid \check{x}_t) \sim \mathcal{N}( \check{x}_t , \kappa t)$ for some constant $\kappa \geq 0$.
  The parameter, $\kappa$, must be learned.
 % We've taken an adhoc approach in this paper, although more intelligent and theoretically sound alternatives likely exist.
  For each curve in $S_k$ we create a synthetic curve using the initial position and speed and integrating the corresponding vector-field, $s\, X_k$.
  So for each curve, $x_i(t)$, of $S_k$, we have a synthesized curve $x_{i,synth}(t)$ as well.
  We then measure the standard deviation of $(x_i(t) - x_{i,synth}(t)) / t$ over $i$ and at few time, $t \in \{ 0, 100, 200 \}$ in order to obtain $\kappa$.

 \subsection{Evaluating Performance}
In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
We implemented our model as well as the evaluation code in Python 2.6. Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Our Algorithm}}   
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Kitani et al.}} 
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Random Walk} }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.2cm}
	\hspace{.5cm}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 2.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 8.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 12.33s$
	\end{minipage}

	\caption{We behave qualitatively and quantitatively better than kitani in this example. Our algorithm starts out with a dense distribution, which spreads with time, whereas Kitani et al.'s algorithm is very confident in its mistake.}

	\label{fig:death-1-2}
\end{figure}

\begin{figure}
	\centering
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Our Algorithm}   }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Kitani et al.}  }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Random Walk} }  
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.png}
	\end{minipage}

	\vspace{0.2cm}
	\hspace{.5cm}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 1.66s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 4.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 12.33s$
	\end{minipage}

	\caption{We predict this trajectory very well, and while Kitani et al. captures the same behaviour, however their motion model doesn't adequately keep up with the ground truth of the pedestrian.}

	\label{fig:bookstore-1-2}
\end{figure}

Our data came from the Stanford Drone Dataset in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles). 
Our test data used the bicycle annotations. 
We did a 2-fold cross validation on 4 different scenes of varying complexity. 
Our analysis used the same partitions of the trajectories for all three predictors.

Our data came from the Stanford Drone Dataset, in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles). Our datasets used the bicycle data. We did a 2-fold cross validation on 4 different scenes of varying complexity. Our analysis used the same partitions of the trajectories for all three predictors.

The algorithm provided by Kitani et al. was given additional information when making predictions that neither ours nor the random walk model were provided, namely the endpoint of the test trajectory. Without these data the implementation of the predictor provided by the authors devolves into a random walk.

The output distributions of the three algorithms were compared using their integrals over the cells of a suitably fine regular grid over our domain. These integrals are used to visualize the distributions in Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2}.
In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. The dimensions of the bounding box were the average bounding box dimensions in the training data set. Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves. Figure \ref{fig:auc_vs_time} shows the Area Under the Curve of these ROC curves versus time.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/the_results.eps}
	\caption{We see that at all times we outperform Random Walk and Kitani et al.'s algorithm. The initial dip in Kitani et al.'s algorithm is due to their initial distribution being very sharp.}
	\label{fig:auc_vs_time}
\end{figure}

Disregarding small time scales where any algorithm will be reasonably accurate, our predictor behaves much better than any of the others at moderately large $t$, where we capture the  behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. Their encoding of velocity results in the prediction lagging behind the actual observed motion. Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales. 
\begin{table}
	\begin{center}
		\caption{Runtimes}
		\label{tab:time}
		\renewcommand{\arraystretch}{1.5}% Spread rows out...
		\begin{tabular}{||c | c c  c ||} 
			\hline
			& Our Predictor & Random Walk & Kitani et al. \\ [0.5ex] 
			\hline 
			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.0169s & 2.1E-7s & 0.0614s \\
			\hline

		\end{tabular}

	\end{center}
\end{table}

The run time per frame for each algorithm was generated using total run time for 400 frames, averaged across several agents and scenes, divided by the number of frames. This is shown in Table~\ref{tab:time}. Our algorithm implementation leveraged its embarassing parallelism using a relatively naive parallelization scheme, which split the computation of frames among 18 cores. It's highly probable that our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. Kitani et al. was timed with minimal modification to the source code provided by the authors.

Figure \ref{fig:auc_vs_time} shows that at all time scales our algorithm has better predictive capability than Kitani et. al's. We also run over 3 times faster than their algorithm, at over 60FPS.  Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2} show the qualitative superiority of our predictions on several scenes.

\section{Conclusion} 
\label{sec:conclusion}

\Ram{Matt can you please take a crack at this ....}
Avenues of improvement:
update predictions when given additional measurements

The conclusion goes here.

\section*{Acknowledgments}
Kris Kitani was critical in aiding the comparison to his algorithm.

%\appendix
%\subsection{Inverse Optimal Control} \label{app:ioc}
%One of the valuable aspect of the motion model of \cite{Kitani2012} was that the trajectories were solutions of optimal control problems.
%This is a desirable property for a motion model of pedestrians/cyclists/vehicles in that it seems reasonable to assume that we all move through the world with an intent to get somewhere.
%This framework was dubbed \emph{inverse optimal control} (IOC).
%Here we illustrate how solutions of \eqref{eq:ode} can fit within the IOC framework.
%
%\begin{thm}
%	Let $X$ be a vector-field such that $\| X(x) \|_2 = 1$ for all $x \in \mathbb{R}^2$.
%Then, for each initial condition $x_0 \in \mathbb{R}^2$, there exists a neighborhood $U$, a Riemannian metric, $g$, and a cost function, $f$, 
%such that solutions of \eqref{eq:ode} are also solutions of the optimal control problem
%\begin{align}
%	p^* = \inf_{ \| v_t \|_g = s } \int_0^T f( x_t) dt \label{eq:IOC}
%\end{align}
%where $v_t = \frac{d x_t}{dt}$, $\| \cdot \|_g$ is the norm with respect to $g$ and the infimum is taken over all differentiable curves in $U$ which originate from $x_0$.
%\end{thm}
%
%\begin{proof}
%	By rescaling time, we may let $s$ be any positive real number.
%	Therefore, without loss of generality, we will seek an $f$ and $g$ such that $s=1$.
%	In this case a solution of \eqref{eq:IOC} is generated by the vector-field $X = \frac{\nabla f}{ \| \nabla f\|}$ where $\nabla$ is the Riemannian gradient with respect to some metric (possibly non-Euclidean).
%	Therefore, our goal is to illustrate that there exists a metric $g$ and a function $f$ such that the given $X$ is given by $\frac{\nabla f}{ \| \nabla f\|}$.
%
%	Because $\| X \| = 1$ globally, it has no fixed points.
%	From the flow-box theorem \cite[Theorem 4.1.14]{MTA} we can assert that for any open set, $U \subset \mathbb{R}^2$ there exists a local diffeomorphism which transforms $X_k$ into a flat vector-field.
%	Mathematically, this asserts the existence of a map $\Phi : U \to V \subset \mathbb{R}^2$ such that the push-forward of $X$, which we will denote by $\tilde{X}$, and given by
%	\begin{align*}
%		\tilde{X}(\tilde{x}) = \Phi_* X(\tilde{x}) := \left. D\Phi \right|_{\Phi^{-1}(\tilde{x}) } \cdot X(  \Phi^{-1}(\tilde{x}) ),
%	\end{align*}
%	is just the flat vector field $(1,0)$ for all $x \in U$.
%	We could view the coordinate function $\tilde{f}(x) = x^0$ as a cost function on $V$,
%	and we may set $g_V$ to be equal to the standard flat metric on $V$ inherited from $\mathbb{R}^2$.
%	In this case we observe that $\tilde{X}$ generates solutions to the optimization problem
%	\begin{align*}
%		\tilde{p}^* = \inf_{ \| \tilde{v}_t \|_{g_V} \leq 1} = \int_0^T \tilde{f}( \tilde{x}_t ) dt
%	\end{align*}
%	By a change of variables, it follows that the original vector field, $X$, then solves the optimization problem \eqref{eq:IOC}
%	where $g = \Phi^* g_V$ is the pull-back metric, and $f = \Phi^* \tilde{f}$ is the pull-back of $\tilde{f}$.
%\end{proof}

%% Use plainnat to work nicely with natbib.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2017-02-15 pedestrian_forecasting_iros_2017.tex
still working on the edits

@@
 More specifically, we assume that each $X_k$ has the property that $\| X_k(x) \| = 1$ for all $x \in D$.
-This property ensures that speed is constant in time
+This property ensures that speed is constant in time.
 % and it has the further advantage of being the (local) generator of solutions to an optimal navigation problem (see Appendix \ref{app:ioc}).
@@
 For each initial condition, $x_0$, we can solve \eqref{eq:ode} as an initial value problem, to obtain a point $\check{x}_t$ with initial condition $\check{x}_0 = x_0$.
-This process, which takes $\check{x}_0$ and outputs $\check{x}_t$, constitutes a map which is termed the flow-map \cite[Ch 4]{MTA}, and which we denote by $\Phi_{k,s}^t$.
+This process of solving the differential equation takes an initial condition, $\check{x}_0$, and outputs a final condition, $\check{x}_t$.
+This constitutes a map which is termed the \emph{flow-map} \cite[Ch 4]{MTA}, and which we denote by $\Phi_{k,s}^t$.
 Explicitly, we have the posterior:
@@
 \end{align}
-We may use this Bayesian network to compute $\rho_t$ efficiently.
+We use this Bayesian network to compute $\rho_t$ efficiently.
 In particular
@@
 	\rho_t(x_t ) &:= \Pr( x_t \mid \hat{x}_0, \hat{v}_0 ) \\
-	&= \left( \sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds \right) \\
+	&= \left( \sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds \right)  \nonumber \\
 	&\quad + \Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 ). \label{eq:decomposition}
@@
 where the error is bounded by $\int| \varepsilon_s | ds \leq  TV(x_t, \hat{x}_0, \hat{v}_0) \Delta s$
-where $TV(x_t, \hat{x}_0, \hat{v}_0)$ is the sum, with respect to $k$, of the total variation of $\Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 )$ with respect to $s$ for fixed $x_t, \hat{x}_0, \hat{v}_0$.
-Since this error term can be controlled, the problem of solving $\rho_t(x_t)$ is reduced to that of efficiently computing $ \Pr( x_t, k , s_j  \mid \hat{x}_0, \hat{v}_0)$
+where $TV(x_t, \hat{x}_0, \hat{v}_0)$ is the sum, with respect to $k$, of the total variation of $\Pr( x_t, k , s, \hat{x}_0, \hat{v}_0 )$ with respect to $s$ for fixed $x_t, \hat{x}_0, \hat{v}_0$.
+Since this error term can be controlled, the problem of solving $\rho_t(x_t)$ is reduced to that of efficiently computing $ \Pr( x_t, k , s_j, \hat{x}_0, \hat{v}_0)$
 for a fixed collection of $s_j$'s.
@@
 Assuming, for the moment, that such a convolution can be performed efficiently, we focus on computation of $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$.
-
 Again, \eqref{eq:pgm} implies:
@@
  where $\Psi( \hat{v}_0 ;k,s,x_0) := \left. \Pr( \hat{v}_0 \mid v_0) \right|_{v_0 = s X_k(x_0)}$.
- Assuming the density $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0)$ is of bounded variation in the variable $x_0$ (with all other variables held fixed),
- we may approximate it as a sum of weighted Dirac-delta distributions supported on a regular grid.
+ We may approximate $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0)$ as a sum of weighted Dirac-delta distributions supported on a regular grid, since $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0)$  is of bounded variation in the variable $x_0$ (with all other variables held fixed).
  %This is merely the definition of \emph{Riemann integrable} \cite{Rudin1991}.
@@
  Choose  $L>0$ to be such that $\int_{S_L(\hat{x}_0)} \Pr( x_0 \mid \hat{x}_0) dx_0 = 1 - \varepsilon_{tol}$ for some error tolerance $\varepsilon_{tol}>0$.
- Then, for a given resolution $N_x \in \mathbb{N}$ define the regular grid on $S_L(\hat{x}_0)$ as $\Gamma_L( \hat{x}_0 ; N_x ) := \left\{ x_0^{i,j} \mid i,j = -N_x,\dots,N_x \right\}$, where $x_0^{i,j} = \hat{x}_0 + \frac{L}{2N_x}(i,j)$.
+ Then, for a given resolution $N_x \in \mathbb{N}$ define the regular grid on $S_L(\hat{x}_0)$ as $\Gamma_L( \hat{x}_0 ; N_x ) := \left\{ x_0^{i,j} \mid i,j \in \{ -N_x,\dots,N_x \} \right\}$, where $x_0^{i,j} = \hat{x}_0 + \frac{L}{2N_x}(i,j)$.
  The grid spacing is given by $\Delta x = ( \frac{L}{2N_x}, \frac{L}{2N_x} )$.
-Approximate the smooth distribution $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ as a weighted sum of Dirac-deltas (in the variable $x_0$) supported on $\Gamma_L( \hat{x}_0;N)$:
+We approximate the smooth distribution $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ as a weighted sum of Dirac-deltas (in the variable $x_0$) supported on $\Gamma_L( \hat{x}_0;N)$:
 \begin{align}
@@
 \end{align}
-where $W(k,s,i,j,\hat{x}_0)$ is merely the evaluation of $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ at the grid point $x_0 = x_0^{i,j} \in \Gamma( \hat{x}_0; N)$.
+where $W(k,s,i,j,\hat{x}_0)$ is the evaluation of $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ at the grid point $x_0 = x_0^{i,j} \in \Gamma( \hat{x}_0; N)$.
+More explicitly, this evaluation can be done for each grid point by using only the assumed posterior models in \eqref{eq:pgm}.
 %That is to say
@@
 %The function $W(k,s,i,j, \hat{x}_0)$, can be computed efficiently as product of the posteriors represented in our Bayesian network \eqref{eq:pgm}.
+
 Substitution of \eqref{eq:approximation 1} into the final line of \eqref{eq:push forward} yields:
@@
 	\end{split}
+	\label{eq:makesense}
 \end{align}
 where $\varepsilon_t( \check{x}_t) = \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right)  \varepsilon_0(x_0) dx_0$.
-The first term is computable by flowing the points of the grid, $\Gamma_L(\hat{x}_0; N_x)$, along the vector field $s X_k$.
+The first term of the right hand side of \eqref{eq:makesense} is computable by flowing the points of the grid, $\Gamma_L(\hat{x}_0; N_x)$, along the vector field $s X_k$.
 The second term, $\varepsilon_t$, may be viewed as an error term.
@@
 	The error term, $\varepsilon_t \sim \mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ in the $L^1$-norm, for fixed $k,s,\hat{x}_0$, and $\hat{v}_0$.
-	Moreover, $\| \varepsilon_t \|$ is constant in time.
+	Moreover, $\| \varepsilon_t \|_{L^1}$ is constant in time.
 \end{thm}
@@
 \end{proof}
-This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$.
-This yields the following theorem
+This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$, which yields the following result:
 \begin{thm} \label{thm:main}
@@
 
-This is a powerful result, since if $\Pr(s)$ has a support contained within the domain $[-\bar{s}, \bar{s}]$, then to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $0, \Delta t, \dots, N_t \Delta t$, we only need to compute $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ for $\ell = -N_t,\dots,N_t$ once.
-This is because, once $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ is computed for $\ell = -N_t, \dots ,N_t$ we can use the partition of $[-\bar{s}, \bar{s}]$ given by $\{ \bar{s} \frac{m}{\ell} \}_{m=-\ell}^{\ell}$ to compute a $\Delta s = \bar{s} / \ell$ accurate approximation for each $\ell$, courtesy of Theorem \ref{thm:main}.
+This is a powerful result, since if $\Pr(s)$ has a support contained within the domain $[-\bar{s}, \bar{s}]$, then to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $0, \Delta t, \dots, N_t \Delta t$, we only need to compute $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ for each $\ell \in  \{-N_t,\dots,N_t\}$ once.
+This is because, once $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ is computed for each $\ell \in \{-N_t, \dots ,N_t\}$ we can use the partition of $[-\bar{s}, \bar{s}]$ given by $\{ \bar{s} \frac{m}{\ell} \}_{m=-\ell}^{\ell}$ to compute a $\Delta s = \bar{s} / \ell$ accurate approximation for each $\ell$, courtesy of Theorem \ref{thm:main}.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
@@
 %Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
-For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at $t = -N_t\bar{s}, (1-N_t)\bar{s} ,\dots, N_t \bar{s}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
-Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
+For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
+Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
 If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.

### 2017-02-15 pedestrian_forecasting_iros_2017.tex
Resize images

--- 2017-02-15 pedestrian_forecasting_iros_2017.tex
@@
 	\begin{subfigure}[b]{.45\linewidth}
-		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=55.png}
+		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=55.jpg}
 		\caption{$t=1.83s$}
@@
 		\begin{subfigure}[b]{.45\linewidth}
-			\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=145.png}
+			\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=145.jpg}
 			\caption{$t=4.83$}
@@
 	\begin{subfigure}[b]{.45\linewidth}
-		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=235.png}
+		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=235.jpg}
 		\caption{$t=7.83s$}
@@
 	\begin{subfigure}[b]{.45\linewidth}
-		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=345.png}
+		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=345.jpg}
 		\caption{$t=11.5s$}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.jpg}
 	\end{minipage}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.jpg}
 	\end{minipage}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.jpg}
 	\end{minipage}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.jpg}
 	\end{minipage}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.jpg}
 	\end{minipage}
@@
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.jpg}
 	\end{minipage}
 	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.png}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.jpg}
 	\end{minipage}

### 2017-02-15 pedestrian_forecasting_iros_2017.tex
minor changes to implementation portion of text.

--- 2017-02-15 pedestrian_forecasting_iros_2017.tex
@@
   \subsection{Learning the Vector Fields}
-  Before we begin to learn vector-fields, we must learn the number of possible vector-fields.
+ We begin by identifying the number of possible vector-fields.
   To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
@@
   Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
-  We assume $\Theta_k$ takes the form
+  We assume $\Theta_k$ takes the form:
   \begin{align*}
@@
   In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
- $\Theta_k$ can be learned by computing the velocities observed in the cluster, $S_k$.
-  These velocities may be obtained by a low order finite difference formula.
+ $\Theta_k$ is learned by computing the velocities observed in the cluster, $S_k$.
+  These velocities are obtained by a low order finite difference formula.
   Upon normalizing the velocities, we obtain a unit-length velocity vectors, $v_{i,k}$, anchored at each point, $x_{i,k}$, of $S_k$.
@@
 
-  We begin by considering the nonlinear models first.
+We first considering the nonlinear models.
 %  That is to say, we are concerned with $M = (k,s)$ for an integer $k$ and scalar $s$.
 %  We'd like to compute $\Pr( x_0 \mid k,s)$ for each cluster $k=1,\dots,n$ and each speed $s \in \mathbb{R}$.
-  We make the modeling assumption that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
+  We begin by assuming that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
   Additionally, we assume that $s$ and $k$ are independent.
@@
   This implies that $\Pr(lin) = (n+1)^{-1}$ as well.
-  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.
+%  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.
 
@@
   for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
-  We chose our basis functions to be the collection of tensor products from the first six Legendre polynomials, normalized to the size of the domain.
+  We chose our basis functions to be the collection of tensor products of the first six Legendre polynomials, normalized to the size of the domain.
   Then, one may fit the coefficients $c_{\alpha}$ to the data by using a log-likelihood criterion.
@@
   Where the norm on $\mathbf{c}$ is a sup-norm.
-  We can also bias this optimization towards more regular functions by adding a penalty to the cost function.
+  We bias this optimization towards more regular functions by adding a penalty to the cost function.
     Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.
@@
 In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
-We implemented our model as well as the evaluation code in Python 2.6. Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
+We implement our model as well as the evaluation code in Python 2.6. 
+Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
ran through the implementation.

@@
  \subsection{Evaluating Performance}
-In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
-We implement our model as well as the evaluation code in Python 2.6. 
-Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
+
+This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
+We implement our model as well as our evaluation code in Python 2.6. 
+Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
+Our test data used the bicycle annotations in \cite{Robicquet2016}. 
+We did a 2-fold cross validation on 4 different scenes of varying complexity. 
+Our analysis used the same partitions of the trajectories for all three predictors.
+
+\begin{figure}
+	\centering
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Our Algorithm}   }
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Kitani et al.}  }
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Random Walk} }  
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.2cm}
+	\hspace{.5cm}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 1.66s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 4.33s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 12.33s$
+	\end{minipage}
+
+	\caption{An illustration of the predictions generated by the various algorithms.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.}
+
+	\label{fig:bookstore-1-2}
+\end{figure}
 
@@
 
-	\caption{We behave qualitatively and quantitatively better than kitani in this example.
-Our algorithm starts out with a dense distribution, which spreads with time, whereas Kitani et al.'s algorithm is very confident in its mistake.}
+	\caption{An illustration of the predictions generated by the various algorithms.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.}
 
@@
 
-\begin{figure}
-	\centering
-	\begin{minipage}[c]{0.5cm}
-		\rotatebox{90}{\small{Our Algorithm}   }
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.jpg}
-	\end{minipage}
+%Our data came from the Stanford Drone Dataset in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles).
 
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
-		\rotatebox{90}{\small{Kitani et al.}  }
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.jpg}
-	\end{minipage}
+Note that the implementation of the algorithm in \cite{Kitani2012} required the endpoint of each test trajectory. 
+Without this information the implementation of the predictor provided by the authors devolved into a random walk.
+Neither our algorithm nor the random walk were provided this information.
 
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
-		\rotatebox{90}{\small{Random Walk} }  
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.jpg}
-	\end{minipage}
-
-	\vspace{0.2cm}
-	\hspace{.5cm}
-	\begin{minipage}{0.3\linewidth}
-		\centering
-		$t = 1.66s$
-	\end{minipage}
-	\begin{minipage}{0.3\linewidth}
-		\centering
-		$t = 4.33s$
-	\end{minipage}
-	\begin{minipage}{0.3\linewidth}
-		\centering
-		$t = 12.33s$
-	\end{minipage}
-
-	\caption{We predict this trajectory very well, and while Kitani et al. captures the same behaviour, however their motion model doesn't adequately keep up with the ground truth of the pedestrian.}
-
-	\label{fig:bookstore-1-2}
-\end{figure}
-
-Our data came from the Stanford Drone Dataset in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles). 
-Our test data used the bicycle annotations. 
-We did a 2-fold cross validation on 4 different scenes of varying complexity. 
-Our analysis used the same partitions of the trajectories for all three predictors.
-
-Our data came from the Stanford Drone Dataset, in the form of hand annotated bounding boxes around agents (e.g.
-cars, pedestrians, bicycles).
-Our datasets used the bicycle data.
-We did a 2-fold cross validation on 4 different scenes of varying complexity.
-Our analysis used the same partitions of the trajectories for all three predictors.
-
-The algorithm provided by Kitani et al.
-was given additional information when making predictions that neither ours nor the random walk model were provided, namely the endpoint of the test trajectory.
-Without these data the implementation of the predictor provided by the authors devolves into a random walk.
-
-The output distributions of the three algorithms were compared using their integrals over the cells of a suitably fine regular grid over our domain.
+The output distributions of the three algorithms were compared using their integrals over the cells of a regular grid over our domain. 
 These integrals are used to visualize the distributions in Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2}.
-In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$.
-The dimensions of the bounding box were the average bounding box dimensions in the training data set.
+%In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 
+%The ground truth bounding box at each time instance were set equal to the average bounding box dimension centered at the 
+%were the average bounding box dimensions in the training data set. 
 Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves. 
 Figure \ref{fig:auc_vs_time} shows the Area Under the Curve of these ROC curves versus time.
+Our predictor behaves better than any of the others at moderately large $t$
+%, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
+%Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
+This is despite the algorithm in \cite{Kitani2012} knowing the specific final location of each agent.
+%Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.
+
 \begin{figure}
@@
 	\includegraphics[width=\linewidth]{./figures/the_results.eps}
-	\caption{We see that at all times we outperform Random Walk and Kitani et al.'s algorithm. The initial dip in Kitani et al.'s algorithm is due to their initial distribution being very sharp.}
+	\caption{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.}
 	\label{fig:auc_vs_time}
@@
 
-Disregarding small time scales where any algorithm will be reasonably accurate, our predictor behaves much better than any of the others at moderately large $t$, where we capture the  behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions.
-Their encoding of velocity results in the prediction lagging behind the actual observed motion.
-Even given significantly better information, Kitani et al.
-generates lower quality predictions than our algorithm at all time scales.
 \begin{table}
 	\begin{center}
-		\caption{Runtimes}
+		\caption{Comparison of runtimes of the various algorithms.}
 		\label{tab:time}
@@
 			\hline
-
 		\end{tabular}
-
 	\end{center}
@@
 This is shown in Table~\ref{tab:time}. 
-Our algorithm implementation leveraged its embarassing parallelism using a relatively naive parallelization scheme, which split the computation of frames among 18 cores.
-It's highly probable that our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme.
-Kitani et al.
-was timed with minimal modification to the source code provided by the authors.
+Our algorithm implementation leveraged its embarrassing parallelism using a relatively naive Python implementation, which split the computation of frames among 18 cores. 
+%Our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. 
+The algorithm in \cite{Kitani2012} was timed with minimal modification using the source code provided by the authors.
 
-Figure \ref{fig:auc_vs_time} shows that at all time scales our algorithm has better predictive capability than Kitani et.
-al's.
-We also run over 3 times faster than their algorithm, at over 60FPS.
-Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2} show the qualitative superiority of our predictions on several scenes.
+%Figure \ref{fig:auc_vs_time} shows that at all time scales our algorithm has better predictive capability than Kitani et. al's. 
+%We also run over 3 times faster than their algorithm, at over 60FPS.  
+%Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2} show the qualitative superiority of our predictions on several scenes.
 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
one sentence added to make it sound stronger

@@
 \end{itemize}
+The method proposed by this paper is able to work three times faster than the existing state of the art while improving upon its performance over long time horizons.
 
@@
 Figure \ref{fig:auc_vs_time} shows the Area Under the Curve of these ROC curves versus time.
-Our predictor behaves better than any of the others at moderately large $t$
+Our predictor behaves better than any of the other compared methods at moderately large $t$.
 %, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
 %Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
-This is despite the algorithm in \cite{Kitani2012} knowing the specific final location of each agent.
+This is despite providing the algorithm in \cite{Kitani2012} with the specific final location of each agent.
 %Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
removed some extra text

@@
 \end{figure}
-
-Autonomous systems are increasingly being deployed in and around humans.
- The ability to accurately model and anticipate human behavior is of great value to maximize safety, confidence, and effectiveness in human-centric environments. 
- The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of possible action over a set a of possible behaviors. 
- Since the set of plausible behaviors in arbitrary situations may be vast, this work constrains the problem to anticipating the possible future locations of people within a bounded area.
-This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).
-
- With a particular focus on the AV application several additional design criteria become important. 
- First, false negative rates for unoccupied regions must be minimized. 
- The misclassification of space in this way has obvious safety issues. 
- Second, speed is paramount. 
- To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories. 
- The margins between humans and vehicles are small and both frequently operate at the boundary. 
- Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.
-
-This paper presents an algorithm for real-time, long-term prediction of pedestrian behavior which can subsequently be used by autonomous agents.
-As depicted in Figure \ref{fig:}, this approach relies upon ...
 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
added a few more minor edits to the captions

@@
 	\caption{The performance of the presented algorithm captures the most probable routes that a pedestrian chooses.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the \textit{virdis} color palette.
 In this example which was captured at thirty frames per second, the presented algorithm took 0.0158s per frame in a Python implementation.
-In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.}
+}
 	\label{fig:gates-1-2}
@@
 	\caption{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the \textit{virdis} color palette.
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.}
@@
 	\caption{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the \textit{virdis} color palette.
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.}

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
corrected algorithm
conclusion

--- 2017-02-16 pedestrian_forecasting_iros_2017.tex
@@
 \Ram{Matt can you please take a crack at this ....}
-Avenues of improvement:
-update predictions when given additional measurements
-
-The conclusion goes here.
+This paper presents a real-time approach to probabilistic pedestrian modeling.
+We demonstrate the ability to accurately predict the final location of a pedestrian with comparable accuracy to a state-of-the-art with large gains in three critical areas: 1) prediction speed is highly efficient enabling tight loop real-time control;  2) our prediction uncertainty bound much more accurately reflect potential paths for the human limiting false negatives and maximize the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a much wider variety of real world applications.
+These three features have great promise for robotic applications and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces.
+For future directions we plan to optimize the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
reformulated the disccussion after theorem 3

--- 2017-02-16 pedestrian_forecasting_iros_2017.tex
@@
 We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
-Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on the support of $\Pr(s)$, 
+Let us assume that $\Pr(s \mid k)$ is compactly supported for all $k = 1,\dots,n$, and the supported is always contained
+in some interval $[ - \bar{s} , \bar{s} ]$ for some $\bar{s} > 0$.
+Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on $[-\bar{s}, \bar{s}]$, 
 we can conclude that the integral term in \eqref{eq:decomposition} can be approximated by
@@
 		\begin{split}
-		&\Delta s \sum_{i,j,k,\ell} W(k,s_\ell,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{s_\ell t}( x_0^{i,j}) } \\
+		&\Delta s \sum_{i,j,k,m} W(k,s_m,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{s_m t}( x_0^{i,j}) } \\
 		&+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 )
@@
 
-This is a powerful result, since if $\Pr(s)$ has a support contained within the domain $[-\bar{s}, \bar{s}]$, then to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $0, \Delta t, \dots, N_t \Delta t$, we only need to compute $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ for each $\ell \in  \{-N_t,\dots,N_t\}$ once.
-This is because, once $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ is computed for each $\ell \in \{-N_t, \dots ,N_t\}$ we can use the partition of $[-\bar{s}, \bar{s}]$ given by $\{ \bar{s} \frac{m}{\ell} \}_{m=-\ell}^{\ell}$ to compute a $\Delta s = \bar{s} / \ell$ accurate approximation for each $\ell$, courtesy of Theorem \ref{thm:main}.
+This is a powerful result, because it allows us to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $t=\Delta t, \dots, N_t \Delta t$ by a single computation of $\Phi_{k,1}^{\ell \Delta t \bar{s}}( x_0^{i,j})$ for each $\ell \in  \{-N_t,\dots,N_t\}$, $k \in \{1,\dots,n\}$, and $x_0^{i,j} \in \Gamma_L( \hat{x}_0 , N_x)$.
+To explain this, assume we are given $\Phi_{k,1}^{m \Delta t \bar{s}}(x_0^{i,j})$ for all $|m| \leq \ell$.
+We can then use the regular partition
+\begin{align}
+	\mathcal{P}_\ell := \left\{ \frac{m}{\ell} \bar{s} \right\}_{m=-\ell}^{\ell} \label{eq:partition}
+\end{align}
+of $[-\bar{s}, \bar{s}]$ to compute a Riemann sum approximation using Theorem \ref{thm:main}.
+The partition $\mathcal{P}_\ell$ has a width of size, $\Delta s = \bar{s} / \ell$, and substitution in \eqref{eq:approximation 3} yields an approximation of $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ given by
+\begin{align*}
+	\begin{split}
+		&\frac{\bar{s}}{\ell} \sum_{i,j,k,m} W(k,m \bar{s} / \ell ,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{m \bar{s} \Delta t}( x_0^{i,j}) } \\
+		&\qquad+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 ).
+	\end{split}
+\end{align*}
+As we've already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
@@
 \begin{algorithm}[t]
-		  \caption{Algorithm to Compute  $\rho_t$.}
+		  \caption{Algorithm to Compute  $\rho_t$ at times $t= \Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
 		  \label{alg:1}
 		  \begin{algorithmic}[1]
-		    \Require $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.
-		    \ForEach {$m \in \{ -\ell,\dots,\ell \}$}
-		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
+		    \Require $\bar{s} > 0, N_t \in \mathbb{N}, \Delta t > 0,   \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$,
+		    routines for $\Pr(x_0 \mid M), \Pr(x_0 \mid \hat{x}_0), \Pr(v_0 \mid \hat{v}_0 ), \Pr(M \in \mathcal{M}), \Pr(s \mid k)$ .
+		    \For {$\ell = 1,\dots,N_t$}
+		    \State Compute and store $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
+		    \For {$m = -\ell,\dots,\ell$}
+		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ using \eqref{eq:pgm}.
 		    \EndFor  
-		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} \label{alg1:step2}.
+		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
+		    with partition $\mathcal{P}_\ell$ of \eqref{eq:partition}.
+		   \label{alg1:step2}.
 		    \State Normalize with respect to $\check{x}_t$ to obtain an approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
 		    \State Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$
+		    \EndFor
 		  \end{algorithmic}

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
small updates

--- 2017-02-16 pedestrian_forecasting_iros_2017.tex
@@
 The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of an action over a set of possible behaviors. 
-Since the set of plausible human behaviors is vast, this work focuses on anticipating the possible future locations of people within a bounded area. 
+Since the set of plausible human behaviors is vast, this work focuses on anticipating the possible future locations of pedestrians within a bounded area. 
 This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).
@@
 The output distributions of the three algorithms were compared using their integrals over the cells of a regular grid over our domain. 
-These integrals are used to visualize the distributions in Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2}.
+These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
 %In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
small changes

@@
 \label{sec:conclusion}
-
-\Ram{Matt can you please take a crack at this ....}
 This paper presents a real-time approach to probabilistic pedestrian modeling.
-We demonstrate the ability to accurately predict the final location of a pedestrian with comparable accuracy to a state-of-the-art with large gains in three critical areas: 1) prediction speed is highly efficient enabling tight loop real-time control;  2) our prediction uncertainty bound much more accurately reflect potential paths for the human limiting false negatives and maximize the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a much wider variety of real world applications.
+We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with large additional contribtions in three critical areas: 1) prediction speed is highly efficient enabling tight loop real-time control;  2) our prediction uncertainty bound much more accurately reflect potential paths for the human limiting false negatives and maximize the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a much wider variety of real world applications.
 These three features have great promise for robotic applications and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces.

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
editing through implementation

@@
 \begin{abstract}
-That autonomous vehicles will come to dominate our streets is imminent.
-This motivates the need for a real-time probabilistic forecasting algorithm for pedestrians, cyclists, and other agents,
-as it forms a necessary step in assessing the risk (and therefore the cost) we should expect to incur.
-In this paper, we present a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations.
-The resulting algorithm is embarrassingly parallel, and trained on historical trajectory information within a fixed scene.
-When compared with MDP-based methods, our algorithm appears to be superior from the standpoint of precision and recall.
+The success of autonomous systems will depend upon their ability to safely navigate human-centric environments.
+This motivates the need for a real-time, probabilistic forecasting algorithm for pedestrians, cyclists, and other agents since these predictions will form a necessary step in assessing the risk of any action.
+This paper presents a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations that are learned from historical trajectory information within a fixed scene.
+The resulting algorithm is embarrassingly parallel and is able to work at real-time speeds using a naive Python implementation.
+The quality of predicted locations of agents generated by the proposed algorithm is validated on a variety of examples and considerably higher than existing state of the art approaches over long time horizons.
+%When compared with MDP-based methods, this algorithm 
 \end{abstract}
@@
 The input to the algorithm at runtime is a noisy measurement of position and velocity, $\hat{x}_0, \hat{v}_0 \in \mathbb{R}^2$.
-If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$.
+If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$ for each time $t \in \{\Delta t, \dots, N_t \Delta t \}$ for some user-specified $N_t \in \mathbb{N}$ and $\Delta t \in \mathbb{R}$.
 
@@
 \subsection{The Variables of the Model}
-The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,T]$ for some $T>0$.
+The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,N_t \Delta t]$.
 We denote the position and velocity at time $t=0$ by $x_0$ and $v_0$ respectively.
@@
 
-This is a powerful result, because it allows us to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $t=\Delta t, \dots, N_t \Delta t$ by a single computation of $\Phi_{k,1}^{\ell \Delta t \bar{s}}( x_0^{i,j})$ for each $\ell \in  \{-N_t,\dots,N_t\}$, $k \in \{1,\dots,n\}$, and $x_0^{i,j} \in \Gamma_L( \hat{x}_0 , N_x)$.
-To explain this, assume we are given $\Phi_{k,1}^{m \Delta t \bar{s}}(x_0^{i,j})$ for all $|m| \leq \ell$.
+This is a powerful result since it allows us to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $t \in \{\Delta t, \dots, N_t \Delta t\}$ with a single computation of $\Phi_{k,1}^{\ell \Delta t \bar{s}}( x_0^{i,j})$ for each $\ell \in  \{-N_t,\dots,N_t\}$, $k \in \{1,\dots,n\}$, and $x_0^{i,j} \in \Gamma_L( \hat{x}_0 , N_x)$.
+To appreciate this, assume we are given $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for all $m \in \{ -\ell, \ldots, \ell \}$.
 We can then use the regular partition
@@
 \end{align*}
-As we've already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
+As we have already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
@@
 \begin{algorithm}[t]
-		  \caption{Algorithm to Compute  $\rho_t$ at times $t= \Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
+		  \caption{Algorithm to compute  $\rho_t$ for each $t \in  \{\Delta t, 2 \Delta t, \dots, N_t \Delta t$\}.}
 		  \label{alg:1}
 		  \begin{algorithmic}[1]
-		    \Require $\bar{s} > 0, N_t \in \mathbb{N}, \Delta t > 0,   \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$,
-		    routines for $\Pr(x_0 \mid M), \Pr(x_0 \mid \hat{x}_0), \Pr(v_0 \mid \hat{v}_0 ), \Pr(M \in \mathcal{M}), \Pr(s \mid k)$ .
-		    \For {$\ell = 1,\dots,N_t$}
-		    \State Compute and store $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
-		    \For {$m = -\ell,\dots,\ell$}
-		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ using \eqref{eq:pgm}.
+		    \Require $\bar{s} > 0$, $N_t \in \mathbb{N}$, $\Delta t > 0$,  $ \{ X_k \}_{k=1}^{n}$, $\hat{x}_0$, $\hat{v}_0$, $\Gamma_L( \hat{x}_0, N_x)$, $\Pr(x_0 \mid M), \Pr(x_0 \mid \hat{x}_0), \Pr(v_0 \mid \hat{v}_0 ), \Pr(M \in \mathcal{M}), \text{ and } \Pr(s \mid k)$.
+		    \For {$\ell \in \{1,\dots,N_t\}$}
+		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
+		    \For {$m \in \{-\ell,\dots,\ell\}$}
+		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
 		    \EndFor  
-		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
-		    with partition $\mathcal{P}_\ell$ of \eqref{eq:partition}.
-		   \label{alg1:step2}.
-		    \State Normalize with respect to $\check{x}_t$ to obtain an approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
-		    \State Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$
+		    \State Compute $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
+%		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
+		    \hspace*{12pt} with partition $\mathcal{P}_\ell$ of \eqref{eq:partition}.\label{alg1:step2}
+		    \State Apply Bayes' Theorem to obtain $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
+		    \State Compute $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$ via \eqref{eq:convolve}.
 		    \EndFor
@@
 %Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
-For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
+For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
 Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
@@
 Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
-Our test data used the bicycle annotations in \cite{Robicquet2016}. 
-We did a 2-fold cross validation on 4 different scenes of varying complexity. 
-Our analysis used the same partitions of the trajectories for all three predictors.
+Our test data used a total  annotations in \cite{Robicquet2016}. 
+We did a 2-fold cross validation by using $20\%$ of the data for testing and the remainder for training.
+We used the same partitions of the trajectories for all three predictors.
+Our analysis was conducted on 4 different scenes of varying complexity with a total of $350$ total annotations.
 
@@
 			\hline
-			& Our Predictor & Random Walk & Kitani et al. \\ [0.5ex] 
+			& Our Algorithm & Random Walk & Kitani et al. \\ [0.5ex] 
 			\hline 
@@
 This paper presents a real-time approach to probabilistic pedestrian modeling. 
-We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with large additional contribtions in three critical areas: 1) prediction speed is highly efficient enabling tight loop real-time control;  2) our prediction uncertainty bound much more accurately reflect potential paths for the human limiting false negatives and maximize the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a much wider variety of real world applications.
-These three features have great promise for robotic applications and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces.
-For future directions we plan to optimize the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
+We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in three critical areas: 1) prediction speed is efficient enabling tight loop real-time control;  2) our error bound on the predictions accurately reflects potential paths for the human limiting false negatives and maximizing the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a wide variety of real world applications.
+These three features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces.
+We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
 

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
fixed equation after the partition equation

@@
 of $[-\bar{s}, \bar{s}]$ to compute a Riemann sum approximation using Theorem \ref{thm:main}.
-The partition $\mathcal{P}_\ell$ has a width of size, $\Delta s = \bar{s} / \ell$, and substitution in \eqref{eq:approximation 3} yields an approximation of $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ given by
+The partition $\mathcal{P}_\ell$ has a width of size, $\Delta s = \bar{s} / \ell$, and substitution in \eqref{eq:approximation 3} yields an approximation of $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at time $t=\ell \Delta t$
+given by
 \begin{align*}

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
fixed some oversights in the algorithm
fixed a math typo

@@
 		    \For {$\ell \in \{1,\dots,N_t\}$}
-		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
+		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$ and all $k=1,\dots, n$.
 		    \For {$m \in \{-\ell,\dots,\ell\}$}
-		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
+		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ for all $i,j=-N_x,\dots, N_x$ and $k=1,\dots,n$.
 		    \EndFor  
@@
 %Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
-For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
+For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t \Delta t \bar{s} ,\dots, N_t \Delta t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
 Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
trying to clarify algorithm 1

@@
 As we have already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
+Now if we want to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at $t=(\ell + 1 ) \Delta t$, then by reusing the earlier computation one would only need to compute $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for $m \in \{ -( \ell + 1 ), (\ell + 1 )\}$ along with the weights.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
-The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.
 
@@
 		    \For {$\ell \in \{1,\dots,N_t\}$}
-		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$ and all $k=1,\dots, n$.
+		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$ and \hspace*{12pt} for all $k \in  \{1,\dots, n\}$.
 		    \For {$m \in \{-\ell,\dots,\ell\}$}
-		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ for all $i,j=-N_x,\dots, N_x$ and $k=1,\dots,n$.
+		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ for all \hspace*{28pt} $i,j\in \{-N_x,\dots, N_x\}$ and $k\in \{1,\dots,n\}$.
 		    \EndFor  
@@
 %Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
+The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.
 For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t \Delta t \bar{s} ,\dots, N_t \Delta t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
added one detail to Ram's ammendment

@@
 As we have already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
-Now if we want to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at $t=(\ell + 1 ) \Delta t$, then by reusing the earlier computation one would only need to compute $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for $m \in \{ -( \ell + 1 ), (\ell + 1 )\}$ along with the weights.
+Now if we want to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at $t=(\ell + 1 ) \Delta t$, then by reusing the earlier computation one would only need to compute $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for $m \in \{ -( \ell + 1 ), (\ell + 1 )\}$ along with the weights to obtain an approximation using the partition $\mathcal{P}_{\ell + 1}$.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.

### 2017-05-05 response.tex
Paper revision

\documentclass[usenames,dvipsnames]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[letterpaper, total={6in, 8in}]{geometry}
\usepackage{color}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{booktabs}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\providecommand{\response}[1]{
\noindent
\noindent\colorbox{gray!20}{
\parbox{\textwidth}{
\setlength{\parskip}{.1in}
\setlength{\parindent}{.1in}
#1}
}
}

\title{\LARGE Real-Time Certified Probabilistic Pedestrian Forecasting \\ \textbf{Author's Response} }

\author{Henry O. Jacobs, Owen Hughes, \\ Matt Johnson-Roberson, and Ram Vasudevan }

\begin{document}
\maketitle

%\todo{More formal version of: We would like to thank the reviewers for their comments. The feedback is very helpful, and we have attempted to address everything.}

We would like to begin by thanking the reviewers and the editor for their careful reading and review of our paper.
The comments provided were incredibly helpful and insightful, and have strengthened the paper considerably.
Based on the recommendations of the reviewers and the editor, we have made several modifications to the paper.
%The changes made are presented below as a highlighted manuscript that indicates additions in green and removals in red.
\todo{The changes made are summarized in detail below and a highlighted manuscript is attached that indicates additions in red.}	
% See the response to reviewers for a detailed breakdown of the changes made.

\section*{Reviewer 1}
\begin{enumerate}

\begin{item}
\textbf{This paper presents a novel real-time probabilistic forecasting method
for pedestrian trajectories via observing the historical trajectories
in a particular scene. The problem is well motivated, formulated and
handled. The reported experimental evaluations show pivotal improvement
against the current state-of-the-art.}

Thank you for the positive assessment.
\end{item}

\begin{item}
\textbf{With the current form of the results it is hard to compare the
improvements in terms of accuracy gained. This can be overcome by
including results such as physical distance between predicted paths and
ground truth observations. Such matrices are provided in the baseline
models [7, 9, 13] with accuracy values in meters. }

Thank you for noticing the difficulty in comparing the improvements. We calculated the MHD from each point in the trajectory to sampled points from each distribution which is summarized in Figure \ref{MHD} below.

We feel that both plots add something to the analysis, so we've included the following passage to justify both our metrics.

\response{ In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction. We used the area under the curve of the ROC curve as our measure of this. The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety. The area under this curve is a standard measure of the quality of a predictor. Figure \ref{AUC} shows the analysis of the AUC of each algorithm versus time. In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are. Figure \ref{MHD} shows MHD plotted against time.}

\end{item}
\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{paper/figures/mhd_results.eps}
	\caption{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution. The method from [13] does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times. At long timescales the MHD to the trajectory of most algorithms converges. Our method increases positional uncertainty with time, and so [7] outperforms us because they know the end point and we do not, and [13] places too much significance on the positions of other pedestrians at large time scales.}
	\label{MHD}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{paper/figures/the_results.eps}
	\caption{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of [7] is due to their confidence in their initial estimate. We sampled the S-LSTM model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}
	\label{AUC}
\end{figure}

\begin{item}
\textbf{How the model can incorporate the interactions among pedestrians (i.e
group motion) and the influences from neighbouring pedestrians (for
instances such as collision avoidance)?}
\end{item}
Thank you for this insightful question. Social forces are a natural avenue for future work, and so we included the following paragraph in the paper.

\response{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.  In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  Incorporating a social force $F_{i}$ acting on the $i$th can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.}

\begin{item}
\textbf{Presentation and organisation
The motivation behind choosing start/end points based clustering is not
clear. Does clustering based on start/end points help to identify
different motion models? Isn't it better to cluster considering the
entire trajectory? }
\end{item}
This is a relevant question, thank you. As stated in Morris and Trivedi (below), our choice of clustering technique is not nearly as important as the chosen distance metric. Our success with distance functions that took a whole trajectory was limited on the data we tested, particularly when compared against our method. We have several hypotheses as to why this is true.  For one, the spatial scale of the data we test against showcases important features (e.g. curb cuts and storefronts), which lead to clustered start points and end points. Similarly, we tested against high quality data in which trajectories with irreconcilable occlusions were rare and thus could be omitted. See the comment below for the text we added to the paper.

\begin{item}
\textbf{Do you cluster different scenes provided in campus dataset [13] (i.e
Valley, Fangron) separately or all together? When motion models are
learnt, are they learnt in scene specific manner or altogether?}
\end{item}
Thank you for noticing that this is unclear. Our algorithm trains on each scene independently, so clustering, vector field learning, and learning the potential functions are all done for each Valley, Fangron, etc. We've added the following text to the paper:

\response{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}

\begin{item}
\textbf{The motivation behind choosing the affinity propagation clustering
algorithm is not clearly stated. 
What is the distance measure used (i.e Euclidian, Ö) ? Not specifying
such information makes reproduction of this work not feasible. }

This was an oversight on our part, thank you for noticing it. The distance used is actually a custom distance function on $\mathbb{R}^4$, which disregards the involution $(x_1, x_2, x_3, x_4) \to (x_3, x_4, x_1, x_2)$, and so doesn't care about parity of trajectories. We have included the following clarification:

\response{We then cluster in $\mathbb{R}^4$ using Affinity Propagation[17] and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$. This function measures the distance between the endpoints irrespective of their ordering. The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g. from a storefront to the sidewalk at the edge of a scene. On our data, other distance measures from \todo{[Morris and Trivedi]} and \todo{[Lee 2007]} didn't identify coherent motion models. \todo{come up with a good story for why these fail so spectacularly.}}

\end{item}

\begin{item}
\textbf{Adequacy of Citation
I believe literature review section can be improved using a sub section
on social force models, which is extensively applied for pedestrian
trajectory forecasting. 
Helbing and Molnar, 1995; Koppula and Saxena, 2013; Pellegrini et al.,
2010; Yamaguchi et al., 2011; Xu et al., 2012; Wang et al. 2008;
Hospedales et al. 2009; Emonet et al. 2011; Yi et al. 2015.}
\end{item}

We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature to reflect the relevant work.

\response{Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
	[yamaguchi] and [yi] do this by minimizing learned energy functions. }
\todo{FINISHTHIS}
\begin{item}
\textbf{Furthermore a comprehensive review on trajectory clustering algorithms
is required. This should be used to motivate the reason of choosing
affinity propagation algorithm in section 4. 
Morris and Trivedi 2009; Giannotti et al. 2007; Lee et al. 2007;
Ester et al. 1996 }

Thank you for this suggestion and for the list of relevant titles. We included a note about different algorithms as well as a justification for why we chose our metric. The note can be seen in point 6 above.

\end{item}
\end{enumerate}

\section*{Reviewer 2}
\begin{enumerate}
\begin{item}
\textbf{Keep the good work going. Interesting approach.}
\end{item}

Thank you for the kind comments.
\begin{item}
\textbf{
Need to be clearer and
dive a little bit deeper into the technical approach. }
Thank you for this criticism. We hope we adequately addressed it by strengthening the description of our method for learning the motion model, as well as quantifying the way that we tested our data. We also changed the language below to better reflect that the specific implementation we chose is an example of how to use the larger framework.

\response{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
\end{item}
\begin{item}
\textbf{
Need to use and apply more deep learning approaches such as social LSTM
or other machine methods.}
Thank you for the suggestion. We hope that we have addressed it by including a review of literature of social force models, including S-LSTM, as well as a section on how social factors can be integrated into our model. In addition, we compared against S-LSTM in addition to [7] from the paper. Thefigures drawing the predictions of the S-LSTM model are ommitted from the paper because like many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero. We changed the subtitle of the figure to reflect this, which is shown below.

\response{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent. The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.
}
\response{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle. The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}

\end{item}

\end{enumerate}

\section*{Reviewer 5}
\begin{enumerate}
\begin{item}
\textbf{
It would be more interesting if the authors show whether
the prediction algorithm would transfer the learned "knowledge" of
forecasting in novel scenes.}
Thank you for your comment. As it is, this is a shortcoming of our model, but we see it as an immediate avenue for future work. As can be read below, the environmental decomposition that the segmentation algorithm in [12] returns as well as an analogue of the methods [11] proposes for characterizing and transferring our vector fields from one scene to another.

\response{As future work, we propose using scene segmentation from [12], as well as the semantic context descriptors and routing scores from [11] to show how vector fields can be transferred to novel scenes. It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in [11].}

\end{item}
\begin{item}
\textbf{
More statistically result is expected
rather than analyzing "four different scenes", and comparing run-time
by "averaged across several agents and scenes" (how many?).}

Thank you for noticing this. We've clarified the papers in the passages listed below to clarify the list of scenes we ran in comparison, and how many agents total were run. We clarified that the set of agents used to evaluate were the same that were used to determine run times. Note that we attempted to compare on all of the scenes in the Stanford Drone Dataset, but the difficulty in getting the other models to converge in a reasonable amount of time prevented us from doing so.

\response{ Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [13], with a total of 142 trajectories analyzed.}

\response{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the data used in the quality analysis.}
\end{item}

\begin{item}
\textbf{
Besides, it is better to explain the meaning of ROC and AUC in section IV.}

This is a necessary clarification, thank you for the suggestion. We have included a description of ROC and AUC, as well as a justification for why we chose to evaluate the methods using this criterion. You can read the text that we added to address this in the second point made by Reviewer 1.

\end{item}
\end{enumerate}

% \section*{Associate Editor?}
% \begin{enumerate}
% \begin{item}
% \textbf{The manuscript presents an interesting approach toward pedestrian
% forecasting which is quite different than current approaches.}  
% \end{item}
% \begin{item}
% \textbf{The
% presentation could be improved by providing more concrete performance
% measures and comparison with more state-of-the-art.For example newer
% models that utilize social interactions.  Please see the reviewer
% comments for more detailed recommendations.}
% Thank you for the suggestions, we feel that they have added to the quality of the paper. As stated in the individual responses to reviewers, we tested using a variant of the MHD method used in [7], and found that while we still outperformed [7] and [13] in the intermediate time scales, it didn't capture our desire to test for safety as well as proximity to ground truth as well as the AUC of the ROC curves. The plot of measured MHD is included \todo{(?)}.

% The addition of a comparison to social interaction models was very valuable, and we thank you for the suggestion. We have included a section in our literature review on these models, as well as a comparison in section IV to the S-LSTM model from [13].

% \end{item}
% \end{enumerate}

\end{document}

### 2017-05-05 paper/pedestrian_forecasting_iros_2017.tex
Paper revision

--- 2017-02-16 pedestrian_forecasting_iros_2017.tex
@@
 \usepackage{algorithm}
+\usepackage{marginnote}
 %\usepackage{algorithm2e}
@@
 \newcommand{\Hoj}[1]{\textcolor{blue}{#1}}
+\newcommand{\rtext}[1]{\textcolor{red}{#1}}
 
@@
 \author{Henry O. Jacobs, Owen Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
+\thanks{This work was supported by Ford Motor Company.}
  \thanks{ H.O. Jacobs, O. Hughes, and R.~Vasudevan are with the Department of Mechanical Engineering, University of Michigan, Ann Arbor, MI 48109
@@
 \section{Implementation and Experimental results} \label{sec:implementation}
-  Given this model, we describe a specific implementation and the process of fitting the model to a dataset.
+  \rtext{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
   In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
@@
   %In principal, any clustering algorithm could accomplish this.
-  In particular, we use the start and end point for each trajectory to obtain a point in $\mathbb{R}^4$.
-  We then cluster in $\mathbb{R}^4$ using Affinity propagation \cite{FreyDueck2007}.
+\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation\cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
+This function measures the distance between the endpoints irrespective of their ordering.
+The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g.
+from a storefront to the sidewalk at the edge of a scene.
+On our data, other distance measures from [Morris and Trivedi] and [Lee 2007] didn't identify coherent motion models.}
   This clustering of the end-points induces a clustering of the trajectories.
@@
 
-  \subsection{Learning the Noise Model}
-  Finally, we assume that the true position is related to the model by Gaussian noise with a growing variance.
-  In particular, we assume $\Pr( x_t \mid \check{x}_t) \sim \mathcal{N}( \check{x}_t , \kappa t)$ for some constant $\kappa \geq 0$.
-  The parameter, $\kappa$, must be learned.
- % We've taken an adhoc approach in this paper, although more intelligent and theoretically sound alternatives likely exist.
-  For each curve in $S_k$ we create a synthetic curve using the initial position and speed and integrating the corresponding vector-field, $s\, X_k$.
-  So for each curve, $x_i(t)$, of $S_k$, we have a synthesized curve $x_{i,synth}(t)$ as well.
-  We then measure the standard deviation of $(x_i(t) - x_{i,synth}(t)) / t$ over $i$ and at few time, $t \in \{ 0, 100, 200 \}$ in order to obtain $\kappa$.
-
- \subsection{Evaluating Performance}
-
-This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
-We implement our model as well as our evaluation code in Python 2.6. 
-Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
-Our test data used a total  annotations in \cite{Robicquet2016}. 
-We did a 2-fold cross validation by using $20\%$ of the data for testing and the remainder for training.
-We used the same partitions of the trajectories for all three predictors.
-Our analysis was conducted on 4 different scenes of varying complexity with a total of $350$ total annotations.
-
   \begin{figure}
@@
 
-	\caption{An illustration of the predictions generated by the various algorithms.
-In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
-The likelihood of detection is depicted using the \textit{virdis} color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.}
+	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
+The algorithm from \cite{Robicquet2016} quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
 
@@
 
-	\caption{An illustration of the predictions generated by the various algorithms.
-In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
-The likelihood of detection is depicted using the \textit{virdis} color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.}
+	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Robicquet2016} passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
 
@@
 \end{figure}
+
+  \subsection{Learning the Noise Model}
+  Finally, we assume that the true position is related to the model by Gaussian noise with a growing variance.
+  In particular, we assume $\Pr( x_t \mid \check{x}_t) \sim \mathcal{N}( \check{x}_t , \kappa t)$ for some constant $\kappa \geq 0$.
+  The parameter, $\kappa$, must be learned.
+ % We've taken an adhoc approach in this paper, although more intelligent and theoretically sound alternatives likely exist.
+  For each curve in $S_k$ we create a synthetic curve using the initial position and speed and integrating the corresponding vector-field, $s\, X_k$.
+  So for each curve, $x_i(t)$, of $S_k$, we have a synthesized curve $x_{i,synth}(t)$ as well.
+  We then measure the standard deviation of $(x_i(t) - x_{i,synth}(t)) / t$ over $i$ and at few time, $t \in \{ 0, 100, 200 \}$ in order to obtain $\kappa$.
+
+ \subsection{Evaluating Performance}
+
+This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
+We implement our model as well as our evaluation code in Python 2.6. 
+Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
+Our test data used a total  annotations in \cite{Robicquet2016}. 
+\rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
+We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
+\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
 
@@
 Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
-Figure \ref{fig:auc_vs_time} shows the Area Under the Curve of these ROC curves versus time.
+\rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction.
+We used the area under the curve of the ROC curve as our measure of this.
+The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety.
+The area under this curve is a standard measure of the quality of a predictor.
+Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
+In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are.
+Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.}
 Our predictor behaves better than any of the other compared methods at moderately large $t$.
@@
 	\includegraphics[width=\linewidth]{./figures/the_results.eps}
-	\caption{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.}
+	\caption{\rtext{A comparison of the AUC of the various algorithms.
+Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
+We sampled the model from \cite{Robicquet2016} 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}}
 	\label{fig:auc_vs_time}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{./figures/MHD_results.eps}
+	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
+The method from \cite{Robicquet2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
+At long timescales the MHD to the trajectory of most algorithms converges.
+Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Robicquet2016} places too much significance on the positions of other pedestrians at large time scales.}}
+	\label{fig:mhd_vs_time}
 \end{figure}
@@
 
-The run time per frame for each algorithm was generated using total run time for 400 frames, averaged across several agents and scenes, divided by the number of frames. 
+\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the data used in the quality analysis.}
 This is shown in Table~\ref{tab:time}. 
@@
 We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
+\rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.
+In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
+Incorporating a social force $F_{i}$ acting on the $i$th can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
 

### 2017-05-05 response.tex
More

@@
 %The changes made are presented below as a highlighted manuscript that indicates additions in green and removals in red.
-\todo{The changes made are summarized in detail below and a highlighted manuscript is attached that indicates additions in red.}	
+The changes made are summarized in detail below and a highlighted manuscript is attached that indicates additions in red.
 % See the response to reviewers for a detailed breakdown of the changes made.
@@
 Incorporating a social force $F_{i}$ acting on the $i$th can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.}
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \todo{[Helbing and Molnar]}.}
 
@@
 algorithm is not clearly stated. 
-What is the distance measure used (i.e Euclidian, Ö) ? Not specifying
+What is the distance measure used (i.e Euclidian, \"{O}) ? Not specifying
 such information makes reproduction of this work not feasible. }
@@
 We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature to reflect the relevant work.
+\todo{Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
+	[yamaguchi] incorporates social grouping into a discrete-time dynamical system model of pedestrian movement using a learned energy function to determine the dynamics that incorporates collision/group details, and learns social groups using an improvement on [Pellegrini].
+	[Yi] does it by learning a scene energy map, a moving pedestrian energy map, and a stationary energy map by maximizing likelihood, then computes trajectories by minimizing this map between the start and the end point.
+	[Emonet] uses heirarchical dirichlet processes to infer behavioural motifs in videos. 
+	[Hospedales 2009] Uses an improvement on LDA, does online topic discovery, unclear if it does prediction.
+	[helbing and Molnar] describes pedestrian behaviour according to an interaction potential, causing pedestrians to avoid each other.
+	[Pellegrini]  uses a third-order CRF to group pedestrians based on behaviour. No prediction.
+	[robicquet]  uses LSTM to learn pedestrian behaviours, and predict based on them.
+	[zhou] }
 
-\response{Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
-	[yamaguchi] and [yi] do this by minimizing learned energy functions. }
-\todo{FINISHTHIS}
+\response{Along different lines, many authors have focused on using social interaction as a main part of their motion model.
+\todo{[yamaguchi]} incorporates social grouping into the energy function used to determine the velocity of a discrete-time dynamical system, whiel [Yi] incorporates pedestrians into an energy map, and calculate the prediction between two points as the lowest cost path.
+[helbing and molnar] introduced the idea of an actual Social Force, which was shown to be effective on modern data by \todo{[A.
+Lerner, Y.
+Chrysanthou]}, and \todo{[pellegrini 2010]}.
+\todo{[emonet]} and \todo{[hospedales]} both operate on raw video data, using Heirarchical Dirichlet Processes and an extension of LDA to find behaviours in scenes.
+Though these methods are good at detecting anomalous behaviour, without an actual underlying motion model their individual predictive power is limited.
+\todo{robicquet} uses a Long Short-Term Memory network to learn and predict social behaviours.
+}
+
 \begin{item}

### 2017-05-06 response.tex
More stuff

@@
 We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature to reflect the relevant work.
-\todo{Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
-	[yamaguchi] incorporates social grouping into a discrete-time dynamical system model of pedestrian movement using a learned energy function to determine the dynamics that incorporates collision/group details, and learns social groups using an improvement on [Pellegrini].
+
+Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
+	[yamaguchi] incorporates social grouping into a discrete-time dynamical system model of pedestrian movement using a learned energy function (minimizing over energy function to find velocity, local minimum) to determine the dynamics that incorporates collision/group details, and learns social groups using an improvement on [Pellegrini].
+Learns weight parameters by throwing simplex method at an LP version of the problem and hopes it converges on a local min.
+Scene obstacles are modeled as pedestrians.
+Destination forecasted by learned SVM on destinations of observed trajectories.
+Uses a classifier to determine groups.
+	Mentions:
+	Mentions Helbing, similarity to physics
+	"The assumption of discretized choice allows efficient prediction
+	with analytical solutions despite the large number of
+	factors considered in the model [1, 10]. However, due to the
+	nature of the discretization, the behavioral prediction tends
+	to show artifacts when metric is continuous."
+	[pellegrini: you'll never walk alone ] ``An advantage of continuous model is the flexibility of constructing
+	complex models, however, previous work focuses
+	on individual motivation of the behavior [you'll never walk alone, 11], and the
+	only social context is collision avoidance.''
+	[As in [you'll], our model assumes that each pedestrian makes a
+	decision on the velocity based on various environmental
+	and social factors in the scene]
+	Yamaguchi uses collision avoidance from [you'll]
+	[pellegrini: improving ]: Yamaguchi attempts a simpler method of social grouping, with better computation time.
+
+	Notes on CRFs from [Yamaguchi]: "The CRF formulation indirectly
+	encodes a behavioral model. Our focus is to build an
+	explicit behavioral model which can exploit the rich behavioral
+	context in social interactions,"
+
 	[Yi] does it by learning a scene energy map, a moving pedestrian energy map, and a stationary energy map by maximizing likelihood, then computes trajectories by minimizing this map between the start and the end point.
-	[Emonet] uses heirarchical dirichlet processes to infer behavioural motifs in videos. 
-	[Hospedales 2009] Uses an improvement on LDA, does online topic discovery, unclear if it does prediction.
-	[helbing and Molnar] describes pedestrian behaviour according to an interaction potential, causing pedestrians to avoid each other.
-	[Pellegrini]  uses a third-order CRF to group pedestrians based on behaviour. No prediction.
-	[robicquet]  uses LSTM to learn pedestrian behaviours, and predict based on them.
-	[zhou] }
+These are learned by using gradient ascent to maximize the likelihood of the data under the energy maps.
+Requires endpoint.
+No mention of real-time, or run-time.
+
+	Mentions:
+	[emonet]: Spatio-temporal dependency on motion patterns could be
+	included in topic models [emonet] - [31].
+	[hospedales] included in last
+	[helbing] and abnormal event detection [helbing].
+	[zhou] It can be used for various applications
+	including pedestrian walking path prediction [1], [zhou],
+	``Previous studies [zhou], [17]–[19] have shown that the walking
+	behavior of an individual can be influenced by a variety
+	of factors including scene layout (e.g. entrances, exits,
+	walls, and obstacles), inter-person variations on the choice
+	of source and destination, and interactions with other moving
+	pedestrians. ''
+	``A mixture model of dynamic
+	agents (MDA) is proposed by Zhou et al. [2], which can learn
+	parameters automatically. Typical agent-based models also
+	include the self-driven particle model [43], and the reciprocal
+	velocity obstacle model [44].''
+	``Most previous methods [2], [17], [18], [40]–[42] are local
+	models. Decisions are made based on local environments and
+	the interactions with nearby people. These models may make
+	reasonable decisions when there are not so many people in
+	the scene and the scene is not that complex.''
+	``. Two existing approaches,
+	i.e., the MDA model [zhou ]and an unsupervised visual prediction
+	approach (UVP) [54], are used for comparison.''
+	`` This is because only the motion pattern
+	information are modeled for MDA [2] and UVP [54], while the
+	influences of stationary objects (e.g. stationary crowd groups)
+	are ignored''
+	[wang: unsupervised] ``traffic
+	flow segmentation [3]–[5],''
+
+	[Emonet] uses heirarchical dirichlet processes on optical flow of downsampled video, discretize to 8 cardinal directions, then find the most common low-level words, and do heirarchical dirichlet processes to generate topics there.
+Can't predict individual pedestrians?
+	[Hospedales 2009] Uses an improvement on LDA, does online topic discovery in real time.
+Adds heirarchical modeling, to reflect how simple actions can be turned into behaviours.
+Doesn't have explicit motion model though, Doesn't necessarily work for a single pedestrian?
+	[helbing and Molnar] describes pedestrian behaviour according to an interaction potential, causing pedestrians to avoid each other. Doesn't include a complex motion model.
+
+	[alahi]  uses LSTM to learn pedestrian behaviours, and predict based on them.
+	[zhou] 
+	[wang1: Trajectory analysis]
+	[wang2: Unsupervised Activity Perception]
+	[Xu]
+	[Pellegrini1:Improving data]
+	[Pellegrini2: wrong turn]
+	[Pellegrini3 : you'll never walk alone]
+	[Koppula]
 
 	\response{Along different lines, many authors have focused on using social interaction as a main part of their motion model.
-\todo{[yamaguchi]} incorporates social grouping into the energy function used to determine the velocity of a discrete-time dynamical system, whiel [Yi] incorporates pedestrians into an energy map, and calculate the prediction between two points as the lowest cost path.
-[helbing and molnar] introduced the idea of an actual Social Force, which was shown to be effective on modern data by \todo{[A.
-Lerner, Y.
-Chrysanthou]}, and \todo{[pellegrini 2010]}.
+\todo{[yamaguchi]} incorporates social grouping into the energy function used to determine the velocity of a discrete-time dynamical system, while \todo{[Yi]} incorporates pedestrians into an energy map, and calculate the prediction between two points as the lowest cost path.
+Both of these approaches  [helbing and molnar] introduced the idea of an actual Social Force, which was shown to be effective on modern data by \todo{[]}, and \todo{[pellegrini 2010]}.
+\todo{[pellegrini 2010]} uses a stochastic linear trajectory avoidance.
 \todo{[emonet]} and \todo{[hospedales]} both operate on raw video data, using Heirarchical Dirichlet Processes and an extension of LDA to find behaviours in scenes.

### 2017-05-06 paper/pedestrian_forecasting_iros_2017.tex
More stuff

@@
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm from \cite{Robicquet2016} quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
+The algorithm from [asdfasdf] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
 
@@
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in \cite{Robicquet2016} passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
+The algorithm in [asdfasdf] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
 
@@
 Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
-Our test data used a total  annotations in \cite{Robicquet2016}. 
 \rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
-\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
+\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [asdfasdf], with a total of 142 trajectories analyzed.}
 
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the model from \cite{Robicquet2016} 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}}
+We sampled the model from [asdfasdf] 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}}
 	\label{fig:auc_vs_time}
@@
 	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
-The method from \cite{Robicquet2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
+The method from [asdfasdf] does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
 At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Robicquet2016} places too much significance on the positions of other pedestrians at large time scales.}}
+Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and [asdfasdf] places too much significance on the positions of other pedestrians at large time scales.}}
 	\label{fig:mhd_vs_time}

### 2017-05-06 response.tex
Add lit review

--- 2017-05-05 response.tex
@@
 On our data, other distance measures from \todo{[Morris and Trivedi]} and \todo{[Lee 2007]} didn't identify coherent motion models.
-\todo{come up with a good story for why these fail so spectacularly.}}
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, but the cumulative effect on the distance measures is enough to hinder the clustering.
+}
 
@@
 
-\response{Other classes of algorithm focus on predicting social behaviour based on  social interactions of people 
-	[yamaguchi] and [yi] do this by minimizing learned energy functions. }
-\todo{FINISHTHIS}
+	\response{On the other hand, authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
+Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated tools.
+Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] which incorporates collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of Linear Trajectory Avoidance (LTA).
+However LTA suffers from not planning for other pedestrian’s positions at future times.
+[emonet], [hospedales], and [wang: unsupervised] all take optical flow as input.
+[wang] and [emonet] both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
+classes of motion within the scene), where [hospedales] substitutes a Markov model.
+These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
+[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate human affordances relating to human objectives within the scene.
+[Tay], [Wang jw], and [Trautman] create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
+[alahi] uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
+[alahi] outperformed LTA, and social forcing based on [yamaguchi]  and IGP from [trautman].
+Along with its quick run time, this motivated our decision to test our algorithm against it.
+}
+
 \begin{item}

### 2017-05-06 paper/pedestrian_forecasting_iros_2017.tex
Add lit review

@@
 However, since the motion model was a Markov jump process which required the application of a sample based technique, we should expect the same painful trade-off between error and speed to occur as in \cite{Karasev2016} and \cite{Ballan2016}.
+
+\rtext{On the other hand, authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
+Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated tools.
+Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] which incorporates collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of Linear Trajectory Avoidance (LTA).
+However LTA suffers from not planning for other pedestrian’s positions at future times.
+[emonet], [hospedales], and [wang: unsupervised] all take optical flow as input.
+[wang] and [emonet] both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
+classes of motion within the scene), where [hospedales] substitutes a Markov model.
+These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
+[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate human affordances relating to human objectives within the scene.
+[Tay], [Wang jw], and [Trautman] create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
+[alahi] uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
+[alahi] outperformed LTA, and social forcing based on [yamaguchi]  and IGP from [trautman].
+Along with its quick run time, this motivated our decision to test our algorithm against it.
+}
 

### 2017-05-07 paper/response.tex
Bibliography shit

--- 2017-05-06 response.tex
@@
 The changes made are summarized in detail below and a highlighted manuscript is attached that indicates additions in red.
+\textbf{This letter is largely self contained, so references here are to the bibliography in this file}.
 % See the response to reviewers for a detailed breakdown of the changes made.
@@
 ground truth observations. Such matrices are provided in the baseline
-models [7, 9, 13] with accuracy values in meters. }
+models \cite{Kitani2012} \cite{Karasev2016} \cite{Robicquet2016} with accuracy values in meters. }
 
 Thank you for noticing the difficulty in comparing the improvements.
-We calculated the MHD from each point in the trajectory to sampled points from each distribution which is summarized in Figure \ref{MHD} below.
+We addressed this by adding a comparison to the MHD from .
+Specifically, we calculated the MHD from each point in the trajectory to sampled points from each distribution.
+The results aresummarized in Figure \ref{MHD} below.
 
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{paper/figures/mhd_results.eps}
+	\includegraphics[width=0.6\textwidth]{figures/mhd_results.eps}
 	\caption{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
-The method from [13] does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
+The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
 At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so [7] outperforms us because they know the end point and we do not, and [13] places too much significance on the positions of other pedestrians at large time scales.}
+Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}
 	\label{MHD}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{paper/figures/the_results.eps}
+	\includegraphics[width=0.6\textwidth]{figures/the_results.eps}
 	\caption{A comparison of the AUC of the various algorithms.
-Note that the initial dip in the performance of [7] is due to their confidence in their initial estimate.
-We sampled the S-LSTM model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}
+Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
+We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
 	\label{AUC}
@@
 
-\response{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.
+\response{Though including social forces into the model is out of the scope of this paper, its incorporation could begin as follows.
 In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
-Incorporating a social force $F_{i}$ acting on the $i$th can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \todo{[Helbing and Molnar]}.}
+Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995} .}
 
@@
 \begin{item}
-\textbf{Do you cluster different scenes provided in campus dataset [13] (i.e
+\textbf{Do you cluster different scenes provided in campus dataset \cite{Alahi2016} (i.e
 Valley, Fangron) separately or all together? When motion models are
@@
 
-\response{We then cluster in $\mathbb{R}^4$ using Affinity Propagation[17] and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
+\response{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
 This function measures the distance between the endpoints irrespective of their ordering.
@@
 from a storefront to the sidewalk at the edge of a scene.
-On our data, other distance measures from \todo{[Morris and Trivedi]} and \todo{[Lee 2007]} didn't identify coherent motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, but the cumulative effect on the distance measures is enough to hinder the clustering.
+On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models.
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.
 }
@@
 
-We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature to reflect the relevant work.
+We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature review to reflect the relevant work.
 
-	\response{On the other hand, authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
-Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated tools.
-Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] which incorporates collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of Linear Trajectory Avoidance (LTA).
+	\response{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
+Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods.
+Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
 However LTA suffers from not planning for other pedestrian’s positions at future times.
-[emonet], [hospedales], and [wang: unsupervised] all take optical flow as input.
-[wang] and [emonet] both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
-classes of motion within the scene), where [hospedales] substitutes a Markov model.
+\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
+\cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
+classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
 These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
-[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate human affordances relating to human objectives within the scene.
-[Tay], [Wang jw], and [Trautman] create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
-[alahi] uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
-[alahi] outperformed LTA, and social forcing based on [yamaguchi]  and IGP from [trautman].
-Along with its quick run time, this motivated our decision to test our algorithm against it.
+\cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
+\cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
+\cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
+\cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}.
+Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.
 }
@@
 We included a note about different algorithms as well as a justification for why we chose our metric.
+We were unable to test against the methods of Giannotti and Ester.
 The note can be seen in point 6 above.
@@
 Thank you for the suggestion.
-We hope that we have addressed it by including a review of literature of social force models, including S-LSTM, as well as a section on how social factors can be integrated into our model.
-In addition, we compared against S-LSTM in addition to [7] from the paper.
-Thefigures drawing the predictions of the S-LSTM model are ommitted from the paper because like many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero.
+We hope that we have addressed it by including a review of literature of social motivated models, including S-LSTM, as well as a section on how social factors can be integrated into our model.
+In addition, we compared against S-LSTM in addition to \cite{Kitani2012} from the paper.
+The figures corresponding to the predictions of the S-LSTM model are ommitted from the paper because as happened for many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero.
 We changed the subtitle of the figure to reflect this, which is shown below.
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
+The algorithm from \cite{Alahi2016} quickly passes beyond the boundaries of the scene, so the plots are ommitted here.
 }
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} passes beyond the boundaries of the scene, and so the plots are ommitted here.}
 
@@
 As it is, this is a shortcoming of our model, but we see it as an immediate avenue for future work.
-As can be read below, the environmental decomposition that the segmentation algorithm in [12] returns as well as an analogue of the methods [11] proposes for characterizing and transferring our vector fields from one scene to another.
+As can be read below, the environmental decomposition that the segmentation algorithm in \cite{Walker2014} returns as well as an analogue of the methods \cite{Ballan2016} proposes for characterizing and transferring vector fields from one scene to another could reasonably be applied to our method..
 
-\response{As future work, we propose using scene segmentation from [12], as well as the semantic context descriptors and routing scores from [11] to show how vector fields can be transferred to novel scenes.
-It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in [11].}
+\response{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
 
@@
 Thank you for noticing this.
-We've clarified the papers in the passages listed below to clarify the list of scenes we ran in comparison, and how many agents total were run.
+We have clarified the paper in the passages listed below to clarify the list of scenes we ran in comparison, and how many agents total were run.
 We clarified that the set of agents used to evaluate were the same that were used to determine run times.
@@
 
-\response{ Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [13], with a total of 142 trajectories analyzed.}
+\response{ Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
 
-\response{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the data used in the quality analysis.}
+\response{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
 \end{item}
@@
 % \end{enumerate}
-
+\bibliographystyle{IEEEtran}
+\bibliography{references}
 \end{document}

### 2017-05-07 paper/pedestrian_forecasting_iros_2017.tex
Bibliography shit

@@
 
-\rtext{On the other hand, authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
-Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated tools.
-Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] which incorporates collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of Linear Trajectory Avoidance (LTA).
+\rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
+Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated methods.
+Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
 However LTA suffers from not planning for other pedestrian’s positions at future times.
@@
 These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
-[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate human affordances relating to human objectives within the scene.
+[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
 [Tay], [Wang jw], and [Trautman] create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
 [alahi] uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
-[alahi] outperformed LTA, and social forcing based on [yamaguchi]  and IGP from [trautman].
-Along with its quick run time, this motivated our decision to test our algorithm against it.
+[alahi] outperformed LTA, the social forcing based on [yamaguchi],  and IGP from [trautman].
+Along with its quick run time, this establishes [alahi] as state-of-the-art for socially based models, and so we our algorithm to theirs.
 }
@@
   %In principal, any clustering algorithm could accomplish this.
-\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation\cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
+\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation[17] and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
 This function measures the distance between the endpoints irrespective of their ordering.
@@
 from a storefront to the sidewalk at the edge of a scene.
-On our data, other distance measures from [Morris and Trivedi] and [Lee 2007] didn't identify coherent motion models.}
+On our data, other distance measures from \todo{[Morris and Trivedi]} and \todo{[Lee 2007]} didn't identify coherent motion models.
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
+We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
   This clustering of the end-points induces a clustering of the trajectories.
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm from [asdfasdf] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
+The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
 
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in [asdfasdf] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
 
@@
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
-\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [asdfasdf], with a total of 142 trajectories analyzed.}
+\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [13], with a total of 142 trajectories analyzed.}
 
@@
 At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and [asdfasdf] places too much significance on the positions of other pedestrians at large time scales.}}
+Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and [asdfasdf] places too much significance on the initial positions of other pedestrians at large time scales.}}
 	\label{fig:mhd_vs_time}
@@
 		\renewcommand{\arraystretch}{1.5}% Spread rows out...
-		\begin{tabular}{||c | c c  c ||} 
+		\begin{tabular}{||c | c c  c c ||} 
 			\hline
-			& Our Algorithm & Random Walk & Kitani et al. \\ [0.5ex] 
+			& Our Algorithm & Random Walk & Kitani et al. & S-LSTM \\ [0.5ex] 
 			\hline 
-			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.0169s & 2.1E-7s & 0.0614s \\
+			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.00443s & 2.1E-7s & 0.0706s & 0.0134s \\
 			\hline
@@
 
-\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the data used in the quality analysis.}
+\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
 This is shown in Table~\ref{tab:time}. 
@@
 In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
-Incorporating a social force $F_{i}$ acting on the $i$th can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
+Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
 Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.

### 2017-05-07 paper/pedestrian_forecasting_iros_2017.tex
All citations should be resolved for now.

@@
 \rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
-Early work by [helbig] and [Xu] describe pedestrians interactions using physically motivated methods.
-Several models [yamaguchi, yi, pellegrini you’ll never] derive their motion models from [helbing] who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
+Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods.
+Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
 However LTA suffers from not planning for other pedestrian’s positions at future times.
-[emonet], [hospedales], and [wang: unsupervised] all take optical flow as input.
-[wang] and [emonet] both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
-classes of motion within the scene), where [hospedales] substitutes a Markov model.
+\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
+\cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
+classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
 These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
-[koppula and saxena] predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
-[Tay], [Wang jw], and [Trautman] create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
-[alahi] uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
-[alahi] outperformed LTA, the social forcing based on [yamaguchi],  and IGP from [trautman].
-Along with its quick run time, this establishes [alahi] as state-of-the-art for socially based models, and so we our algorithm to theirs.
-}
+\cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
+\cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories.
+\cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
+\cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}.
+Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.}
 
@@
   %In principal, any clustering algorithm could accomplish this.
-\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation[17] and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
+\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
 This function measures the distance between the endpoints irrespective of their ordering.
@@
 from a storefront to the sidewalk at the edge of a scene.
-On our data, other distance measures from \todo{[Morris and Trivedi]} and \todo{[Lee 2007]} didn't identify coherent motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
+On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models.
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.
+We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
@@
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
-\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from [13], with a total of 142 trajectories analyzed.}
+\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
 
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the model from [asdfasdf] 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large lowered their performance.}}
+We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.Though including social forces into the model is out of the scope of this paper, its incorporation could begin as follows.
+In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
+Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995} .}}
 	\label{fig:auc_vs_time}
@@
 	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
-The method from [asdfasdf] does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
+The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
 At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and [asdfasdf] places too much significance on the initial positions of other pedestrians at large time scales.}}
+Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}}
 	\label{fig:mhd_vs_time}

### 2017-05-07 paper/pedestrian_forecasting_iros_2017.tex
Preliminary numbers

@@
 
+\marginnote{\#1.1}
 \rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians.
@@
 \section{Implementation and Experimental results} \label{sec:implementation}
+\marginnote{\#4.1}
   \rtext{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
@@
   %In principal, any clustering algorithm could accomplish this.
+  \marginnote{\#4.2}
 \rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
@@
 On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.
-We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
-We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.}
   This clustering of the end-points induces a clustering of the trajectories.
@@
 
-	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+	\caption{
+		\rtext{An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
@@
 The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
-
+	\marginnote{\#4.7}
 	\label{fig:bookstore-1-2}
@@
 The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
-
+	\reversemarginpar\marginnote{\#4.8}
 	\label{fig:death-1-2}
@@
 Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
+\reversemarginpar\marginnote{\#4.3}
 \rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene.
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
+\reversemarginpar\marginnote{\#4.4}
 \rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
@@
 Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
-\rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction.
+\reversemarginpar\marginnote{\#4.5} \rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction.
 We used the area under the curve of the ROC curve as our measure of this.
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.Though including social forces into the model is out of the scope of this paper, its incorporation could begin as follows.
-In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
-Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995} .}}
+We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on ourdated pedestrian measurements at moderate-to-large timescales lowered their performance.}}
+	\normalmarginpar\marginnote{\#4.9}
 	\label{fig:auc_vs_time}
@@
 Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}}
+	\normalmarginpar\marginnote{\#4.10}
 	\label{fig:mhd_vs_time}
@@
 
+\normalmarginpar\marginnote{\#4.6}
 \rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
@@
 We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
-\rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.
+\reversemarginpar\marginnote{\#5.1}\rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.
 In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.
 Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \reversemarginpar\marginnote{\#5.2}\rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
 It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}

### 2017-05-07 paper/response.tex
Preliminary numbers
made lots of comments and changes to the response letter

@@
 \usepackage{booktabs}
+\usepackage{marginnote}
+
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
+\newcommand{\Ram}[1]{{\normalsize{\textbf{({\color{green}Ram:\ }#1)}}}}
 
@@
 
-\begin{item}
+\begin{item} \label{newmetrics}
 \textbf{With the current form of the results it is hard to compare the
@@
 
-Thank you for noticing the difficulty in comparing the improvements.
-We addressed this by adding a comparison to the MHD from .
-Specifically, we calculated the MHD from each point in the trajectory to sampled points from each distribution.
-The results aresummarized in Figure \ref{MHD} below.
+Thank you for the suggestion. 
+The objective of this paper was to develop a predictor that would be effective for use in the autonomous vehicle context. 
+In this instance, to ensure safety it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
+Since the ROC curve plots the true positive rate (i.e.
+the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e.
+the proportion of the generated set that does not correspond to the ground truth trajectory), it evaluates this aforementioned safety criteria for autonomous vehicles exactly.
 
-We feel that both plots add something to the analysis, so we've included the following passage to justify both our metrics.
+On the other hand the Modified Hausdorff Distance (MHD) from \Ram{which paper and in what form} measures the average geometric distance between the ground-truth observed trajectory and the generated set of predictions.
+Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
+Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
+If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
+Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+Whereas the MHD would describe this generated set of predictions as informative.
 
-\response{ In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction.
-We used the area under the curve of the ROC curve as our measure of this.
-The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety.
-The area under this curve is a standard measure of the quality of a predictor.
+Nevertheless, we addressed this concern by adding a comparison to the MHD which is summarized in the Figure \ref{MHD} below and included in the resubmitted version of our manuscript. 
+\Ram{it would be useful if we summarized what we were seeing the MHD plot here.
+Meaning it would be helpful if you could say something to the effect that with respect to MHD we are good in the medium term.
+In addition it would be nice if we could see some sample trajectories from lstm to really illustrate that they aren't including the ground truth trajectory but are just close geometrically which contributes to their low MHD score.
+That image should be included in the resubmitted text and in this response letter}
+We have included the following passage to explain both metrics in the revised submission:
+
+\reversemarginpar\marginnote{\#4.5}
+\response{ In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
+In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
+ROC curves which plot the true positive rate (i.e.
+the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e.
+the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly.
+The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
+The closer that this AUC is to one, the better the prediction algorithm.
 Figure \ref{AUC} shows the analysis of the AUC of each algorithm versus time. 
-In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are.
-Figure \ref{MHD} shows MHD plotted against time.}
+In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+Figure \ref{MHD} shows MHD plotted against time.
+Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
+Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
+If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
+Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+Whereas the MHD would describe this generated set of predictions as informative.
+}
 
@@
 	\includegraphics[width=0.6\textwidth]{figures/mhd_results.eps}
-	\caption{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
-The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
-At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}
+	\caption{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point samples from each distribution.
+The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
+At longer timescales the MHD to the trajectory of most algorithms converges.
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms which assume that the end point of the trajectory is unknown.}
+   \reversemarginpar\marginnote{\#4.10}
 	\label{MHD}
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times to give them the best chance in this analysis \Ram{why does sampling a $100$ times give them the best chance?}, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
+	\reversemarginpar\marginnote{\#4.9}
 	\label{AUC}
@@
 \end{item}
-Thank you for this insightful question. Social forces are a natural avenue for future work, and so we included the following paragraph in the paper.
+Thank you for this insightful question. 
+Since the focus of this paper was on developing a method to generate predictions at real-time for the autonomous vehicle context, we did not consider the effect of incorporating interactions between predictions.
+However, social forces are a natural avenue for future work, and so we included the following paragraph in the paper:
 
-\response{Though including social forces into the model is out of the scope of this paper, its incorporation could begin as follows.
+\response{Though including social forces into the model is outside of the scope of this paper, its incorporation could begin as follows.  
 In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
 Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995} .}
+Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995}. 
+The algorithms for generating real-time prediction would then generalize to this new vector-field definition.
+}
 
@@
 \end{item}
-This is a relevant question, thank you.
-As stated in Morris and Trivedi (below), our choice of clustering technique is not nearly as important as the chosen distance metric.
-Our success with distance functions that took a whole trajectory was limited on the data we tested, particularly when compared against our method.
+Thank you for this relevant question.
+As stated in \cite{Morris2009}, the choice of clustering technique is not nearly as important as the distance metric. 
+Our success with distance functions that took a whole trajectory was limited on the data we tested and we found more success using a custom distance metric that only considered the start and end points of a trajectory as we describe in Comment \ref{distmetric} to Reviewer 1.
 We have several hypotheses as to why this is true.  
-For one, the spatial scale of the data we test against showcases important features (e.g.
-curb cuts and storefronts), which lead to clustered start points and end points.
+For one, the spatial scale of the data we test against showcases important features (e.g. curb cuts and storefronts), which lead to clustered start points and end points. 
 Similarly, we tested against high quality data in which trajectories with irreconcilable occlusions were rare and thus could be omitted.
-See the comment below for the text we added to the paper.
 
@@
 Thank you for noticing that this is unclear. 
-Our algorithm trains on each scene independently, so clustering, vector field learning, and learning the potential functions are all done for each Valley, Fangron, etc.
-We've added the following text to the paper:
+Our algorithm trains on each scene independently, so clustering, vector field learning, and learning the potential functions are all done for each scene separately. 
+To make this more clear, we added the following text to the paper:
 
+\reversemarginpar\marginnote{\#4.3}
 \response{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
@@
 
-\begin{item}
-\textbf{The motivation behind choosing the affinity propagation clustering
-algorithm is not clearly stated. 
-What is the distance measure used (i.e Euclidian, \"{O}) ? Not specifying
-such information makes reproduction of this work not feasible. }
+\begin{item} \label{distmetric}
+\textbf{The motivation behind choosing the affinity propagation clustering algorithm is not clearly stated. 
+What is the distance measure used (i.e Euclidian) ? 
+Not specifying such information makes reproduction of this work not feasible. }
 
 This was an oversight on our part, thank you for noticing it. 
-The distance used is actually a custom distance function on $\mathbb{R}^4$, which disregards the involution $(x_1, x_2, x_3, x_4) \to (x_3, x_4, x_1, x_2)$, and so doesn't care about parity of trajectories.
+The distance used is a custom distance function on $\mathbb{R}^4$, which disregards the involution $(x_1, x_2, x_3, x_4) \to (x_3, x_4, x_1, x_2)$, and so does not care about parity of trajectories. 
+\Ram{You are using words that roboticists do not know (e.g. involution).  
+Can you please explicitly give the distance function and explain its properties and why you selected for those properties?
+For example you say parity of trajectories.
+What does parity mean and why is it important that the distance function ignore it?
+In your response below what is $x_1,x_2,x_3,x_4$ correspond to in the scene? 
+It may make more sense to use $(x_1,y_1)$ and $(x_2,y_2)$ since this is more indicative of the relationship between the points.
+What is $\bf{y}$ in your definition below?
+Please be more clear, again its easier for us to reduce the text then it is for us to add.}
 We have included the following clarification:
 
+\reversemarginpar\marginnote{\#4.2}
 \response{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
@@
 from a storefront to the sidewalk at the edge of a scene.
-On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.
+On this dataset, other distance measures \Ram{are these distance functions that utilize the entire trajectory?} from \cite{Morris2009} and \cite{Lee2007} didn't \Ram{avoid contractions} identify coherent \Ram{what does coherent mean?} motion models.
+This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering \Ram{what does hinder clustering mean?
+Please be more explicit}.
 }
@@
 
-We agree that this adds to the paper, thank you for the suggestion. We added a section to our literature review to reflect the relevant work.
+We agree that this adds to the paper, thank you for the suggestion. 
+We added the following section to our literature review to reflect the relevant work:
 
+\reversemarginpar\marginnote{\#1.1}
 	\response{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians. 
 	Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. 
-Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
-However LTA suffers from not planning for other pedestrian’s positions at future times.
+	Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporates collision avoidance through an interaction potential.
+	However this method suffers from not planning for other pedestrian'€ positions at future times. 
 	\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
@@
 classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
-These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
+	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
 	 \cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
 	 \cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. 
-\cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
-\cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}.
-Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.
+	 \cite{Alahi2016} uses Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents will interact and is the state-of-the-art algorithm for socially based models while also having a quick run time.
 }
@@
 Thank you for this suggestion and for the list of relevant titles.
-We included a note about different algorithms as well as a justification for why we chose our metric.
-We were unable to test against the methods of Giannotti and Ester.
-The note can be seen in point 6 above.
+ We included a note about different algorithms as well as a justification for why we chose our metric which is described in Comment \ref{distmetric} to Reviewer 1. 
+ %We were unable to test against the methods of Giannotti and Ester.\Ram{why not?}.
 
@@
 Thank you for the kind comments.
+
 \begin{item}
@@
 dive a little bit deeper into the technical approach. }
+
 Thank you for this criticism. 
-We hope we adequately addressed it by strengthening the description of our method for learning the motion model, as well as quantifying the way that we tested our data.
+We hope we adequately addressed it by strengthening the description of our method for learning the motion model, as well as quantifying the way that we tested our data as is described in our responses to the suggestions by other reviewers.
+
+\Ram{I don't think we need this addition to the text.}  
 We also changed the language below to better reflect that the specific implementation we chose is an example of how to use the larger framework.
 
+\reversemarginpar\marginnote{\#4.1}
 \response{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
 \end{item}
+
 \begin{item}
@@
 or other machine methods.}
+
 Thank you for the suggestion. 
 We hope that we have addressed it by including a review of literature of social motivated models, including S-LSTM, as well as a section on how social factors can be integrated into our model. 
-In addition, we compared against S-LSTM in addition to \cite{Kitani2012} from the paper.
-The figures corresponding to the predictions of the S-LSTM model are ommitted from the paper because as happened for many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero.
+In addition, we compared our approach against S-LSTM in addition to the method proposed in \cite{Kitani2012}. 
+The figures corresponding to the predictions of the S-LSTM model are omitted from the paper because as happened for many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero \Ram{I don't like this.
+We should include these predictions somehow anyway.}.
 We changed the subtitle of the figure to reflect this, which is shown below.
 
+\reversemarginpar\marginnote{\#4.8}
 \response{An illustration of the predictions generated by the various algorithms.
@@
 }
-\response{An illustration of the predictions generated by the various algorithms.
-In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
-The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in \cite{Alahi2016} passes beyond the boundaries of the scene, and so the plots are ommitted here.}
 
@@
 forecasting in novel scenes.}
+
 Thank you for your comment. 
-As it is, this is a shortcoming of our model, but we see it as an immediate avenue for future work.
-As can be read below, the environmental decomposition that the segmentation algorithm in \cite{Walker2014} returns as well as an analogue of the methods \cite{Ballan2016} proposes for characterizing and transferring vector fields from one scene to another could reasonably be applied to our method..
-
-\response{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+Though this was not the focus of this paper, the low-dimensional parameterization of the vector fields make transfer learning more amenable. 
+As it is, this is as an immediate avenue for future work which we describe in the following addition to the text: 
+\response{We also hope to do transfer learning \Ram{I flipped the previous two words.} with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
 It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
@@
 
-Thank you for noticing this.
-We have clarified the paper in the passages listed below to clarify the list of scenes we ran in comparison, and how many agents total were run.
-We clarified that the set of agents used to evaluate were the same that were used to determine run times.
+Thank you for noticing this oversight on our part. 
+We have modified the paper in the passages listed below to clarify the list of scenes we ran in comparison, and how many agents total were run. 
+The set of agents used during evaluation were the same that were used to determine run times. 
 Note that we attempted to compare on all of the scenes in the Stanford Drone Dataset, but the difficulty in getting the other models to converge in a reasonable amount of time prevented us from doing so.
 
+\reversemarginpar\marginnote{\#4.4}
 \response{ Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
 
+\reversemarginpar\marginnote{\#4.6}
 \response{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
@@
 This is a necessary clarification, thank you for the suggestion. 
-We have included a description of ROC and AUC, as well as a justification for why we chose to evaluate the methods using this criterion.
-You can read the text that we added to address this in the second point made by Reviewer 1.
+We have included a description of ROC and AUC, as well as a justification for why we chose to evaluate the methods using this criterion in our response to Comment \ref{newmetrics} for Reviewer 1.
 

### 2017-05-07 paper/response.tex
stupid typos fixed
some more typos fixed

@@
 \end{item}
+
 Thank you for this insightful question. 
@@
 \end{item}
+
 Thank you for this relevant question.
@@
 \end{item}
+
 Thank you for noticing that this is unclear. 
@@
 It would be more interesting if the authors show whether
-the prediction algorithm would transfer the learned "knowledge" of
+the prediction algorithm would transfer the learned ``knowledge" of
 forecasting in novel scenes.}
@@
 More statistically result is expected
-rather than analyzing "four different scenes", and comparing run-time
+rather than analyzing ``four different scenes", and comparing run-time
 by "averaged across several agents and scenes" (how many?).}

### 2017-05-08 paper/response.tex
Home stretch

--- 2017-05-07 paper/response.tex
@@
 \newcommand{\Ram}[1]{{\normalsize{\textbf{({\color{green}Ram:\ }#1)}}}}
-
 \providecommand{\response}[1]{
@@
 
-\author{Henry O. Jacobs, Owen Hughes, \\ Matt Johnson-Roberson, and Ram Vasudevan }
+\author{Henry O. Jacobs, Owen K. Hughes, \\ Matt Johnson-Roberson, and Ram Vasudevan }
 
@@
 
-On the other hand the Modified Hausdorff Distance (MHD) from \Ram{which paper and in what form} measures the average geometric distance between the ground-truth observed trajectory and the generated set of predictions.
+On the other hand the Modified Hausdorff Distance (MHD) from \cite{Dubuisson1994} measures the average geometric distance between the ground-truth observed trajectory and the generated set of predictions.
 Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
@@
 Nevertheless, we addressed this concern by adding a comparison to the MHD which is summarized in the Figure \ref{MHD} below and included in the resubmitted version of our manuscript.
-\Ram{it would be useful if we summarized what we were seeing the MHD plot here.
+We outperform \cite{Kitani2012} and \cite{Alahi2016} in the intermediate term, but at short times \cite{Alahi2016} puts more weight on the direction of the observed trajectory, and so their narrow distribution is closer to the trajectory.
+At large scales, \cite{Kitani2012} outperforms us because they know the endpoint of the trajectory.
+If you look at Figure \ref{bookstore} and \ref{death}, you'll see that while they are geometrically closer to the observed trajectory, they don't include it.
+%\Ram{it would be useful if we summarized what we were seeing the MHD plot here.
 Meaning it would be helpful if you could say something to the effect that with respect to MHD we are good in the medium term.
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model $100$ times to give them the best chance in this analysis \Ram{why does sampling a $100$ times give them the best chance?}, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times in order to organically extract a less concentrated probability distribution from their algorithm, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
 	\reversemarginpar\marginnote{\#4.9}
@@
 This was an oversight on our part, thank you for noticing it. 
-The distance used is a custom distance function on $\mathbb{R}^4$, which disregards the involution $(x_1, x_2, x_3, x_4) \to (x_3, x_4, x_1, x_2)$, and so does not care about parity of trajectories. 
-\Ram{You are using words that roboticists do not know (e.g. involution).  
-Can you please explicitly give the distance function and explain its properties and why you selected for those properties?
-For example you say parity of trajectories.
-What does parity mean and why is it important that the distance function ignore it?
-In your response below what is $x_1,x_2,x_3,x_4$ correspond to in the scene? 
-It may make more sense to use $(x_1,y_1)$ and $(x_2,y_2)$ since this is more indicative of the relationship between the points.
-What is $\bf{y}$ in your definition below?
-Please be more clear, again its easier for us to reduce the text then it is for us to add.}
+The distance used is a custom distance function.
+Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
+We define the points $\mathbf{a}_1 = (x_{A, \mathrm{start}}, y_{A, \mathrm{start}}, x_{A, \mathrm{end}}, y_{A, \mathrm{end}})$, $\mathbf{a}_2 = (x_{A, \mathrm{end}}, y_{A, \mathrm{end}}, x_{A, \mathrm{start}}, y_{A, \mathrm{start}})$, and $\mathbf{b} = (x_{B, \mathrm{start}}, y_{B, \mathrm{start}}, x_{B, \mathrm{end}}, y_{B, \mathrm{end}})$.
+We then define our distance measure as $d(A, B) :=  \min \left\{ d_e(\mathbf{a}_1, \mathbf{b}), d_e(\mathbf{a}_2, \mathbf{b}) \right\}$, for the euclidean distance $d_e$ in $\mathbb{R}^4$.
+This is designed such that transposing endpoints of a trajectory doesn't matter, and so a trajectory that starts at point A and ends at point B is close to a trajectory that starts at point B and ends at point A.
 We have included the following clarification:
@@
 \reversemarginpar\marginnote{\#4.2}
-\response{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
-This function measures the distance between the endpoints irrespective of their ordering.
+\response{
+	We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
+	We define the points $\mathbf{a}_1 = (x_{A, \mathrm{start}}, y_{A, \mathrm{start}}, x_{A, \mathrm{end}}, y_{A, \mathrm{end}})$, $\mathbf{a}_2 = (x_{A, \mathrm{end}}, y_{A, \mathrm{end}}, x_{A, \mathrm{start}}, y_{A, \mathrm{start}})$, and $\mathbf{b} = (x_{B, \mathrm{start}}, y_{B, \mathrm{start}}, x_{B, \mathrm{end}}, y_{B, \mathrm{end}})$.
+	We then define our distance measure as $d(A, B) :=  \min \left\{ d_e(\mathbf{a}_1, \mathbf{b}), d_e(\mathbf{a}_2, \mathbf{b}) \right\}$, for the euclidean distance $d_e$ in $\mathbb{R}^4$. 
+	This function measures the distance between the endpoints irrespective of their ordering, which means that a trajectory that starts from point A and ends at point B will be close to a trajectory that starts from point B and ends at point A.
 	The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g.
 from a storefront to the sidewalk at the edge of a scene.
-On this dataset, other distance measures \Ram{are these distance functions that utilize the entire trajectory?} from \cite{Morris2009} and \cite{Lee2007} didn't \Ram{avoid contractions} identify coherent \Ram{what does coherent mean?} motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering \Ram{what does hinder clustering mean?
-Please be more explicit}.
-}
+On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did.
+While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.}
 
@@
 
-\Ram{I don't think we need this addition to the text.}  
-We also changed the language below to better reflect that the specific implementation we chose is an example of how to use the larger framework.
-
-\reversemarginpar\marginnote{\#4.1}
-\response{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
 \end{item}
@@
 In addition, we compared our approach against S-LSTM in addition to the method proposed in \cite{Kitani2012}. 
-The figures corresponding to the predictions of the S-LSTM model are omitted from the paper because as happened for many of the agents, the S-LSTM predictions quickly exit the bounds of the scene, and so the visualization is identically zero \Ram{I don't like this.
-We should include these predictions somehow anyway.}.
-We changed the subtitle of the figure to reflect this, which is shown below.
+We included the visualizations in Figures \ref{bookstore} and \ref{death} below, also which are also updated in the paper.
 
 \reversemarginpar\marginnote{\#4.8}
-\response{An illustration of the predictions generated by the various algorithms.
+
+\begin{figure}
+	\centering
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Our Algorithm}   }
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Kitani et al.}  }
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{S-LSTM} }  
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=100.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=300.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=360.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Random Walk} }  
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.2cm}
+	\hspace{.5cm}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 1.66s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 4.33s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 12.33s$
+	\end{minipage}
+
+	\caption{
+		An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm from \cite{Alahi2016} quickly passes beyond the boundaries of the scene, so the plots are ommitted here.
-}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
+The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+	\label{bookstore}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Our Algorithm}}   
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Kitani et al.}} 
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{S-LSTM}} 
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=100.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=300.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=380.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{Random Walk} }
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.jpg}
+	\end{minipage}
+
+	\vspace{0.2cm}
+	\hspace{.5cm}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 2.33s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 8.33s$
+	\end{minipage}
+	\begin{minipage}{0.3\linewidth}
+		\centering
+		$t = 12.33s$
+	\end{minipage}
+
+	\caption{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+	\label{death}
+\end{figure}
 
@@
 As it is, this is as an immediate avenue for future work which we describe in the following addition to the text: 
-\response{We also hope to do transfer learning \Ram{I flipped the previous two words.} with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+\response{We also hope to do transfer learning with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
 It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}

### 2017-05-08 paper/pedestrian_forecasting_iros_2017.tex
Home stretch

@@
 %%% first author
-\author{Henry O. Jacobs, Owen Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
+\author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
 \thanks{This work was supported by Ford Motor Company.}
@@
 The likelihood of detection is depicted using the \textit{virdis} color palette.
-In this example which was captured at thirty frames per second, the presented algorithm took 0.0158s per frame in a Python implementation.
+In this example which was captured at thirty frames per second, the presented algorithm took 0.00465s per frame in a Python implementation.
 }
@@
 	Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. 
-Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA).
-However LTA suffers from not planning for other pedestrian’s positions at future times.
+	Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporates collision avoidance through an interaction potential.
+	However this method suffers from not planning for other pedestrian' positions at future times. 
 	\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
@@
 classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
-These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data.
+	These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
 	\cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
 	\cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. 
-\cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact.
-\cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}.
-Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.}
+	\cite{Alahi2016} uses Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents will interact and is the state-of-the-art algorithm for socially based models while also having a quick run time.}
 
@@
 \section{Implementation and Experimental results} \label{sec:implementation}
-\marginnote{\#4.1}
-  \rtext{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
+  Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.
   In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
@@
   \marginnote{\#4.2}
-\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$.
-This function measures the distance between the endpoints irrespective of their ordering.
+\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
+	We define the points $\mathbf{a}_1 = (x_{A, \mathrm{start}}, y_{A, \mathrm{start}}, x_{A, \mathrm{end}}, y_{A, \mathrm{end}})$, $\mathbf{a}_2 = (x_{A, \mathrm{end}}, y_{A, \mathrm{end}}, x_{A, \mathrm{start}}, y_{A, \mathrm{start}})$, and $\mathbf{b} = (x_{B, \mathrm{start}}, y_{B, \mathrm{start}}, x_{B, \mathrm{end}}, y_{B, \mathrm{end}})$.
+	We then define our distance measure as $d(A, B) :=  \min \left\{ d_e(\mathbf{a}_1, \mathbf{b}), d_e(\mathbf{a}_2, \mathbf{b}) \right\}$, for the euclidean distance $d_e$ in $\mathbb{R}^4$. 
+	This function measures the distance between the endpoints irrespective of their ordering, which means that a trajectory that starts from point A and ends at point B will be close to a trajectory that starts from point B and ends at point A.
 	The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g.
 from a storefront to the sidewalk at the edge of a scene.
-On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models.
-This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.}
+On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did.
+While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.}
   This clustering of the end-points induces a clustering of the trajectories.
@@
 		\begin{minipage}[c]{0.5cm}
+			\rotatebox{90}{\small{S-LSTM} }  
+		\end{minipage}
+		\begin{minipage}[c]{0.3\linewidth}
+			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=100.jpg}
+		\end{minipage}
+		\begin{minipage}[c]{0.3\linewidth}
+			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=300.jpg}
+		\end{minipage}
+		\begin{minipage}[c]{0.3\linewidth}
+			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=360.jpg}
+		\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
 		\rotatebox{90}{\small{Random Walk} }  
@@
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
+The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
 	\marginnote{\#4.7}
@@
 	\begin{minipage}[c]{0.5cm}
+		\rotatebox{90}{\small{S-LSTM}} 
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=100.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=300.jpg}
+	\end{minipage}
+	\begin{minipage}[c]{0.3\linewidth}
+		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=380.jpg}
+	\end{minipage}
+
+	\vspace{0.1cm}
+	\begin{minipage}[c]{0.5cm}
 		\rotatebox{90}{\small{Random Walk} }
@@
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
+The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
 	\reversemarginpar\marginnote{\#4.8}
@@
 Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
-\reversemarginpar\marginnote{\#4.5} \rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction.
-We used the area under the curve of the ROC curve as our measure of this.
-The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety.
-The area under this curve is a standard measure of the quality of a predictor.
-Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
-In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are.
-Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.}
+\reversemarginpar\marginnote{\#4.5} \rtext{In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
+	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
+	ROC curves which plot the true positive rate (i.e.
+the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e.
+the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly.
+	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
+	The closer that this AUC is to one, the better the prediction algorithm.
+	Figure \ref{ref:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. 
+	In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+Figure \ref{ref:mhd_vs_time} shows MHD plotted against time.
+	Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
+	Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
+	If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
+	Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+	Whereas the MHD would describe this generated set of predictions as informative.}
 Our predictor behaves better than any of the other compared methods at moderately large $t$.
@@
 We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
-\reversemarginpar\marginnote{\#5.1}\rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.
+\reversemarginpar\marginnote{\#5.1}\rtext{Though including social forces into the model is outside of the scope of this paper, its incorporation could begin as follows.
 	In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
 	Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
-Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \reversemarginpar\marginnote{\#5.2}\rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
+	Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995}. 
+	The algorithms for generating real-time prediction would then generalize to this new vector-field definition.} \reversemarginpar\marginnote{\#5.2}\rtext{We also hope to do transfer learning with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
 It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}

### 2017-05-08 paper/response.tex

@@
 On the other hand the Modified Hausdorff Distance (MHD) from \cite{Dubuisson1994} measures the average geometric distance between the ground-truth observed trajectory and the generated set of predictions.
-Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
-Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
-If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
-Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the best way to evaluate the utility of an algorithm in the autonomous vehicle context.
+Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory, but is close to the ground-truth trajectory geometrically. 
+If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space which the person occupied. 
+Notice that in this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
 Whereas the MHD would describe this generated set of predictions as informative.
@@
 Nevertheless, we addressed this concern by adding a comparison to the MHD which is summarized in the Figure \ref{MHD} below and included in the resubmitted version of our manuscript. 
-We outperform \cite{Kitani2012} and \cite{Alahi2016} in the intermediate term, but at short times \cite{Alahi2016} puts more weight on the direction of the observed trajectory, and so their narrow distribution is closer to the trajectory.
-At large scales, \cite{Kitani2012} outperforms us because they know the endpoint of the trajectory.
-If you look at Figure \ref{bookstore} and \ref{death}, you'll see that while they are geometrically closer to the observed trajectory, they don't include it.
+Our approach outperforms the methods proposed in \cite{Kitani2012} and \cite{Alahi2016} in the intermediate term. 
+At short times, the method described in \cite{Alahi2016} puts more weight on the direction of the observed trajectory which results in a narrow distribution that is closer to the trajectory. 
+If you look at Figures \ref{bookstore} and \ref{death}, one can observe that the predictions generated using the method in \cite{Alahi2016} is geometrically closer to the observed trajectory but does not contain it which makes it less appropriate in the autonomous vehicle context.
+At large scales,  the approached described in\cite{Kitani2012} outperforms all  others since it is given the endpoint of the trajectory. 
 %\Ram{it would be useful if we summarized what we were seeing the MHD plot here.
@@
 \response{ In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
-In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
+In this instance, it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
 ROC curves which plot the true positive rate (i.e.
@@
 Figure \ref{AUC} shows the analysis of the AUC of each algorithm versus time.
+
 In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
@@
 Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
-If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
-Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space which the person occupied. 
+Notice that in this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
 Whereas the MHD would describe this generated set of predictions as informative.
@@
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model $100$ times in order to organically extract a less concentrated probability distribution from their algorithm, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times to organically extract a less concentrated probability distribution from their algorithm, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
 	\reversemarginpar\marginnote{\#4.9}
@@
 Thank you for this criticism. 
-We hope we adequately addressed it by strengthening the description of our method for learning the motion model, as well as quantifying the way that we tested our data as is described in our responses to the suggestions by other reviewers.
+We strengthened the description of our method for learning the motion model and quantifying the way that we tested our data as is described in our responses to the suggestions by other reviewers.
 
@@
 Thank you for the suggestion. 
-We hope that we have addressed it by including a review of literature of social motivated models, including S-LSTM, as well as a section on how social factors can be integrated into our model. 
+We hope that we have addressed this criticism by including a review of literature of social motivated models, including S-LSTM, as well as a section on how social factors can be integrated into our model.
 In addition, we compared our approach against S-LSTM in addition to the method proposed in \cite{Kitani2012}. 
-We included the visualizations in Figures \ref{bookstore} and \ref{death} below, also which are also updated in the paper.
+We included the visualizations in Figures \ref{bookstore} and \ref{death}, which are also included in our resubmission.
 
@@
 	\end{minipage}	
-
 	\caption{
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
 	\label{bookstore}
@@
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
 	\label{death}

### 2017-05-08 paper/response.tex
fixed numbering in response letter and got the size of the text under 9 pages

@@
 \usepackage{marginnote}
+\usepackage{subcaption}
 
@@
 The comments provided were incredibly helpful and insightful, and have strengthened the paper considerably.
-Based on the recommendations of the reviewers and the editor, we have made several modifications to the paper.
+Based on the recommendations of the reviewers and the editor, we have made several modifications to the paper including performing additional analyses with an lstm-based approach and using the Modified Hausdorff Distance metric.
+These additions pushed the paper beyond the eight page limit. 
+As a result we removed all proofs from the resubmitted manuscript, but have included them as instead in a supplementary technical report with our resubmission.
+This technical report will be posted to our websites if the paper is accepted. 
 %The changes made are presented below as a highlighted manuscript that indicates additions in green and removals in red.
@@
 At short times, the method described in \cite{Alahi2016} puts more weight on the direction of the observed trajectory which results in a narrow distribution that is closer to the trajectory. 
-If you look at Figures \ref{bookstore} and \ref{death}, one can observe that the predictions generated using the method in \cite{Alahi2016} is geometrically closer to the observed trajectory but does not contain it which makes it less appropriate in the autonomous vehicle context.
+If you look at Figures \ref{fig:bookstore} and \ref{fig:death}, one can observe that the predictions generated using the method in \cite{Alahi2016} are geometrically closer to the observed trajectory but do not contain it.
+As a result these predictions are less appropriate in the autonomous vehicle context.
 At large scales,  the approached described in\cite{Kitani2012} outperforms all  others since it is given the endpoint of the trajectory. 
@@
 
-\reversemarginpar\marginnote{\#4.5}
+\reversemarginpar\marginnote{\#4.4}
 \response{ In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
-In this instance, it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
+	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
 	ROC curves which plot the true positive rate (i.e.
@@
 	Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
-If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space which the person occupied. 
-Notice that in this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+	If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
+	Notice that the true positive rate would be zero meaning that the generated set of predictions was not useful.
 	Whereas the MHD would describe this generated set of predictions as informative.
 }
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.6\textwidth]{figures/the_results.eps}
+	\caption{A comparison of the AUC of the various algorithms.
+Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times to organically extract a less concentrated probability distribution from their algorithm, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
+	\reversemarginpar\marginnote{\#4.7}
+	\label{AUC}
+\end{figure}
 
@@
 \cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms which assume that the end point of the trajectory is unknown.}
-   \reversemarginpar\marginnote{\#4.10}
+   \reversemarginpar\marginnote{\#4.8}
 	\label{MHD}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.6\textwidth]{figures/the_results.eps}
-	\caption{A comparison of the AUC of the various algorithms.
-Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model $100$ times to organically extract a less concentrated probability distribution from their algorithm, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.}
-	\reversemarginpar\marginnote{\#4.9}
-	\label{AUC}
 \end{figure}
@@
 
-\response{Though including social forces into the model is outside of the scope of this paper, its incorporation could begin as follows.  
-In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
-Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
+\reversemarginpar\marginnote{\#5.1}
+\response{Social forces can be incorporated into this model by
+	%In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
+	including a social force $F_{i}$ acting on the $i$th agent by letting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
 	Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995}. 
-The algorithms for generating real-time prediction would then generalize to this new vector-field definition.
+	The algorithms for generating real-time predictions would then generalize to this definition.
 }
@@
 
-\reversemarginpar\marginnote{\#4.3}
+\reversemarginpar\marginnote{\#4.2}
 \response{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
@@
 
-\reversemarginpar\marginnote{\#4.2}
-\response{
-	We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+\reversemarginpar\marginnote{\#4.1}
+\response{To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
 	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
@@
 \reversemarginpar\marginnote{\#1.1}
-	\response{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians. 
-	Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. 
-	Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporates collision avoidance through an interaction potential.
-	However this method suffers from not planning for other pedestrian'€ positions at future times. 
-	\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
-	 \cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
-classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
+	\response{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}.
+	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
+	However this method suffers from not planning for other pedestrian positions at future times. 
+	Others take optical flow as input \cite{Hospedales2009,Wang2009,Emonet2011}.
+	These approaches use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), or on a Markov model. 
 	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
-	 \cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
-	 \cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. 
-	 \cite{Alahi2016} uses Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents will interact and is the state-of-the-art algorithm for socially based models while also having a quick run time.
-}
+	 Recently methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene \cite{Koppula2016}.
+	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
+	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact and is the state-of-the-art algorithm for socially based models while also having a quick run time \cite{Alahi2016}.}
 
@@
 In addition, we compared our approach against S-LSTM in addition to the method proposed in \cite{Kitani2012}. 
-We included the visualizations in Figures \ref{bookstore} and \ref{death}, which are also included in our resubmission.
+We included the visualizations in Figures \ref{fig:bookstore} and \ref{fig:death}, which are also included in our resubmission.
 
-\reversemarginpar\marginnote{\#4.8}
-
-\begin{figure}
+ \begin{figure}[t!]
+  \begin{subfigure}[t]{0.5\textwidth}
 	\centering
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Our Algorithm}   }
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Kitani et al.}  }
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+		\begin{minipage}[c]{0.45cm}
 			\rotatebox{90}{\small{S-LSTM} }  
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Random Walk} }  
@@
 	\hspace{.5cm}
-	\begin{minipage}{0.3\linewidth}
+	\begin{minipage}{0.28\linewidth}
 		\centering
@@
 	\end{minipage}
-	\begin{minipage}{0.3\linewidth}
+	\begin{minipage}{0.28\linewidth}
 		\centering
@@
 	\end{minipage}
-	\caption{
-		An illustration of the predictions generated by the various algorithms.
+	\caption{Scene 1}
+%	\caption{
+%		\rtext{An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
-	\label{bookstore}
-\end{figure}
-
-\begin{figure}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
+%	\marginnote{\#4.7}
+	\label{fig:bookstore}
+	\end{subfigure} 
+	\hspace*{0.5cm}
+	\begin{subfigure}[t]{0.5\textwidth}
 	\centering
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Our Algorithm}}   
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Kitani et al.}} 
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{S-LSTM}} 
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.45cm}
 		\rotatebox{90}{\small{Random Walk} }
@@
 	\end{minipage}	
-
+	\caption{Scene 2}
+%	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
+%	\reversemarginpar\marginnote{\#4.8}
+	\label{fig:death}
+	\end{subfigure}
 	\caption{An illustration of the predictions generated by the various algorithms.
@@
 The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
-	\label{death}
+	\reversemarginpar \marginnote{\#4.5}
+%	\vspace{-0.7cm}
 \end{figure}
@@
 As it is, this is as an immediate avenue for future work which we describe in the following addition to the text: 
-\response{We also hope to do transfer learning with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
-It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
+\reversemarginpar\marginnote{\#5.2}
+\response{We also plan to explore transfer learning using scene segmentation, as well as the semantic context descriptors and routing scores to show how vector fields can be transferred to novel scenes.
+It appears that the low-order parameterization and unit-length vector field of our model, make it amenable to the methods developed in \cite{Ballan2016}.}
 
@@
 
-\reversemarginpar\marginnote{\#4.4}
+\reversemarginpar\marginnote{\#4.3}
 \response{ Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}

### 2017-05-08 paper/pedestrian_forecasting_iros_2017.tex
fixed numbering in response letter and got the size of the text under 9 pages

@@
 \usepackage{color}  % For Highlighting
+%\usepackage{natbib}
+\usepackage[noadjust]{cite}
 
@@
 \author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
+ \thanks{ H.O. Jacobs, O. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
+{\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
 \thanks{This work was supported by Ford Motor Company.}
- \thanks{ H.O. Jacobs, O. Hughes, and R.~Vasudevan are with the Department of Mechanical Engineering, University of Michigan, Ann Arbor, MI 48109
-{\scriptsize \{\texttt{hojacobs,owhughes,ramv}\}~\texttt{@umich.edu}}}
-\thanks{M. Johnson-Roberson is with the Department of Naval and Marine Engineering, University of Michigan, Ann Arbor, MI 48109\newline
-{\scriptsize \texttt{mattjr@umich.edu}}}
 }
@@
 To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories.
-The margins between humans and vehicles are small and both frequently operate at the boundary. 
+%The margins between humans and vehicles are small and both frequently operate at the boundary. 
 Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.
@@
 	\caption{The performance of the presented algorithm captures the most probable routes that a pedestrian chooses.
-In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
+The dot is the starting point of the trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.
 The likelihood of detection is depicted using the \textit{virdis} color palette.
-In this example which was captured at thirty frames per second, the presented algorithm took 0.00465s per frame in a Python implementation.
+The presented algorithm took $0.00465$s per frame in a Python implementation.
 }
 	\label{fig:gates-1-2}
+	\vspace*{-0.7cm}
 \end{figure}
@@
 This paper is focused on the construction of useful motion models for pedestrians that can aide the task of real-time forecasting for autonomous agents.
-The simplest approach to forecasting with motion models forward integrates a Kalman filter \cite{kalman1960new} based upon the observed heading.
+The simplest approach to forecasting with motion models forward integrates a Kalman filter based upon the observed heading.
 Over short time scales this method may perform well, but the resulting distribution devolves into an imprecise Gaussian mass over longer time scales.
-This is a typical property of the extended Kalman filter and the Kalman filter with deterministic nonlinear drift \cite{Schneider2013}.
+%This is a typical property of the extended Kalman filter and the Kalman filter with deterministic nonlinear drift \cite{Schneider2013}.
 In particular, such models are less useful for forecasts beyond two seconds, especially when a pedestrian turns.
 Nonetheless, these stochastic linear models serve as a reasonable default in the absence of any contextual knowledge.
-For example, \cite{Kooji2014} uses a motion model of this variety, parametrized by latent variables, such as the pedestrian's awareness of nearby vehicles, to predict when and how a pedestrian may cross a street without leveraging any environmental context.
-
-More sophisticated models that attempt to leverage environmental data include inverse optimal control (IOC) based models \cite{Ziebart2008,Ziebart2009,Kitani2012,Xie2013,Karasev2016}.
+%For example, \cite{Kooji2014} uses a motion model of this variety, parametrized by latent variables, such as the pedestrian's awareness of nearby vehicles, to predict when and how a pedestrian may cross a street without leveraging any environmental context.
+
+More sophisticated models that attempt to leverage environmental data include Inverse Optimal Control (IOC) based models \cite{Ziebart2008,Ziebart2009,Kitani2012,Xie2013,Karasev2016}.
 These IOC models have the desirable property of attributing intention and goal-seeking behavior to the agents.
@@
 
-The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
+%The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
+This approach has been generalized in a variety of ways.
 For example, time-dependent information, such as traffic signals, are incorporated in \cite{Karasev2016}, by relaxing the Markov property and considering a switched Markov process.
-Other generalizations in \cite{Karasev2016} include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
-Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method: a Rao-Blackwellized particle filter~\cite{Doucet2000} is employed.
-The resulting accuracy of such methods can only be known in a probabilistic sense in that the error bounds are themselves random variables.
-Moreover, accuracy can come at a large computational expensive which is prohibitive during real-time control.
-
-One limitation of IOC models occurs in cases where there are many locally optimal solutions between a start and end goal.
-In these cases, IOC type methods can yield non-robust and imprecise behavior.
-This can occur, for example, when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
-To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch based on its feature and the behavior of previous agents on similar patches.
-The motion model is a Markov jump process and the relevant posteriors were approximated using sample based techniques similar to \cite{Karasev2016}.
+Other generalizations include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
+Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method.
+The resulting accuracy of such methods, which can come at a large computational expense, can only be known in a probabilistic sense in that the error bounds are themselves random variables.
+%Moreover, accuracy can come at a large computational expense.
+% which is prohibitive during real-time control.
+
+One limitation of IOC models occurs in cases where there are locally optimal solutions between a start and end goals which can yield non-robust and imprecise behavior.
+This can occur when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
+To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch.
+The motion model is a Markov jump process and the relevant posteriors are approximated using sample based techniques similar to \cite{Karasev2016}.
 The objective of \cite{Ballan2016} is not only prediction, but the development of a motion model learned on one scene that could then subsequently be transferred to other scenes. 
 This requires representations of ``objects'' in the scene that do not depend rigidly on the finitely many labels an engineer managed to think of in a late-night brainstorming session.
-For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.
-
-Along similar lines to \cite{Ballan2016}, \cite{Walker2014} constructed an unsupervised approach towards forecasting.
-As before, the motion model was a Markov jump process, and the training set was a collection unlabeled videos.
+%For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.
+
+Recent work has focused on constructing an unsupervised approach towards forecasting  \cite{Walker2014}.
+%As before, the motion model was a Markov jump process, and the training set was a collection unlabeled videos.
 Unlike all the approaches mentioned thus far, the agents in \cite{Walker2014} were not manually specified.
 They were learned by detecting which sort of patches of video were likely to move, and how.
-The resulting predictions outperformed \cite{Kitani2012} when comparing the most likely path with the ground truth using the mean Hausdorff distance.
-As in all methods mentioned thus far, computational speed and accuracy of any predicted posteriors were not a concern of \cite{Walker2014}, so no such results were reported.
-However, since the motion model was a Markov jump process which required the application of a sample based technique, we should expect the same painful trade-off between error and speed to occur as in \cite{Karasev2016} and \cite{Ballan2016}.
-
+The resulting predictions outperformed \cite{Kitani2012} when comparing the most likely path with the ground truth using the Modified Hausdorff Distance.
+As in all methods mentioned thus far, computational speed and accuracy of any predicted posteriors were not a concern, so no such results were reported.
+However, since the motion model was a Markov jump process which required the application of a sample based technique, we should expect the same painful trade-off between error and speed to occur as in \cite{Karasev2016,Ballan2016}.
+
+\reversemarginpar
 \marginnote{\#1.1}
-\rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians. 
-	Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. 
-	Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporates collision avoidance through an interaction potential.
-	However this method suffers from not planning for other pedestrian' positions at future times. 
-	\cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input.
-	\cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e.
-classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model.
+\rtext{Many have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
+	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
+	However this method suffers from not planning for other pedestrian positions at future times. 
+	Others take optical flow as input \cite{Hospedales2009,Wang2009,Emonet2011}.
+	These approaches use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), or on a Markov model. 
 	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
-	\cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene.
-	\cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. 
-	\cite{Alahi2016} uses Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents will interact and is the state-of-the-art algorithm for socially based models while also having a quick run time.}
+	 Recently methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
+	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
+	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.}
 
@@
 
-The primary contributions of this paper are:
-\begin{itemize}
-	\item An accurate motion model for pedestrian forecasting.
-	\item An expedient method for computing approximations of relevant posteriors generated by our motion model.
-	\item Hard error bounds on the proposed approximations.
-\end{itemize}
+The primary contributions of this paper are three-fold: first, an accurate motion model for pedestrian forecasting, second, an expedient method for computing approximations of relevant posteriors generated by our motion model, and finally hard error bounds on the proposed approximations.
 The method proposed by this paper is able to work three times faster than the existing state of the art while improving upon its performance over long time horizons.
-
 For clarification, we should mention that there are a number of things that we do not do.
 For example, we do not concern ourselves with detection and tracking. 
-Nor do we concern ourselves with updating our prediction as new data comes along \cite{Kitani2012}.
+Nor do we concern ourselves with updating our prediction as new data comes along.
 We largely work in a 2D environment with a bird's eye view, operating under the assumption that the data has been appropriately transformed by a third party.
-This is in contrast to \cite{Karasev2016}, which operates from first person video.
-While it would be a straight forward extension to consider such things, it would detract from the presentation.
+%This is in contrast to \cite{Karasev2016}, which operates from first person video.
+While it would be a straight forward extension to consider a first person perspective, it would detract from the presentation.
 %Such questions are important, but beyond the scope of this paper, which concerns the underlying dynamic model.
 
-The rest of the paper is organized as follows:
-\begin{itemize}
-	\item \S \ref{sec:model} describes our motion model as a Bayesian network.
-	\item \S \ref{sec:efficient} describes how to compute probability densities for an agent's position efficiently.
-	\item \S \ref{sec:implementation} demonstrates the model by training and testing it on the Stanford drone dataset \cite{Robicquet2016}.
-\end{itemize}
+The rest of the paper is organized as follows: \S \ref{sec:model} describes our motion model as a Bayesian network, \S \ref{sec:efficient} describes how to compute probability densities for an agent's position efficiently, and \S \ref{sec:implementation} demonstrates the model by training and testing it on the Stanford drone dataset \cite{Robicquet2016}.
 
@@
 
-To numerically compute $\rho_t$, we build a probabilistic graphical model.
-Our model assumes we have noisy information about agents, and each agent moves with some intention through the world in a way that is roughly approximated by a model.
-Our model can be divided into three parts:
+To compute $\rho_t$, we build a probabilistic graphical model, which is composed of three parts:
+%Our model assumes we have noisy information about agents, and each agent moves with some intention through the world.
+%Our model is divided into three parts:
 \begin{enumerate}
@@
 	\item Measurements:  This is represented by our sensor readings $\hat{x}_0$ and $\hat{v}_0$ and are independent of all other variables given the true initial position and velocity, $x_o, v_0$.
-	\item Motion Model: This is represented by a trajectory $\check{x}_t$ and depends on a variety of other variables which are described below.
+	\item Motion Model: This is represented by a trajectory $\check{x}_t$ and depends on a variety of other variables.
 \end{enumerate}
-We now elaborate on these three components, and relate them to one another.
-Many of the choices we make are motivated by a balance between model quality and computational speed (see \S \ref{sec:efficient}).
+We elaborate on these three components next.
+%Many of the choices are motivated by a balance between model quality and computational speed.
 
@@
 \end{align}
-where $X_k$ is a vector-field\footnote{A vector-field, generally speaking, is an assignment of a velocity to each position in some space.
-A vector-field on $\mathbb{R}^n$ is nothing but a map from $\mathbb{R}^n \to \mathbb{R}^n$.} drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
+where $X_k$ is a vector-field\footnote{A vector-field is an assignment of a velocity to each position in some space.
+A vector-field on $\mathbb{R}^n$ is a map from $\mathbb{R}^n \to \mathbb{R}^n$.} drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
 More specifically, we assume that each $X_k$ has the property that $\| X_k(x) \| = 1$ for all $x \in D$.
@@
 where $\Phi^{t}_{k,s}$ is the flow-map of the vector field $s \,X_k$ up to time $t$.
-Note that this flow-map can be evaluated for an initial condition efficiently by just integrating the vector field from that initial condition.
+Note that this flow-map can be evaluated for an initial condition by just integrating the vector field from that initial condition.
 Note the variables $k,s$ and $x_0$ determine $v_0$.
@@
 \end{align}
-In summary, the agent models are parametrized by the set:
-\begin{align}
-	\mathcal{M} = \{ lin \} \cup \left( \mathbb{R} \times \{ 1 , \dots, n \} \right) \label{eq:models}
-\end{align}
-and each flavor determines the type of motion we should expect from the agent model.
+In summary, the agent models are parametrized by the set $\mathcal{M} = \{ lin \} \cup \left( \mathbb{R} \times \{ 1 , \dots, n \} \right)$ whose elements determine the type of agent motion.
 
@@
 \begin{align}
-\begin{tikzpicture}[thick, var/.style={circle,draw,thin,rounded corners,shade,top color=blue!50,minimum size = 4mm}]
+\begin{tikzpicture}[thick, var/.style={circle,draw,thin,rounded corners,shade,top color=blue!50,minimum size = 2mm}]
 	\node[var] (M) {$M$};
@@
 \end{align}
-We use this Bayesian network to compute $\rho_t$ efficiently.
+We use this Bayesian network to compute $\rho_t$.
 In particular
 \begin{align}
-	\rho_t(x_t ) &:= \Pr( x_t \mid \hat{x}_0, \hat{v}_0 ) \\
-	&= \left( \sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds \right)  \nonumber \\
-	&\quad + \Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 ). \label{eq:decomposition}
-\end{align}
-The final term $\Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 )$ is expressible in closed form when the posteriors $\Pr( x_0 \mid lin)$ and $\Pr( v_0 \mid lin,x_0)$ have known expressions ( e.g.
-Gaussians or uniform distributions).
-In this instance, the numerical computation of $\Pr(x_t, lin \mid \hat{x}_0, \hat{v}_0)$ poses a negligible burden.
-Instead the primary computational burden derives from computing $\sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds$.
+	&\rho_t(x_t ) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 ) \\
+	&=\sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds  + \Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 ). \label{eq:decomposition}
+\end{align}
+$\Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 )$ is expressible in closed form when the posteriors $\Pr( x_0 \mid lin)$ and $\Pr( v_0 \mid lin,x_0)$ are known. 
+In this instance, the numerical computation of $\Pr(x_t, lin \mid \hat{x}_0, \hat{v}_0)$ poses a negligible burden and the primary computational burden derives from computing $\sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds$.
 
 \section{Efficient Probability Propagation} \label{sec:efficient}
-As mentioned, many of the modeling choices are born out of a balance between accuracy and real-time computability.
-One of the major modeling choices, that agents move approximately according to a small number of ODEs, is the most prominent such choice.
-This section details how this modeling choice can be leveraged to compute $\rho_t(x_t)$ quickly and accurately.
-
+%As mentioned, many of the modeling choices are born out of a balance between accuracy and real-time computability.
+%One of the major modeling choices, that agents move approximately according to a small number of ODEs, is the most prominent such choice.
+This section details how the modeling of the agent's motion as satisfying an ODE can be leveraged to compute $\rho_t(x_t)$ quickly and accurately.
 To begin, rather than focusing on computing $\rho_t(x_t)$, we describe how to compute the joint probability $\Pr( x_t , \hat{x}_0, \hat{v}_0)$.
@@
 We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
+
 Let us assume that $\Pr(s \mid k)$ is compactly supported for all $k = 1,\dots,n$, and the supported is always contained
 in some interval $[ - \bar{s} , \bar{s} ]$ for some $\bar{s} > 0$.
-Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on $[-\bar{s}, \bar{s}]$, 
-we can conclude that the integral term in \eqref{eq:decomposition} can be approximated by
+Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on $[-\bar{s}, \bar{s}]$,  we can conclude that the integral term in \eqref{eq:decomposition} is approximated by
 \begin{align}
@@
 \begin{align}
-	&\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0) =\\
-	&= \int \Pr( x_t, \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \\
-	&= \int \Pr( x_t \mid \check{x}_t ) \Pr(\check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \label{eq:convolve}
+	\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0) &= \int \Pr( x_t, \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t  \label{eq:convolve} \\
+	&= \int \Pr( x_t \mid \check{x}_t ) \Pr(\check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \nonumber
 \end{align}
@@
 \begin{align}
-	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \\
-	&= \int \Pr( \check{x}_t , x_0, \hat{x}_0, v_0, \hat{v}_0, k,s) dx_0 \, dv_0 \\
+	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) =\int \Pr( \check{x}_t , x_0, \hat{x}_0, v_0, \hat{v}_0, k,s) dx_0 \, dv_0 \nonumber \\
 	&= \int \Pr( \check{x}_t \mid  x_0, k,s, v_0) \Pr( \hat{v}_0 \mid v_0) \cdot \\
@@
 \begin{align}
-	\begin{split}
-		&\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 ) = \\
-		&\qquad \underbrace{\left( \sum_{i,j=-N}^{N} W(k,s,i,j,\hat{x}_0) \delta( x_0 - x_0^{i,j} ) \right)}_{\text{approximation}}
+		&\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 ) = \nonumber \\
+		& \underbrace{\left( \sum_{i,j=-N}^{N} W(k,s,i,j,\hat{x}_0) \delta( x_0 - x_0^{i,j} ) \right)}_{\text{approximation}}
 		+ \underbrace{ \varepsilon_0(x_0) }_{\text{error}}
-	\end{split}
 	\label{eq:approximation 1}
@@
 \end{thm}
-\begin{proof}
-	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
-	We observe
-\begin{align*}
-	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
-	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
-	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
-\end{align*}
-As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
-\end{proof}
+%\begin{proof}
+%	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
+%	We observe
+%\begin{align*}
+%	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
+%	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
+%	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
+%\end{align*}
+%As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
+%\end{proof}
 
@@
 \end{cor}
-\begin{proof}
-	By \eqref{eq:convolve}
-	\begin{align*}
-		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
-	\end{align*}
-	Substitution of \eqref{eq:approximation 1} yields
-	\begin{align*}
-		\begin{split}
-		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
-		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
-		\end{split}
-	\end{align*}
-	where the error term is
-	\begin{align}
-		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
-	\end{align}
-	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
-	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
-	\begin{align}
-		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
-			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
-	\end{align}
-	Implementing the integration over $x_t$ first yields:
-	\begin{align}
-		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
-	\end{align}
-	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
-\end{proof}
+%\begin{proof}
+%	By \eqref{eq:convolve}
+%	\begin{align*}
+%		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
+%	\end{align*}
+%	Substitution of \eqref{eq:approximation 1} yields
+%	\begin{align*}
+%		\begin{split}
+%		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
+%		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
+%		\end{split}
+%	\end{align*}
+%	where the error term is
+%	\begin{align}
+%		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
+%	\end{align}
+%	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
+%	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
+%	\begin{align}
+%		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
+%			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
+%	\end{align}
+%	Implementing the integration over $x_t$ first yields:
+%	\begin{align}
+%		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
+%	\end{align}
+%	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
+%\end{proof}
 
 Corollary \ref{cor:error} justifies using \eqref{eq:approximation 2} as an approximation of $\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0)$.
-In particular, this reduces the problem of computing $\rho_t(x_t)$ to the problem of computing the weights $W(k,s,i,j,\hat{x}_0)$ and the points $\Phi_{k,s}^t(x_0^{i,j})$ for all $k,s$ and points $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
+This reduces the problem of computing $\rho_t(x_t)$ to the problem of computing the weights $W(k,s,i,j,\hat{x}_0)$ and the points $\Phi_{k,s}^t(x_0^{i,j})$ for all $k,s$ and points $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
 We can reduce this burden further by exploiting the following symmetry:
@@
 \end{thm}
-\begin{proof}
-	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
-	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
-	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
-	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
-	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
-	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
-\end{proof}
+%\begin{proof}
+%	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
+%	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
+%	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
+%	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
+%	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
+%	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
+%\end{proof}
 This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$, which yields the following result:
@@
 \end{thm}
-\begin{proof}
-	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
-	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
-	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
-\end{proof}
+%\begin{proof}
+%	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
+%	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
+%	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
+%\end{proof}
 
@@
 given by
-\begin{align*}
-	\begin{split}
-		&\frac{\bar{s}}{\ell} \sum_{i,j,k,m} W(k,m \bar{s} / \ell ,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{m \bar{s} \Delta t}( x_0^{i,j}) } \\
+\begin{align}
+		&\frac{\bar{s}}{\ell} \sum_{i,j,k,m} W(k,m \bar{s} / \ell ,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{m \bar{s} \Delta t}( x_0^{i,j}) }  \nonumber \\
 		&\qquad+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 ).
-	\end{split}
-\end{align*}
+\end{align}
 As we have already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
-Now if we want to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at $t=(\ell + 1 ) \Delta t$, then by reusing the earlier computation one would only need to compute $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for $m \in \{ -( \ell + 1 ), (\ell + 1 )\}$ along with the weights to obtain an approximation using the partition $\mathcal{P}_{\ell + 1}$.
+If we want to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ at $t=(\ell + 1 ) \Delta t$, then by reusing the earlier computation one only need to compute $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for $m \in \{ -( \ell + 1 ), (\ell + 1 )\}$ along with the weights to obtain an approximation using the partition $\mathcal{P}_{\ell + 1}$.
 %Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
+
+%\begin{algorithm}[t]
+%		    \KwData{ $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.}
+%		    \For{$\ell \in \{ -N_t, \dots, N_t\}$}{
+%		    	Compute $\Phi_{k,1}^{\bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0,N)$ \;
+%			\For{$m \in \{ -\ell,\dots,\ell \}$}{
+%				Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ \;
+%			}
+%			Use Theorem \ref{thm:main} to obtain a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$\;
+%			Normalize with respect to $\check{x}_t$ to obtain an equally accurate approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$ \;
+%			Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$.
+%		    }
+%		    \KwResult{An approximation of $\rho_t(x_t)$ at $t= 0,\Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
+%		    \caption{Our algorithm for computing $\rho_t$}
+%		    		    \label{alg:main}
+%\end{algorithm}
+
+%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
+%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
+The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.
+For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t \Delta t \bar{s} ,\dots, N_t \Delta t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
+Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
+If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
+%We summarize this in the following theorem:
+%\begin{thm}
+%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
+%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
+%\end{thm}
 
@@
 
-%\begin{algorithm}[t]
-%		    \KwData{ $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.}
-%		    \For{$\ell \in \{ -N_t, \dots, N_t\}$}{
-%		    	Compute $\Phi_{k,1}^{\bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0,N)$ \;
-%			\For{$m \in \{ -\ell,\dots,\ell \}$}{
-%				Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ \;
-%			}
-%			Use Theorem \ref{thm:main} to obtain a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$\;
-%			Normalize with respect to $\check{x}_t$ to obtain an equally accurate approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$ \;
-%			Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$.
-%		    }
-%		    \KwResult{An approximation of $\rho_t(x_t)$ at $t= 0,\Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
-%		    \caption{Our algorithm for computing $\rho_t$}
-%		    		    \label{alg:main}
-%\end{algorithm}
-
-%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
-%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
-The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.
-For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t \Delta t \bar{s} ,\dots, N_t \Delta t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
-Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
-If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
-%We summarize this in the following theorem:
-%\begin{thm}
-%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
-%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
-%\end{thm}
-
 \section{Implementation and Experimental results} \label{sec:implementation}
@@
  We begin by identifying the number of possible vector-fields.
-  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
   %In principal, any clustering algorithm could accomplish this.
-  \marginnote{\#4.2}
-\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+  \normalmarginpar
+\marginnote{\#4.1}
+\rtext{To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
 	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
@@
   Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
-  We assume $\Theta_k$ takes the form:
-  \begin{align*}
-  	\Theta_k(x) = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x),
-  \end{align*}
-  for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
-  In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
+  We assume $\Theta_k = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x)$ for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
+  We choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
+
  $\Theta_k$ is learned by computing the velocities observed in the cluster, $S_k$.
@@
   We learn $\Theta_k$ by defining the cost-function:
-  \begin{align*}
+  \begin{align}
   	C[ \Theta_k] = \sum_i \langle v_{i,k} , ( \cos(\Theta_k( x_{i,k}) , \sin( \Theta_k( x_{i,k} ) ) \rangle
-  \end{align*}
+  \end{align}
   which penalizes $\Theta_k$ for producing a misalignment with the observed velocities at the observed points of $S_k$.
@@
   This means that we only need to learn $\Pr( x_0 \mid k)$, $\Pr(k)$, and $\Pr(s)$.
-
   We let $\Pr(k) = (n+1)^{-1}$ and $\Pr(s) \sim \mathcal{U}( [-s_{\max}, s_{\max} ] )$  where $s_{\max}>0$ is the largest observed speed in the dataset.  
@@
 %  That is to say, 
-  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is given by:
-  \begin{align*}
-  	V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)
-  \end{align*}
-  for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
+  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is  $V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)$ for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
   We chose our basis functions to be the collection of tensor products of the first six Legendre polynomials, normalized to the size of the domain.
@@
   The resulting (convex) optimization problem takes the form:
-  \begin{align*}
-  	\mathbf{c}^* = \inf_{ |\mathbf{c}| } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
-  \end{align*}
+  \begin{align}
+  	\mathbf{c}^* = \inf_{ \mathbf{c} } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
+  \end{align}
   Where the norm on $\mathbf{c}$ is a sup-norm.
-  We bias this optimization towards more regular functions by adding a penalty to the cost function.
+  We bias this optimization towards smooth functions by adding a penalty to the cost function.
     Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.
@@
   \subsection{Learning the Measurement Model}
-  We assume a Gaussian noise model.  
-  That is
-  \begin{align*}
-  	\Pr( \hat{x}_0 \mid x_0 ) \sim \mathcal{N}( x_0 , \sigma_x) \quad, \quad \Pr( \hat{v}_0 \mid v_0 ) \sim \mathcal{N}( v_0, \sigma_v).
-  \end{align*}
+  We assume a Gaussian noise model (i.e. $\Pr( \hat{x}_0 \mid x_0 ) \sim \mathcal{N}( x_0 , \sigma_x)$ and $\Pr( \hat{v}_0 \mid v_0 ) \sim \mathcal{N}( v_0, \sigma_v)$).
   Therefore, our model is parametrized by the standard deviations $\sigma_x$ and $\sigma_v$.
@@
 
-  \begin{figure}
+  \begin{figure*}[t!]
+  \begin{subfigure}[t]{0.48\textwidth}
 	\centering
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Our Algorithm}   }
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Kitani et al.}  }
@@
 	\vspace{0.1cm}
-		\begin{minipage}[c]{0.5cm}
+		\begin{minipage}[c]{0.47cm}
 			\rotatebox{90}{\small{S-LSTM} }  
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Random Walk} }  
@@
 	\end{minipage}
-
-	\caption{
-		\rtext{An illustration of the predictions generated by the various algorithms.
+	\caption{Scene 1}
+%	\caption{
+%		\rtext{An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent.
-The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
-	\marginnote{\#4.7}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
+%	\marginnote{\#4.7}
 	\label{fig:bookstore-1-2}
-\end{figure}
-
-\begin{figure}
+	\end{subfigure} 
+	\hspace*{0.5cm}
+	\begin{subfigure}[t]{0.48\textwidth}
 	\centering
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Our Algorithm}}   
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Kitani et al.}} 
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{S-LSTM}} 
@@
 	\vspace{0.1cm}
-	\begin{minipage}[c]{0.5cm}
+	\begin{minipage}[c]{0.47cm}
 		\rotatebox{90}{\small{Random Walk} }
@@
 	\end{minipage}	
-
-	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+	\caption{Scene 2}
+%	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
 The likelihood of detection is depicted using the virdis color palette.
-Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in [13] is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
-	\reversemarginpar\marginnote{\#4.8}
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
+%	\reversemarginpar\marginnote{\#4.8}
 	\label{fig:death-1-2}
-\end{figure}
+	\end{subfigure}
+	\caption{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+	\normalmarginpar \marginnote{\#4.5}
+	\vspace{-0.7cm}
+\end{figure*}
+
+%\begin{figure}
+%	\centering
+%	\begin{minipage}[c]{0.5cm}
+%		\rotatebox{90}{\small{Our Algorithm}}   
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.jpg}
+%	\end{minipage}
+%	
+%	\vspace{0.1cm}
+%	\begin{minipage}[c]{0.5cm}
+%		\rotatebox{90}{\small{Kitani et al.}} 
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.jpg}
+%	\end{minipage}
+%	
+%	\vspace{0.1cm}
+%	\begin{minipage}[c]{0.5cm}
+%		\rotatebox{90}{\small{S-LSTM}} 
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=100.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=300.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=380.jpg}
+%	\end{minipage}
+%	
+%	\vspace{0.1cm}
+%	\begin{minipage}[c]{0.5cm}
+%		\rotatebox{90}{\small{Random Walk} }
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.jpg}
+%	\end{minipage}
+%	\begin{minipage}[c]{0.3\linewidth}
+%		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.jpg}
+%	\end{minipage}
+%	
+%	\vspace{0.2cm}
+%	\hspace{.5cm}
+%	\begin{minipage}{0.3\linewidth}
+%		\centering
+%		$t = 2.33s$
+%	\end{minipage}
+%	\begin{minipage}{0.3\linewidth}
+%		\centering
+%		$t = 8.33s$
+%	\end{minipage}
+%	\begin{minipage}{0.3\linewidth}
+%		\centering
+%		$t = 12.33s$
+%	\end{minipage}
+%	
+%	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
+In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
+The likelihood of detection is depicted using the virdis color palette.
+Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
+%	\reversemarginpar\marginnote{\#4.8}
+%	\label{fig:death-1-2}
+%\end{figure}
 
@@
 
-This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
-We implement our model as well as our evaluation code in Python 2.6. 
-Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
-\reversemarginpar\marginnote{\#4.3}
-\rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
+This section establishes our methods performance and compares it to the model from \cite{Kitani2012}, \cite{Alahi2016}, and a random walk.
+We implement our model as well as our evaluation code in Python 2.6\footnote{https://github.com/ramvasudevan/iros2017\_pedestrian\_forecasting}.
+\marginnote{\#4.2}
+\rtext{We performed a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
 	We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
-\reversemarginpar\marginnote{\#4.4}
+\marginnote{\#4.3}
 \rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
@@
 
-The output distributions of the three algorithms were compared using their integrals over the cells of a regular grid over our domain. 
+The output distributions of the four algorithms were compared using their integrals over the cells of a regular grid over our domain. 
 These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
@@
 %were the average bounding box dimensions in the training data set. 
-Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
-\reversemarginpar\marginnote{\#4.5} \rtext{In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
+Because our predictions all share the same scale, we amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
+\marginnote{\#4.4} \rtext{In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
 	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
@@
 	The closer that this AUC is to one, the better the prediction algorithm.
-	Figure \ref{ref:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. 
-	In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
-Figure \ref{ref:mhd_vs_time} shows MHD plotted against time.
+	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. }
+
+	\rtext{In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.
 	Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context.
@@
 	If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
-	Notice that In this instance the true positive rate would be zero meaning that the generated set of predictions was not useful.
+	Notice that the true positive rate would be zero meaning that the generated set of predictions was not useful.
 	Whereas the MHD would describe this generated set of predictions as informative.}
@@
 %Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{./figures/the_results.eps}
-	\caption{\rtext{A comparison of the AUC of the various algorithms.
-Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on ourdated pedestrian measurements at moderate-to-large timescales lowered their performance.}}
-	\normalmarginpar\marginnote{\#4.9}
-	\label{fig:auc_vs_time}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{./figures/MHD_results.eps}
-	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution.
-The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times.
-At long timescales the MHD to the trajectory of most algorithms converges.
-Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}}
-	\normalmarginpar\marginnote{\#4.10}
-	\label{fig:mhd_vs_time}
-\end{figure}
 
@@
 	\end{center}
+	\vspace*{-0.6cm}
 \end{table}
 
-\normalmarginpar\marginnote{\#4.6}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{./figures/the_results.eps}
+	\caption{\rtext{A comparison of the AUC of the various algorithms.
+Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}}
+	\normalmarginpar\marginnote{\#4.7}
+	\label{fig:auc_vs_time}
+	\vspace*{-0.75cm}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{./figures/MHD_results.eps}
+	\caption{\rtext{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
+The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}}
+	\reversemarginpar\marginnote{\#4.8}
+	\label{fig:mhd_vs_time}
+	\vspace*{-0.7cm}
+\end{figure}
+
+\reversemarginpar
+\marginnote{\#4.6}
 \rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
 This is shown in Table~\ref{tab:time}. 
-Our algorithm implementation leveraged its embarrassing parallelism using a relatively naive Python implementation, which split the computation of frames among 18 cores. 
+Our algorithm implementation leveraged its parallelism, which split the computation of frames among 18 cores. 
 %Our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. 
@@
 This paper presents a real-time approach to probabilistic pedestrian modeling. 
-We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in three critical areas: 1) prediction speed is efficient enabling tight loop real-time control;  2) our error bound on the predictions accurately reflects potential paths for the human limiting false negatives and maximizing the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a wide variety of real world applications.
-These three features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces.
-We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
-\reversemarginpar\marginnote{\#5.1}\rtext{Though including social forces into the model is outside of the scope of this paper, its incorporation could begin as follows.
-	In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
-	Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
+We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in two critical areas: 1) prediction speed is efficient to enable real-time control and 2) our error bound on the predictions accurately reflects potential paths for the human.
+These two features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
+
+%We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop. 
+\reversemarginpar\marginnote{\#5.1}\rtext{Social forces can be incorporated into this model by
+	%In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
+	including a social force $F_{i}$ acting on the $i$th agent by letting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  
 	Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995}. 
-	The algorithms for generating real-time prediction would then generalize to this new vector-field definition.} \reversemarginpar\marginnote{\#5.2}\rtext{We also hope to do transfer learning with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes.
-It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
-
-\section*{Acknowledgments}
-Kris Kitani was critical in aiding the comparison to his algorithm.
+	The algorithms for generating real-time predictions would then generalize to this definition.} 
+	\reversemarginpar\marginnote{\#5.2}\rtext{We also plan to explore transfer learning using scene segmentation, as well as the semantic context descriptors and routing scores to show how vector fields can be transferred to novel scenes.
+It appears that the low-order parameterization and unit-length vector field of our model, make it amenable to the methods developed in \cite{Ballan2016}.}
+
+%\section*{Acknowledgments}
+%Kris Kitani was critical in aiding in comparison.
 

### 2017-05-08 cleanpaper/pedestrian_forecasting_iros_2017.tex
created separate technical report and clean version of the paper

--- 2017-05-08 paper/pedestrian_forecasting_iros_2017.tex
@@
 
-\reversemarginpar
-\marginnote{\#1.1}
-\rtext{Many have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
+Many have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
 	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
@@
 	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
-	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.}
+	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.
 
@@
   %In principal, any clustering algorithm could accomplish this.
-  \normalmarginpar
-\marginnote{\#4.1}
-\rtext{To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
 	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
@@
 On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did.
-While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.}
+While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.
 	This clustering of the end-points induces a clustering of the trajectories.
@@
 The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
-	\normalmarginpar \marginnote{\#4.5}
 	\vspace{-0.7cm}
@@
 We implement our model as well as our evaluation code in Python 2.6\footnote{https://github.com/ramvasudevan/iros2017\_pedestrian\_forecasting}.
-\marginnote{\#4.2}
-\rtext{We performed a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
-	We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
-\marginnote{\#4.3}
-\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}
+We performed a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
+	We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.
+Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.
 
@@
 Because our predictions all share the same scale, we amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
-\marginnote{\#4.4} \rtext{In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
+In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
 	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
@@
 	The closer that this AUC is to one, the better the prediction algorithm.
-	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. }
-
-	\rtext{In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
+
+	In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
 	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.
@@
 	Notice that the true positive rate would be zero meaning that the generated set of predictions was not useful.
-	Whereas the MHD would describe this generated set of predictions as informative.}
+	Whereas the MHD would describe this generated set of predictions as informative.
 Our predictor behaves better than any of the other compared methods at moderately large $t$.
@@
 	\includegraphics[width=0.9\linewidth]{./figures/the_results.eps}
-	\caption{\rtext{A comparison of the AUC of the various algorithms.
+	\caption{A comparison of the AUC of the various algorithms.
 Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate.
-We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}}
-	\normalmarginpar\marginnote{\#4.7}
+We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}
 	\label{fig:auc_vs_time}
@@
 	\includegraphics[width=0.9\linewidth]{./figures/MHD_results.eps}
-	\caption{\rtext{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
+	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
 The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
-\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}}
-	\reversemarginpar\marginnote{\#4.8}
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}
 	\label{fig:mhd_vs_time}
@@
 
-\reversemarginpar
-\marginnote{\#4.6}
-\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
+The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.
 This is shown in Table~\ref{tab:time}. 
@@
 %We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop. 
-\reversemarginpar\marginnote{\#5.1}\rtext{Social forces can be incorporated into this model by
+Social forces can be incorporated into this model by
 	%In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  
@@
 	Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995}. 
-	The algorithms for generating real-time predictions would then generalize to this definition.} 
-	\reversemarginpar\marginnote{\#5.2}\rtext{We also plan to explore transfer learning using scene segmentation, as well as the semantic context descriptors and routing scores to show how vector fields can be transferred to novel scenes.
-It appears that the low-order parameterization and unit-length vector field of our model, make it amenable to the methods developed in \cite{Ballan2016}.}
+	The algorithms for generating real-time predictions would then generalize to this definition.
+	We also plan to explore transfer learning using scene segmentation, as well as the semantic context descriptors and routing scores to show how vector fields can be transferred to novel scenes.
+It appears that the low-order parameterization and unit-length vector field of our model, make it amenable to the methods developed in \cite{Ballan2016}.
 

### 2017-05-08 paper/pedestrian_forecasting_iros_2017.tex
created separate technical report and clean version of the paper

@@
 	\end{subfigure}
-	\caption{An illustration of the predictions generated by the various algorithms.
+	\caption{\rtext{An illustration of the predictions generated by the various algorithms.
 In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory.
@@
 Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in \cite{Kitani2012} are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle.
-The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
+The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
 	\normalmarginpar \marginnote{\#4.5}

### 2017-05-08 technicalreport/pedestrian_forecasting_iros_2017-technicalreport.tex
created separate technical report and clean version of the paper

--- 2017-05-08 cleanpaper/pedestrian_forecasting_iros_2017.tex
@@
 \end{thm}
-%\begin{proof}
-%	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
-%	We observe
-%\begin{align*}
-%	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
-%	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
-%	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
-%\end{align*}
-%As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
-%\end{proof}
+\begin{proof}
+	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
+	We observe
+\begin{align*}
+	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
+	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
+	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
+\end{align*}
+As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
+\end{proof}
 
@@
 \end{cor}
-%\begin{proof}
-%	By \eqref{eq:convolve}
-%	\begin{align*}
-%		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
-%	\end{align*}
-%	Substitution of \eqref{eq:approximation 1} yields
-%	\begin{align*}
-%		\begin{split}
-%		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
-%		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
-%		\end{split}
-%	\end{align*}
-%	where the error term is
-%	\begin{align}
-%		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
-%	\end{align}
-%	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
-%	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
-%	\begin{align}
-%		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
-%			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
-%	\end{align}
-%	Implementing the integration over $x_t$ first yields:
-%	\begin{align}
-%		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
-%	\end{align}
-%	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
-%\end{proof}
+\begin{proof}
+	By \eqref{eq:convolve}
+	\begin{align*}
+		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
+	\end{align*}
+	Substitution of \eqref{eq:approximation 1} yields
+	\begin{align*}
+		\begin{split}
+		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
+		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
+		\end{split}
+	\end{align*}
+	where the error term is
+	\begin{align}
+		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
+	\end{align}
+	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
+	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
+	\begin{align}
+		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
+			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
+	\end{align}
+	Implementing the integration over $x_t$ first yields:
+	\begin{align}
+		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
+	\end{align}
+	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
+\end{proof}
 
@@
 \end{thm}
-%\begin{proof}
-%	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
-%	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
-%	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
-%	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
-%	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
-%	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
-%\end{proof}
+\begin{proof}
+	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
+	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
+	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
+	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
+	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
+	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
+\end{proof}
 This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$, which yields the following result:
@@
 \end{thm}
-%\begin{proof}
-%	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
-%	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
-%	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
-%\end{proof}
+\begin{proof}
+	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
+	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
+	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
+\end{proof}
 
@@
 We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}
+%	\normalmarginpar\marginnote{\#4.7}
 	\label{fig:auc_vs_time}
@@
 \cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}
+%	\reversemarginpar\marginnote{\#4.8}
 	\label{fig:mhd_vs_time}

### 2017-05-09 cleanpaper/pedestrian_forecasting_iros_2017.tex
Everything is done

@@
 The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
-\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is known.}
 	\label{fig:mhd_vs_time}

### 2017-05-09 paper/pedestrian_forecasting_iros_2017.tex
Everything is done

@@
 The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
-\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}}
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is known.}}
 	\reversemarginpar\marginnote{\#4.8}

### 2017-06-09 revision/paper/response.tex
Create final version of the paper.

--- 2017-05-08 paper/response.tex
@@
 These additions pushed the paper beyond the eight page limit. 
-As a result we removed all proofs from the resubmitted manuscript, but have included them as instead in a supplementary technical report with our resubmission.
+As a result we removed all proofs from the resubmitted manuscript, but have included them instead in a supplementary technical report with our resubmission.
 This technical report will be posted to our websites if the paper is accepted. 
 %The changes made are presented below as a highlighted manuscript that indicates additions in green and removals in red.
-The changes made are summarized in detail below and a highlighted manuscript is attached that indicates additions in red.
-\textbf{This letter is largely self contained, so references here are to the bibliography in this file}.
+The changes made are summarized in detail in the letter below.
+Following the letter is an annotated version of the paper showing the changes made, followed by the technical report.
+\textbf{This letter is largely self contained, so references here are to the bibliography in this document}.
 % See the response to reviewers for a detailed breakdown of the changes made.
@@
 	\includegraphics[width=0.6\textwidth]{figures/mhd_results.eps}
-	\caption{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point samples from each distribution.
+	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
 The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
@@
 \reversemarginpar\marginnote{\#4.2}
-\response{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
+\response{We performed a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. 
 We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
@@
 \reversemarginpar\marginnote{\#4.1}
-\response{To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
+\response{To do this, we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
 	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ .
@@
 \reversemarginpar\marginnote{\#1.1}
-	\response{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}.
+	\response{Many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
 	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
@@
 	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
-	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact and is the state-of-the-art algorithm for socially based models while also having a quick run time \cite{Alahi2016}.}
+	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact and is the state-of-the-art algorithm for socially based models while also having a rapid computation time \cite{Alahi2016}.}
 
@@
 Though this was not the focus of this paper, the low-dimensional parameterization of the vector fields make transfer learning more amenable. 
-As it is, this is as an immediate avenue for future work which we describe in the following addition to the text: 
+As it is, this is an immediate avenue for future work which we describe in the following addition to the text: 
 \reversemarginpar\marginnote{\#5.2}

### 2017-06-09 revision/paper/response_for_essay.tex
Create final version of the paper.

--- 2017-06-09 revision/paper/response.tex
@@
 
-\title{\LARGE Real-Time Certified Probabilistic Pedestrian Forecasting \\ \textbf{Author's Response} }
+\title{\LARGE A Representative Sample of My Writing \\ \textbf{Author's Response} }
 
@@
 
+I have not uploaded the incorrect document.
+Through this assignment you are seeking a representative sample of my writing.
+I have herein provided a sample of my writing.
+My intent is not to pursue a career in the humanities, rather it is to pursue a Ph.D.
+in the mathematical sciences.
+As a professor in a technical field you advance by writing papers for publication in peer-reviewed journals.
+After submission, manuscripts are returned with criticisms made by experts in your field.
+These criticisms must be addressed by the authors.
+I am co-author on a paper submitted to the IEEE Robotics and Automation Letters.
+After the initial submission of which I penned several sections, the first author left our research group to go into industry.
+I stepped up and did the revisions.
+The proceeding document is the letter I wrote as a supplement to our resubmitted manuscript which explains the changes I made to the paper itself.
+The figures have been ommitted.
+This is an example of my writing proficiency; though not the style you requested, it is the style that matters in my career path.
+If the reader cares to read the full manuscript, I would be more than willing to go over it in person and outline my contributions as well as the argumentative structure of the document, however I have not included it here because of its pending publication.
+
 We would like to begin by thanking the reviewers and the editor for their careful reading and review of our paper.
@@
 %The changes made are presented below as a highlighted manuscript that indicates additions in green and removals in red.
-The changes made are summarized in detail in the letter below.
-Following the letter is an annotated version of the paper showing the changes made, followed by the technical report.
-\textbf{This letter is largely self contained, so references here are to the bibliography in this document}.
+The changes made are summarized in detail in the letter below. \textbf{This letter is largely self contained, so references here are to the bibliography in this document}.
 % See the response to reviewers for a detailed breakdown of the changes made.
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figures/the_results.eps}
 	\caption{A comparison of the AUC of the various algorithms.
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figures/mhd_results.eps}
 	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
@@
   \begin{subfigure}[t]{0.5\textwidth}
-	\centering
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Our Algorithm}   }
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.jpg}
-	\end{minipage}
 
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Kitani et al.}  }
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.jpg}
-	\end{minipage}
-
-	\vspace{0.1cm}
-		\begin{minipage}[c]{0.45cm}
-			\rotatebox{90}{\small{S-LSTM} }  
-		\end{minipage}
-		\begin{minipage}[c]{0.3\linewidth}
-			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=100.jpg}
-		\end{minipage}
-		\begin{minipage}[c]{0.3\linewidth}
-			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=300.jpg}
-		\end{minipage}
-		\begin{minipage}[c]{0.3\linewidth}
-			\includegraphics[width=\linewidth]{./figures/bookstore/lstm_1_2_t=360.jpg}
-		\end{minipage}
-
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Random Walk} }  
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.jpg}
-	\end{minipage}
-
-	\vspace{0.2cm}
-	\hspace{.5cm}
-	\begin{minipage}{0.28\linewidth}
-		\centering
-		$t = 1.66s$
-	\end{minipage}
-	\begin{minipage}{0.3\linewidth}
-		\centering
-		$t = 4.33s$
-	\end{minipage}
-	\begin{minipage}{0.28\linewidth}
-		\centering
-		$t = 12.33s$
-	\end{minipage}
-	\caption{Scene 1}
 %	\caption{
@@
 The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}}
-%	\marginnote{\#4.7}
-	\label{fig:bookstore}
-	\end{subfigure} 
-	\hspace*{0.5cm}
-	\begin{subfigure}[t]{0.5\textwidth}
-	\centering
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Our Algorithm}}   
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.jpg}
-	\end{minipage}
 
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Kitani et al.}} 
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.jpg}
-	\end{minipage}
-
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{S-LSTM}} 
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=100.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=300.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/lstm_1_2_t=380.jpg}
-	\end{minipage}
-
-	\vspace{0.1cm}
-	\begin{minipage}[c]{0.45cm}
-		\rotatebox{90}{\small{Random Walk} }
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.jpg}
-	\end{minipage}
-	\begin{minipage}[c]{0.3\linewidth}
-		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.jpg}
-	\end{minipage}
-
-	\vspace{0.2cm}
-	\hspace{.5cm}
-	\begin{minipage}{0.28\linewidth}
-		\centering
-		$t = 2.33s$
-	\end{minipage}
-	\begin{minipage}{0.3\linewidth}
-		\centering
-		$t = 8.33s$
-	\end{minipage}
-	\begin{minipage}{0.28\linewidth}
-		\centering
-		$t = 12.33s$
-	\end{minipage}	
-	\caption{Scene 2}
 %	\caption{\rtext{An illustration of the predictions generated by the various algorithms.

### 2017-06-09 final/technicalreport/pedestrian_forecasting_iros_2017-technicalreport.tex
Create final version of the paper.

--- 2017-05-08 technicalreport/pedestrian_forecasting_iros_2017-technicalreport.tex
@@
 % paper title
-\title{Real-Time Certified Probabilistic Pedestrian Forecasting}
+\title{\textbf{Technical Report:} \\ Real-Time Certified Probabilistic Pedestrian Forecasting}
 
@@
 The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times.
-\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is unknown.}
+\cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is known.}
 %	\reversemarginpar\marginnote{\#4.8}

### 2017-06-11 final/cleanpaper/pedestrian_forecasting_iros_2017.tex
Do modifications on cleanpaper.

--- 2017-05-09 cleanpaper/pedestrian_forecasting_iros_2017.tex
@@
-\documentclass[letterpaper,10pt,conference]{ieeeconf}
+\documentclass[letterpaper,10pt,conference]{ieeetran}
 %\usepackage{times}
@@
 \IEEEoverridecommandlockouts 
-\overrideIEEEmargins   % Needed to meet printer requirements.
+%\overrideIEEEmargins   % Needed to meet printer requirements.
 \usepackage[T1]{fontenc}
@@
 Other generalizations include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
-Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method.
+Unfortunately the desired posteriors are difficult to compute in closed form, and as a result use sampling based methods.
 The resulting accuracy of such methods, which can come at a large computational expense, can only be known in a probabilistic sense in that the error bounds are themselves random variables.
@@
 	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
-	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.
+	Most recently a method using Long Short-Term Memory (LSTM) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.
 
@@
 the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly.
+	In order to generate true positive rate and false positive rate, a probability distribution has to be thresholded to yield a binary map. 
+	Each chosen threshold creates a point on the ROC curve. We treat every value of each bounding box as a threshold.
 	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
 	The closer that this AUC is to one, the better the prediction algorithm.
+Each method of generating thresholds creates a different AUC, ours in particular corresponds to the AUC variant from \cite{Judd2009}.
 Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
 
-	In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+	In addition, we evaluated the Modified Hausdorff Distance (MHD) \cite{Dubuisson1994} from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
 	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.

### 2017-06-16 final/cleanpaper/pedestrian_forecasting_iros_2017.tex
have reduce the paper down to 8 pages
all the updates on the clean paper will have to be moved to the
technical report as well. Also the citation to the technical report
will have to be added at the end of the first paragraph of Section 3.

@@
 {\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
-\thanks{This work was supported by Ford Motor Company.}
+\thanks{This material is based upon work supported by Ford Motor Company and the National Science Foundation under Grant No. 1562612.}
 }
@@
 
-One limitation of IOC models occurs in cases where there are locally optimal solutions between a start and end goals which can yield non-robust and imprecise behavior.
-This can occur when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
+A limitation of IOC models occurs when there are locally optimal solutions between a start and end goals that can yield non-robust and imprecise behavior.
+This occurs when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
 To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch.
@@
 	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
-	 Recently methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
-	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
-	Most recently a method using Long Short-Term Memory (LSTM) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.
+	 Recently, methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
+	 Others create agent-based models based on Gaussian Processes which perform poorly when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
+	Most recently, a method using Long Short-Term Memory (LSTM) was proposed to learn pedestrian motion models without making assumptions about the manner in which agents interact while having a rapid computation time \cite{Alahi2016}.
 
@@
 \end{align}
-where $X_k$ is a vector-field\footnote{A vector-field is an assignment of a velocity to each position in some space.
-A vector-field on $\mathbb{R}^n$ is a map from $\mathbb{R}^n \to \mathbb{R}^n$.} drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
+where $X_k$ is a vector-field drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
 More specifically, we assume that each $X_k$ has the property that $\| X_k(x) \| = 1$ for all $x \in D$.
@@
 We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
+Due to space limitations, the proofs of all the results presented in this section can be found in \cite{}.
 
@@
 from a storefront to the sidewalk at the edge of a scene.
-On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did.
-While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.
+	On this dataset, distance functions that utilize the entire trajectory did not identify pedestrian intent as well as our method \cite{Morris2009,Lee2007}. 
+	It appears the metrics proposed in \cite{Morris2009} put trajectories that were similar for periods of time together even though they had different intents.
+
 	This clustering of the end-points induces a clustering of the trajectories.
   Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.
-
   For each set $S_k$ we learn a vector-field that is approximately compatible with that set.
@@
 Without this information the implementation of the predictor provided by the authors devolved into a random walk.
-Neither our algorithm nor the random walk were provided this information.
+None of the other tested algorithms required this information.
 
@@
 	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
-	ROC curves which plot the true positive rate (i.e.
-the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e.
-the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly.
-	In order to generate true positive rate and false positive rate, a probability distribution has to be thresholded to yield a binary map. 
-	Each chosen threshold creates a point on the ROC curve. We treat every value of each bounding box as a threshold.
+	ROC curves which plot the true positive rate against the false positive rate evaluate this aforementioned safety criteria for autonomous vehicles exactly. 
+	To generate the true positive rate and false positive rate, a probability distribution has to be thresholded to yield a binary map. 
+	Each chosen threshold creates a point on the ROC curve. 
+	% We treat every value of each bounding box as a threshold.
 	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
 	The closer that this AUC is to one, the better the prediction algorithm. 
-Each method of generating thresholds creates a different AUC, ours in particular corresponds to the AUC variant from \cite{Judd2009}.
-Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
+	%Each method of generating thresholds creates a different AUC, ours in particular corresponds to the AUC variant from \cite{Judd2009}. 
+	Figure \ref{fig:auc_vs_time} illustrates the AUC of each algorithm versus time.
 
@@
 We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in two critical areas: 1) prediction speed is efficient to enable real-time control and 2) our error bound on the predictions accurately reflects potential paths for the human.
-These two features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
+These two features are critical in robotic applications.
+%, and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
 

### 2017-06-17 final/technicalreportarxiv/pedestrian_forecasting_iros_2017-technicalreport.tex
Revision with Citation.

--- 2017-06-09 final/technicalreport/pedestrian_forecasting_iros_2017-technicalreport.tex
@@
 %\usepackage{multicol}
-\usepackage[bookmarks=true]{hyperref}
-\newcommand{\Ram}[1]{\textcolor{red}{#1}}
-\newcommand{\Hoj}[1]{\textcolor{blue}{#1}}
-\newcommand{\rtext}[1]{\textcolor{red}{#1}}
+\usepackage{hyperref}
 
@@
 % paper title
-\title{\textbf{Technical Report:} \\ Real-Time Certified Probabilistic Pedestrian Forecasting}
+\title{Technical Report for Real-Time Certified Probabilistic Pedestrian Forecasting}
 
@@
 	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
-	 Recently methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
-	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
-	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.
+	Recently, methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
+	 Others create agent-based models based on Gaussian Processes which perform poorly when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}.
+	Most recently, a method using Long Short-Term Memory (LSTM) was proposed to learn pedestrian motion models without making assumptions about the manner in which agents interact while having a rapid computation time \cite{Alahi2016}.
 
@@
 from a storefront to the sidewalk at the edge of a scene.
-On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did.
-While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.
+	 On this dataset, distance functions that utilize the entire trajectory did not identify pedestrian intent as well as our method \cite{Morris2009,Lee2007}.
+It appears the metrics proposed in \cite{Morris2009} put trajectories that were similar for periods of time together even though they had different intents.
 	This clustering of the end-points induces a clustering of the trajectories.
@@
 	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
-	ROC curves which plot the true positive rate (i.e.
-the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e.
-the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly.
+	ROC curves which plot the true positive rate against the false positive rate evaluate this aforementioned safety criteria for autonomous vehicles exactly.
+To generate the true positive rate and false positive rate, a probability distribution has to be thresholded to yield a binary map.
+Each chosen threshold creates a point on the ROC curve.
 	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
-	The closer that this AUC is to one, the better the prediction algorithm.
+	The closer that this AUC is to one, the better the prediction algorithm. We treat every value of each bounding box as a threshold.
 	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.
 
-In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
+In addition, we evaluated the Modified Hausdorff Distance (MHD) \cite{Dubuisson1994} from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are.
 	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{./figures/the_results.eps}
+	\includegraphics[width=0.9\linewidth]{./figures/the_results.pdf}
 	\caption{A comparison of the AUC of the various algorithms.
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{./figures/MHD_results.eps}
+	\includegraphics[width=0.9\linewidth]{./figures/mhd_results.pdf}
 	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution.
@@
 We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in two critical areas: 1) prediction speed is efficient to enable real-time control and 2) our error bound on the predictions accurately reflects potential paths for the human.
-These two features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
+These two features are critical in robotic applications. %
+and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
 

### 2017-06-17 final/cleanpaper/pedestrian_forecasting_iros_2017.tex
Revision with Citation.
More modifications.

@@
 \author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
- \thanks{ H.O. Jacobs, O. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
+ \thanks{ H.O. Jacobs, O.K. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
 {\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
@@
 We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
-Due to space limitations, the proofs of all the results presented in this section can be found in \cite{}.
+Due to space limitations, the proofs of all the results presented in this section can be found in \cite{Jacobs2017}.
 
@@
 	Each chosen threshold creates a point on the ROC curve. 
-	% We treat every value of each bounding box as a threshold.
-	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
-	The closer that this AUC is to one, the better the prediction algorithm. 
+	%We treat every value of each bounding box as a threshold.
+	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor \cite{Judd2009}.  
+	The closer that this AUC is to one, the better the prediction algorithm .
 	%Each method of generating thresholds creates a different AUC, ours in particular corresponds to the AUC variant from \cite{Judd2009}. 

### 2017-06-18 final/cleanpaper/pedestrian_forecasting_iros_2017.tex
Fixed Judd citation.
modified the citation to the tech report

@@
 	%We treat every value of each bounding box as a threshold.
-	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor \cite{Judd2009}.  
-	The closer that this AUC is to one, the better the prediction algorithm .
+	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor \cite{judd2009}.  
+	The closer that this AUC is to one, the better the prediction algorithm.
 	%Each method of generating thresholds creates a different AUC, ours in particular corresponds to the AUC variant from \cite{Judd2009}. 

### 2017-06-20 final/cleanpaper/pedestrian_forecasting_iros_2017.tex
Try to fix type 3
making small changes to the text to make it ready for final version

@@
-\documentclass[letterpaper,10pt,conference]{ieeetran}
+\documentclass[letterpaper,10pt,journal,twoside]{ieeetran}
 %\usepackage{times}
 
-\IEEEoverridecommandlockouts 
+%\IEEEoverridecommandlockouts 
 %\overrideIEEEmargins   % Needed to meet printer requirements.
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
+
+\markboth{IEEE Robotics and Automation Letters. Preprint Version. Accepted June, 2017}
+{Jacobs \MakeLowercase{\textit{et al.}}: Real-Time Certified Probabilistic Pedestrian Forecasting}
 
@@
 %%% first author
-\author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
- \thanks{ H.O. Jacobs, O.K. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
+\author{Henry O. Jacobs$^{1}$, Owen K. Hughes$^{1}$, Matthew Johnson-Roberson$^{1}$, and Ram Vasudevan$^{1}$
+\thanks{Manuscript received: February 15, 2017; Revised May 8, 2017; Accepted June 5, 2017.}
+\thanks{This paper was recommended for publication by Wan Kyun Chung upon evaluation of the Associate Editor and Reviewers comments. 
+This material is based upon work supported by Ford Motor Company and the National Science Foundation under Grant No. 1562612.}
+ \thanks{$^{1}$H.O. Jacobs, O.K. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
 {\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
-\thanks{This material is based upon work supported by Ford Motor Company and the National Science Foundation under Grant No. 1562612.}
 }
@@
 \end{abstract}
+\begin{IEEEkeywords}
+Optimization and Optimal Control and Human Detection \& Tracking
+\end{IEEEkeywords}
 
@@
 
-Autonomous systems are increasingly being deployed in and around humans. 
+\IEEEPARstart{A}{utonomous} systems are increasingly being deployed in and around humans. 
 The ability to accurately model and anticipate human behavior is critical to maximizing safety, confidence, and effectiveness of these systems in human-centric environments. 
@@
 	\label{fig:gates-1-2}
-	\vspace*{-0.7cm}
+%	\vspace*{-0.7cm}
 \end{figure}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[b]
+
 		  \caption{Algorithm to compute  $\rho_t$ for each $t \in  \{\Delta t, 2 \Delta t, \dots, N_t \Delta t$\}.}
@@
 The algorithm in \cite{Alahi2016} is confident and close to the trajectory at small times, but their lack of a motion model causes their prediction to compress into a point at intermediate time scales.}
-	\vspace{-0.7cm}
+%	\vspace{-0.7cm}
 \end{figure*}
@@
 	\end{center}
-	\vspace*{-0.6cm}
+%	\vspace*{-0.6cm}
 \end{table}
@@
 	\label{fig:auc_vs_time}
-	\vspace*{-0.75cm}
+%	\vspace*{-0.75cm}
 \end{figure}
@@
 	\label{fig:mhd_vs_time}
-	\vspace*{-0.7cm}
+%	\vspace*{-0.7cm}
 \end{figure}
